### 2021-04-22 article/arxiv.tex
some words

\documentclass{nature}
\bibliographystyle{naturemag}

\usepackage{amssymb}
\usepackage{graphicx}
\graphicspath{{figures/}}
%\usepackage{lineno}
%\linenumbers

% #magic #hack to make figures appear
\makeatletter
\let\saved@includegraphics\includegraphics
\AtBeginDocument{\let\includegraphics\saved@includegraphics}
\renewenvironment*{figure}{\@float{figure}}{\end@float}
\makeatother

\usepackage{color}
\newcommand{\todo}[1]{\textcolor{blue}{#1}}
\newcommand{\arxiv}{arXiv}

% I don't suggest this title; this is a placeholder.
\title{What pandemic?}

\author{Andrew R. Casey$^{\ast,1,2,}$,
        Ilya Mandel$^{\ast,1,2}$,
        and friends?
}

\begin{document}

\maketitle

\begin{affiliations}
	\item School of Physics \& Astronomy, Monash University, Clayton 3800, Victoria, Australia
	\item Center of Excellence for Astrophysics in Three Dimensions (ASTRO-3D), Australia
\end{affiliations}

\begin{abstract}
	% Context.
	`\emph{Publish or perish}' describes the pressure for academics to publish research to ensure a successful career in academia.
	% Define the gap, or question.
	With a global pandemic that has changed how businesses operate, has it also changed the academic publishing system?
	% Here we show.
	Here we show that there is almost no change in expected publishing rates due to the COVID-19 pandemic.
	The exception is biology, where an increase in pre-prints is driven by authors who are not established biologists.
	% Additional conclusions.
	Long-term trends over 14 years show that academics are writing more papers, with more co-authors, and those papers are getting longer. 
	However, the increased quantity does not reflect quality: we find evidence that the level of intellectual contribution needed to warrant co-authorship on a paper has dropped in astronomy, \todo{and elsewhere?}. 
	We find that more leading and co-author papers in the first few years of an academic's career increases their chances to stay in academia, affirming the `\emph{publish or perish}' mantra, but only to a point: after an early career researcher has published enough papers, their chances of continuing in academia remain flat. At that point, they should stop writing papers and do something else.
\end{abstract}

Peer-reviewed publications are the primary measure of productivity in academia. The \arxiv\ is a distribution service for research publications before they are printed in a journal (i.e., a pre-print). A pre-print on the \arxiv\ does not ensure that the contents have already passed the peer-review process, but most material on the \arxiv\ eventually goes through peer-review because it is now standard in many research fields to post pre-prints to the \arxiv\ either during or after the peer-review process.

We retrieved metadata for 1.38 million pre-prints posted on the \arxiv\ between April 2007 and December 2020. The metadata includes the creation date, research field(s), title, author name(s), abstract, and other miscellaneous information. These data show that there is an increasing number of publications each year in nearly every field  (Figure~\ref{fig:number-of-publications-with-time}). Academics are writing more papers, and with more co-authors (Figure~\ref{fig:number-of-authors-with-time}), with a steady decline of single author papers in every field. All papers are getting longer, too: the mean abstract length is increasing, and even the shortest abstracts each year are getting longer (Figure~\ref{fig:abstract-length}). These long-term trends are relatively predictable between years, allowing us to measure changes in academic productivity due to the current pandemic, and variations in authorship networks with time.

In most fields there is no impact on pre-print rates due to the COVID-19 pandemic. We used the number of publications from January 2008 to January 2020 to predict the expected number of publications in 2020 in a given field (see Methods). The true number of publications in 2020 agree excellently with the predictions in most fields. The most significant deviation is biology, where 2020 saw \todo{X} pre-prints posted above expected ($X^{X}_{X}$ expected, $Y$ observed). The most frequent terms in the abstracts of those pre-prints relate to the SARS-COV2 disease.

An increase in biology research during a global pandemic is not so surprising. However, the \arxiv\ data suggests that most of these biology pre-prints are not written by established biologists. The number of unique author names with time (Figure~\ref{fig:unique-authors-with-time}) shows that there is a steady increase in the number of uniquely named authors per field. In 2020, that number spiked as many `new' authors began posting biology pre-prints to the \arxiv. It is unlikely that this spike can solely be explained by a sudden burst of research productivity by very large collaborations, newly formed to tackle the impending pandemic. In Figure~\ref{fig:new-authors-with-time} we show the number of authors who first appear on the \arxiv\ with time. Since our data set begins in 2007, there is a sudden influx of author names who first appear, before there is a steady rise in unique authors as new researchers join the field. In quantitative biology, that steady rise turns to a spike in 2020 (Figure~\ref{fig:new-authors-with-time}) as a sudden influx of authors post pre-prints who have never appeared on a biology pre-print until 2020.  This cohort of authors are not established biologists, and many of them did not collaborate with a biologist. In Figure~\ref{fig:papers-entirely-of-new-authors} we show the number of pre-prints where \emph{all} authors on a paper have never appeared before in the literature of that field. The spike in biology pre-prints in 2020 remains present, representing an influx of pre-prints without any biology expertise. Since it is unlikely that the world-wide number of research-trained quantitative biologists increased overnight, we conclude that \todo{half?} of the quantitative biology pre-prints posted in 2020 are authored by researchers trained in other fields.

\section*{Longevity and authorship valuation}

%	\item Is this increased quantity of publications, and length of publications, a measure of increased quality of research?

We find that the number of authors on a paper has tended to increase in most fields between 2007 and 2020 (Figure~\ref{fig:number-of-authors-with-time}). These trends have been explained by the increasing complexity of research projects that requires expertise from different areas [REF]. If the increasing collaboration size is entirely due to research expertise, then we would expect that the expanding network size would mostly be reflected in established researchers that are experts in their field. Instead, we find that the minimum amount of work necessary to warrant co-authorship has decreased in some fields, which \todo{mostly} accounts for the increase in collaboration sizes with time.

Consider a motivated PhD supervisor with good intentions. If the supervisor encourages their student to contribute to a paper where the supervisor is already involved (but the supervisor is not leading), then the supervisor will be helping to improve the student's visibility (and perceived productivity) in an already competitive landscape. This is a common scenario in academia: an established researcher will assign some task to a PhD student, and the outcome of that task will contribute to a paper in preparation. As a consequence, both the student and supervisor are rightfully listed as co-authors instead of just the supervisor. If the valuation of `intellectual contribution necessary to warrant publication' remained stable with time, then we should expect that the number of co-author publications that overlap with the PhD supervisor(s) will also remain stable with time.

We identified two cohorts of early career researchers (ECRs) in astronomy that are separated by their career start date by five years (see Methods). The distribution of first-author (leading) publications between these cohorts is indistinguishable (\todo{some metrics}, Figure~\ref{fig:longevity}). The relative number of researchers with \todo{up to 5} co-author publications is also the same between both cohorts, but the more recent cohort has a long-tail of ECRs with \todo{X or more} co-author publications. These researchers would appear to be more productive, and to have a more extensive network of collaborators. When we separate the co-author publications to those that do, or do not, include the inferred PhD supervisor(s), it is clear that the increased number of co-author publications is not due to network expansion: those ECRs are publishing with their PhD supervisors and not independently expanding their network. While it is impossible to reliably identify a causal link here, it would appear that those ECR students with increased co-author publications, that overlap heavily with their supervisors, is perhaps due to the supervisors good intention of helping the student expand their network. While individual intentions are good, the net behavioural effect of this as a research field is that the intellectual merit necessary for publication can decrease on average.

\todo{Check whether the increased collaboration size is driven by experts being co-authors, or ECRs as being co-authors}

The data suggest that those extra publications has little or no impact on an ECR's likelihood of remaining in academia. We identified astronomy ECRs that started between 2010-2012 and separated them into two cohorts: those that continued to publish in astronomy, and those that did not. For every first-author paper written by an ECR in the first \todo{X} years, their likelihood of continuing in academia increased. However, this effect flattens out at \todo{$N \approx 6$}: an ECR with 6 first author publications is just as likely to continue in academia as an ECR with 10 first author publications. A similar effect is seen in co-author publications, although the rate is lower. These two effects are related: an ECR with 10 first-author papers and no co-author papers may look as employable (or un-employable) as an ECR with 10 co-author papers and no first-author papers, but for different reasons. Matching these cohorts with citation, network, and institution metrics would likely reveal a more predictive model for whether an individual ECR will remain in academia. But as a cohort, the probability of continuing in astronomy research flattens out at about 50\%: for astronomy ECRs is the top echelon of productivity, half of them will still leave academia.

%\begin{itemize}
%	\item Publish or perish culture to publish more papers per person.
%	\item There are more and more papers per year in each field.
%	\item There is a need to network, both for increasing complexity of projects and to expand connections.
%	\item Single author papers decreasing in every field.
%	\item Those papers are getting longer, too.
%	\item Taking abstract length as a proxy for the length of the paper, the abstract length is increasing in every field.
%	\item Even the 'shortest acceptable' abstract is getting longer per field with time.
%	\item Is this increased quantity of publications, and length of publications, a measure of increased quality of research?
%	\item Is increased network a representation of more collaborative nature, or 
%	\item Authorship currency (need to re-run analysis for all fields and make figures)
%	\item Longevity analysis
%	\item Stop publishing after X papers.
%	\item No apparent change in publication rates due to the pandemic, except:
%	\item Biology, which seems to be driven from people outside of biology
%\end{itemize}

% Figures needed:
% 1. abstract length versus time for all fields
% 2. number of Nth author papers versus time for all fields
% - Number of unique authors per field
% - In biology:
% 	- number of unique authors with time
% 	- number of 'new appearing' authors with time
% 	- number of leading/trailing authors with time
% - Number of pre-prints posted per field with time (month-long granularity)
% - Authorship currency histograms, showing distribution of first, nth, nth-with-overlap, nth-without-overlap
% 	across time in astro-ph, and other fields. 
% - Likelihood of longevity > X years given number of papers posted in first Y years, in astro-ph and other fields.

\section*{References}

\begin{thebibliography}{30}
%\bibitem{Iben_1967} Iben, I., Jr. Stellar Evolution. VI. Evolution from the Main Sequence to the Red-Giant Branch for Stars of Mass 1 $M_{sun}$, 1.25 $M_{sun}$, and 1.5 $M_{sun}$. \emph{Astrophys. J.} \textbf{147}, 624, doi:10.1086/149040 (1967).
\end{thebibliography}

\bibliographystyle{naturemag}

\begin{addendum}
\item[Supplementary Information] ~\\Supplementary information is linked to the online version of the paper at www.nature.com/nature.
\item[Acknowledgements] %\todo{acks}
 \item[Author Information] Reprints and permissions information is
   available at www.nature.com/reprints. The authors declare that they
   have no competing financial interests. Correspondence and requests
   for materials should be addressed to andrew.casey@monash.edu
\end{addendum}

\begin{methods}

\section*{Data retrieval}

The \arxiv\ supports a protocol to access metadata for individual pre-prints, given their identifier. An identifier is defined by the year and month that the pre-print was posted, and the number of pre-prints already posted in that month across all fields. We first retrieved the total number of pre-prints posted to the \arxiv\ each month [REF], and used this to generate all possible identifiers. For each identifier we retrieved the title, author name(s), abstract, research field(s), and the date the pre-print was first posted. For pre-prints posted to \arxiv\ prior to 2007, the primary research category is needed to generate the \arxiv\ identifier. Since this is not readily available for all pre-prints, we chose to exclude pre-prints posted earlier than 2007. This data set includes metadata for 1,379,332 pre-prints posted between 2007-03-30 and 2020-12-30, inclusive.

\section*{Long-term modelling of pre-print counts}

Use a gaussian process.

\section*{Uniquely identifying authors}

The metadata available to us does not include institutional affiliations. In many cases the author names are abbreviated by the submitting author, making it impossible to discern whether `A.~B.~Smith' and `A.~Smith' refer to the same person, even if they are publishing in the same field.

\section*{Code availability}
Sure.

\end{methods}

%\section*{Additional References}
%\begin{thebibliography}{1}
%\bibitem[31]{LAMOST} Luo, A.-L. \emph{et al}. The First Data Release (DR1) of the LAMOST general survey. \emph{Res. Astron. Astrophys.} \textbf{15}, 1095, doi:10.1088/1674-4527/15/8/002 (2015).
%\end{thebibliography}

\bibliographystyle{naturemag}

\end{document}

### 2021-05-06 article/arxiv.tex
article draft

@@
 
+% TODO:
+% [ ] Re-make number of pre-prints posted per sub-field of q-bio into one panel.
+% [ ] Re-make number of pre-prints in q-bio by author as linear.
+% [ ] Nice-ify all figures.
+% [ ] Poisson process model predictions for all primary research fields.
+
 \usepackage{color}
-\newcommand{\todo}[1]{\textcolor{blue}{#1}}
+
+\usepackage[dvipsnames]{xcolor}
+
+\definecolor{mygray}{gray}{0.6}
+
+\newcommand{\todo}[1]{\textcolor{gray}{#1}}
 \newcommand{\arxiv}{arXiv}
@@
         Ilya Mandel$^{\ast,1,2}$,
+        Prashun?, 
         and friends?
@@
 	% Context.
-	`\emph{Publish or perish}' describes the pressure for academics to publish research to ensure a successful career in academia.
+	`\emph{Publish or perish}' is an expression that refers to the pressure placed on academics to consistently publish research to ensure a successful career in academia.
 	% Define the gap, or question.
@@
 	% Here we show.
-	Here we show that there is almost no change in expected publishing rates due to the COVID-19 pandemic.
-	The exception is biology, where an increase in pre-prints is driven by authors who are not established biologists.
-	% Additional conclusions.
-	Long-term trends over 14 years show that academics are writing more papers, with more co-authors, and those papers are getting longer. 
-	However, the increased quantity does not reflect quality: we find evidence that the level of intellectual contribution needed to warrant co-authorship on a paper has dropped in astronomy, \todo{and elsewhere?}.
-	We find that more leading and co-author papers in the first few years of an academic's career increases their chances to stay in academia, affirming the `\emph{publish or perish}' mantra, but only to a point: after an early career researcher has published enough papers, their chances of continuing in academia remain flat.
-At that point, they should stop writing papers and do something else.
+	Here we show that academics are collectively posting just as many publications as if there were no pandemic. 
+	This remains true across nearly all fields of research, including those focussed on theoretical and experimental research.
+	The most significant change in academic publishing is seen in biology, where there was a doubling of pre-prints posted in 2020. 
+	However, most of these extra pre-prints are not written by, or with, biologists.
+	\todo{At least half} of the excess biology pre-prints are entirely authored by researchers from other fields, primarily physicists.
+	No broad impact is found between theoretical and experimental research.
+	Reduced access to experimental laboratories seems to have motivated researchers to publish more papers (overall) based on work that was already in preparation.
+The full impact of the pandemic on academic publishing may not yet be fully realised.
 \end{abstract}
@@
 The \arxiv\ is a distribution service for research publications before they are printed in a journal (i.e., a pre-print).
-A pre-print on the \arxiv\ does not ensure that the contents have already passed the peer-review process, but most material on the \arxiv\ eventually goes through peer-review because it is now standard in many research fields to post pre-prints to the \arxiv\ either during or after the peer-review process.
+A pre-print on the \arxiv\ does not ensure that the contents have already passed peer-review, but most material on the \arxiv\ eventually goes through peer-review because it is now standard in many research fields to post pre-prints to the \arxiv\ either during or after the peer-review process.
+For this reason, the number of pre-prints posted to the \arxiv\ reflects the number of peer-reviewed publications written at any time.
 
@@
 The metadata includes the creation date, research field(s), title, author name(s), abstract, and other miscellaneous information.
-These data show that there is an increasing number of publications each year in nearly every field  (Figure~\ref{fig:number-of-publications-with-time}).
-Academics are writing more papers, and with more co-authors (Figure~\ref{fig:number-of-authors-with-time}), with a steady decline of single author papers in every field.
-All papers are getting longer, too: the mean abstract length is increasing, and even the shortest abstracts each year are getting longer (Figure~\ref{fig:abstract-length}).
-These long-term trends are relatively predictable between years, allowing us to measure changes in academic productivity due to the current pandemic, and variations in authorship networks with time.
+These data show that there is an increasing number of publications each year in nearly every field  (Figure~\ref{fig:pre-prints-segmented-by-field}).
+These long-term trends are relatively predictable from year to year, allowing us to measure any change in academic productivity due to the current pandemic.
 
-In most fields there is no impact on pre-print rates due to the COVID-19 pandemic.
+In most fields there was no impact on pre-print rates due to the COVID-19 pandemic.
 We used the number of publications from January 2008 to January 2020 to predict the expected number of publications in 2020 in a given field (see Methods).
-The true number of publications in 2020 agree excellently with the predictions in most fields.
-The most significant deviation is biology, where 2020 saw \todo{X} pre-prints posted above expected ($X^{X}_{X}$ expected, $Y$ observed).
-The most frequent terms in the abstracts of those pre-prints relate to the SARS-COV2 disease.
+The listed number of publications in 2020 agree excellently with the model predictions in most fields.
+The most significant deviation is in quantitative biology (\arxiv\ code `q-bio'), where 2020 saw about double the number of pre-prints than expected (\todo{$X^{X}_{X}$} expected, \todo{$Y$} observed).
+The increase in quantitative biology is dominated by  pre-prints in the sub-field of \emph{Populations and Evolution} (q-bio.PE; Figure~\ref{fig:q-bio-pre-prints-with-entirely-new-authors}), and the most frequent abstract terms relate to the SARS-COV-2 disease, the COVID-19 virus, or the associated pandemic.
 
-An increase in biology research during a global pandemic is not so surprising.
-However, the \arxiv\ data suggests that most of these biology pre-prints are not written by established biologists.
-The number of unique author names with time (Figure~\ref{fig:unique-authors-with-time}) shows that there is a steady increase in the number of uniquely named authors per field.
-In 2020, that number spiked as many `new' authors began posting biology pre-prints to the \arxiv.
-It is unlikely that this spike can solely be explained by a sudden burst of research productivity by very large collaborations, newly formed to tackle the impending pandemic.
-In Figure~\ref{fig:new-authors-with-time} we show the number of authors who first appear on the \arxiv\ with time.
-Since our data set begins in 2007, there is a sudden influx of author names who first appear, before there is a steady rise in unique authors as new researchers join the field.
-In quantitative biology, that steady rise turns to a spike in 2020 (Figure~\ref{fig:new-authors-with-time}) as a sudden influx of authors post pre-prints who have never appeared on a biology pre-print until 2020.
-This cohort of authors are not established biologists, and many of them did not collaborate with a biologist.
-In Figure~\ref{fig:papers-entirely-of-new-authors} we show the number of pre-prints where \emph{all} authors on a paper have never appeared before in the literature of that field.
-The spike in biology pre-prints in 2020 remains present, representing an influx of pre-prints without any biology expertise.
-Since it is unlikely that the world-wide number of research-trained quantitative biologists increased overnight, we conclude that \todo{half?} of the quantitative biology pre-prints posted in 2020 are authored by researchers trained in other fields.
+An increase in biology research during a global pandemic is not surprising.
+However, the data suggests that at \todo{least half} of these additional biology pre-prints were not written by established biologists.
+The number of new authors appearing in any field is representative of how quickly a field is growing.
+The number of new authors in quantitative biology peaked in 2020 (Figure~\ref{fig:new-authors-segmented-by-field}), while the number of new authors in all other fields remained steady.
+The sudden influx of new authors cannot be explained by large newly-formed collaborations working together to tackle the impending pandemic (Figure~\ref{fig:q-bio-pre-prints-segmented-by-author-count}).
 
-\section*{Longevity and authorship valuation}
+The increase in pre-prints (and new authors) is driven by small (1-4) groups of authors who had never before posted pre-prints to quantitative biology before, either as leading- or co-author  (Figure~\ref{fig:pre-prints-with-entirely-new-authors}).
+It is possible that the COVID-19 pandemic inspired hundreds of established biologists, who had never posted pre-prints to the \arxiv\ before, to quickly use these services disseminate their research.
+This can only explain some of the excess biology pre-prints in 2020.
+We examined the number of pre-prints posted in \emph{Populations and Evolution} (q-bio.PE) where those author names had never appeared in any quantitative biology (q-bio) field before, but those names do appear in other \arxiv\ fields (Figure~\ref{fig:q-bio-pre-prints-with-author-overlap}).
+This gives a measure of the number of common names between different fields before the pandemic, which is about 1-3\%.
+In 2020 most of the pre-prints in \emph{Populations and Evolution} that are \emph{entirely} written by `new' authors (non-biologists) are author names from other fields, primarily physicists and mathematicians.
+The fields with the largest number of unique names are, in order: computer science, physics, condensed matter, math, and astrophysics.
+While it is more likely to have names overlapping between biology and computer science (the largest field by unique names) by chance, the newly authored pre-prints in quantitative biology in 2020 are driven by physicists and mathematicians.
+Taking the number of pre-prints in \emph{Populations and Evolution} (q-bio.PE) where no author has appeared before in that sub-field, at least 40\% of these pre-prints are written by researchers from other fields.
+A careful cross-match of biology pre-prints with Google Scholar profiles shows that this statement is not conflated by two researchers with the same name that work in different fields.
+While some biologists continued to post to \arxiv\ for the first time in 2020, the sharp increase in biology pre-prints in 2020 is driven by researchers from other fields.
 
-%	\item Is this increased quantity of publications, and length of publications, a measure of increased quality of research?
+This could be viewed as a good or bad thing, depending on your view.
+Is it still considered multi-disciplinary research -- a research trait fertilised by many institutions worldwide -- to publish in another field without any subject domain experts?
+Are academics spending time trying to advance efforts against the most significant problem to face society this year, or are some jumping on a citation-laden bandwagon?
+The impact of these pre-prints is hard to measure.
+The number of citations are routinely used as a measure of research impact, but that metric becomes less useful in a period where pre-prints are being written (and cross-citing each other) faster than those pre-prints are being read.
 
-We find that the number of authors on a paper has tended to increase in most fields between 2007 and 2020 (Figure~\ref{fig:number-of-authors-with-time}).
-These trends have been explained by the increasing complexity of research projects that requires expertise from different areas [REF].
-If the increasing collaboration size is entirely due to research expertise, then we would expect that the expanding network size would mostly be reflected in established researchers that are experts in their field.
-Instead, we find that the minimum amount of work necessary to warrant co-authorship has decreased in some fields, which \todo{mostly} accounts for the increase in collaboration sizes with time.
+While the number of pre-prints within primary research fields has remained largely unchanged during the pandemic, there have been noticeable changes within specific sub-fields.
+Border restrictions have obviously limited field work and the capacity to attend conferences.
+In the field of high energy physics (lattice), pre-prints tend to peak around December each year as research presented at the International Symposium on Lattice Field Theory conference is posted on \arxiv.
+In 2020 no Lattice Field Theory conference was held\footnote{The last Lattice Field Theory conference was coincidentally held in Wuhan, Hubei, China.}, so no accompanying pre-prints were posted (Figure~\ref{fig:new-authors-segmented-by-field}).
 
-Consider a motivated PhD supervisor with good intentions.
-If the supervisor encourages their student to contribute to a paper where the supervisor is already involved (but the supervisor is not leading), then the supervisor will be helping to improve the student's visibility (and perceived productivity) in an already competitive landscape.
-This is a common scenario in academia: an established researcher will assign some task to a PhD student, and the outcome of that task will contribute to a paper in preparation.
-As a consequence, both the student and supervisor are rightfully listed as co-authors instead of just the supervisor.
-If the valuation of `intellectual contribution necessary to warrant publication' remained stable with time, then we should expect that the number of co-author publications that overlap with the PhD supervisor(s) will also remain stable with time.
+Even limiting access to compute facilities may have had a substantial effect on theoretical research.
+Within the field of computational physics (physics.comp-ph) the first half of 2020 saw a slight increase in pre-prints that was consistent with the rise in recent years.
+By June 2020 that peak began to fall, and continued to do so until the extent of the data (January 2021), where new pre-prints were at the lowest level seen in about five years.
+If this drop is a causal effect of the pandemic, it is the strongest relative decline seen in any sub-field.
 
-We identified two cohorts of early career researchers (ECRs) in astronomy that are separated by their career start date by five years (see Methods).
-The distribution of first-author (leading) publications between these cohorts is indistinguishable (\todo{some metrics}, Figure~\ref{fig:longevity}).
-The relative number of researchers with \todo{up to 5} co-author publications is also the same between both cohorts, but the more recent cohort has a long-tail of ECRs with \todo{X or more} co-author publications.
-These researchers would appear to be more productive, and to have a more extensive network of collaborators.
-When we separate the co-author publications to those that do, or do not, include the inferred PhD supervisor(s), it is clear that the increased number of co-author publications is not due to network expansion: those ECRs are publishing with their PhD supervisors and not independently expanding their network.
-While it is impossible to reliably identify a causal link here, it would appear that those ECR students with increased co-author publications, that overlap heavily with their supervisors, is perhaps due to the supervisors good intention of helping the student expand their network.
-While individual intentions are good, the net behavioural effect of this as a research field is that the intellectual merit necessary for publication can decrease on average.
+Experimental laboratories have been forced to close, or remain open with reduced time and capacity.
+Access to laboratories is critical for condensed matter research, so it is reasonable to expect a sharp decline in pre-prints, but the exact opposite is observed: there were 20\% more pre-prints posted on material science (cond-mat.mtrl-sci) than in previous years (Figure~\ref{fig:winners-and-losers-of-covid}).
+Materials science is a sub-field of condensed matter focusses on laboratory methods and techniques.
+Usually these techniques are developed as part of an ongoing experiment, and then published separately once the primary (`more interesting') research is complete.
+Without access to laboratories, this increase may be a consequence of `writing up what we can': a motivation to publish existing work that would otherwise be neglected due to ongoing laboratory work.
+However, there were no accompanying sharp declines in any other categories of materials science: overall there were more publications across all fields of condensed matter, not fewer.
+Many research projects take months or years to complete.
+It is reasonable to suspect that the full effect of the pandemic has yet to be realised on pre-print counts, while academics publish what they can within ongoing restrictions.
 
-\todo{Check whether the increased collaboration size is driven by experts being co-authors, or ECRs as being co-authors}
+\noindent\todo{physics.soc-ph saw a bump due to COVID-related research}
 
-The data suggest that those extra publications has little or no impact on an ECR's likelihood of remaining in academia.
-We identified astronomy ECRs that started between 2010-2012 and separated them into two cohorts: those that continued to publish in astronomy, and those that did not.
-For every first-author paper written by an ECR in the first \todo{X} years, their likelihood of continuing in academia increased.
-However, this effect flattens out at \todo{$N \approx 6$}: an ECR with 6 first author publications is just as likely to continue in academia as an ECR with 10 first author publications.
-A similar effect is seen in co-author publications, although the rate is lower.
-These two effects are related: an ECR with 10 first-author papers and no co-author papers may look as employable (or un-employable) as an ECR with 10 co-author papers and no first-author papers, but for different reasons.
-Matching these cohorts with citation, network, and institution metrics would likely reveal a more predictive model for whether an individual ECR will remain in academia.
-But as a cohort, the probability of continuing in astronomy research flattens out at about 50\%: for astronomy ECRs is the top echelon of productivity, half of them will still leave academia.
+ \begin{figure}
+     \includegraphics[width=\textwidth]{pre-prints-segmented-by-field}
+     \caption{Most fields saw no change in the number of pre-prints posted due to the COVID-19 pandemic.
+The number of pre-prints continued to increase each year in most fields.
+\todo{The model.} The exception is quantitative biology (q-bio), where the spike in 2020 is in part caused by COVID-19 related pre-prints authored by people who are non-established biologists.}
+     \label{fig:pre-prints-segmented-by-field}
+ \end{figure}
 
-%\begin{itemize}
-%	\item Publish or perish culture to publish more papers per person.
-%	\item There are more and more papers per year in each field.
-%	\item There is a need to network, both for increasing complexity of projects and to expand connections.
-%	\item Single author papers decreasing in every field.
-%	\item Those papers are getting longer, too.
-%	\item Taking abstract length as a proxy for the length of the paper, the abstract length is increasing in every field.
-%	\item Even the 'shortest acceptable' abstract is getting longer per field with time.
-%	\item Is this increased quantity of publications, and length of publications, a measure of increased quality of research?
-%	\item Is increased network a representation of more collaborative nature, or 
-%	\item Authorship currency (need to re-run analysis for all fields and make figures)
-%	\item Longevity analysis
-%	\item Stop publishing after X papers.
-%	\item No apparent change in publication rates due to the pandemic, except:
-%	\item Biology, which seems to be driven from people outside of biology
-%\end{itemize}
+\begin{figure}
+	\includegraphics[width=\textwidth]{q-bio-pre-prints-with-entirely-new-authors}
+	\caption{The number of pre-prints posted in sub-fields of quantitative biology where all authors have never before appeared in the biology literature. \todo{Re-make to single panel.}}
+	\label{fig:q-bio-pre-prints-with-entirely-new-authors}
+\end{figure}
 
-% Figures needed:
-% 1. abstract length versus time for all fields
-% 2. number of Nth author papers versus time for all fields
-% - Number of unique authors per field
-% - In biology:
-% 	- number of unique authors with time
-% 	- number of 'new appearing' authors with time
-% 	- number of leading/trailing authors with time
-% - Number of pre-prints posted per field with time (month-long granularity)
-% - Authorship currency histograms, showing distribution of first, nth, nth-with-overlap, nth-without-overlap
-% 	across time in astro-ph, and other fields. 
-% - Likelihood of longevity > X years given number of papers posted in first Y years, in astro-ph and other fields.
+\begin{figure}
+	\includegraphics[width=\textwidth]{new-authors-segmented-by-field}
+	\caption{The number of new author names appearing in the literature by field.
+Fields established well before 2007 (e.g., astro-ph) show an apparent influx of authors at the time the data starts (2007).
+These author names are dominated by established academics.
+The slow change in new author count after 2010 approximates the net number of new academics joining the field.
+Note the spike in new authors in quantitative biology in 2020.}
+		\label{fig:new-authors-segmented-by-field}
+\end{figure}
+
+  \begin{figure}
+     \includegraphics[width=\textwidth]{q-bio-pre-prints-segmented-by-author-count}
+     \caption{The increase in quantitative biology pre-prints in 2020 cannot be attributed to large collaborations.
+Here we segment q-bio pre-prints by the number of authors, showing that in 2020 a sharp increase was observed for single-author papers and small (2-4 authors) collaborations.}
+     \label{fig:q-bio-pre-prints-segmented-by-author-count}
+  \end{figure}
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{pre-prints-with-entirely-new-authors}
+	\caption{The number of pre-prints appearing per field where all authors have never appeared in the literature for that field before.
+A sharp increase remains present for quantitative biology (q-bio), representing an influx of pre-prints entirely written by authors who had never posted to \arxiv/q-bio before.}
+		\label{fig:pre-prints-with-entirely-new-authors}
+\end{figure}
+
+  \begin{figure}
+     \includegraphics[width=\textwidth]{q-bio-pre-prints-with-author-overlap}
+     \caption{Many physicists and mathematicians posted pre-prints to quantitative biology in 2020 for the first time.
+Pre-prints in \emph{Populations and Evolution} (q-bio.PE) that are written entirely by authors who have never before appeared in quantitative biology (q-bio).
+The area shows the number of pre-prints where those authors have appeared in other fields.
+     }
+   	\label{fig:q-bio-pre-prints-with-author-overlap}
+	\end{figure}
+
+% \begin{figure}
+%     \includegraphics[width=\textwidth]{pre-prints-segmented-by-author-count}
+%     \caption{Each year academics are writing more papers, with more co-authors.
+The typical size of the largest collaboration papers continues to increase.
+Single author papers are becoming rarer with time.}
+%     \label{fig:pre-prints-segmented-by-author-count}
+% \end{figure}
+
+% \begin{figure}
+%     \includegraphics[width=\textwidth]{pre-prints-abstract-length}
+%     \caption{The typical length of pre-print abstracts indicates that academics are writing longer papers (top).
+Even the shortest 10\% of abstracts in a given time bin are increasing in length (bottom).}
+%     \label{fig:pre-prints-abstract-length}
+% \end{figure}
+
+%\begin{figure}
+%	\includegraphics[width=\textwidth]{unique-authors-segmented-by-field}
+%	\caption{The number of unique author names in each field as a function of time.}
+%		\label{fig:unique-authors-segmented-by-field}
+%\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{winners-and-losers-of-covid}
+	\caption{The number of pre-prints in selected sub-fields where the pandemic may have caused a change in publication rates.}
+		\label{fig:winners-and-losers-of-covid}
+\end{figure}
 
@@
 \item[Supplementary Information] ~\\Supplementary information is linked to the online version of the paper at www.nature.com/nature.
-\item[Acknowledgements] %\todo{acks}
+\item[Acknowledgements] Peter Skands (Monash University) 
+						and Ross Young (University of Adelaide) for comments on publication trends.
+
  \item[Author Information] Reprints and permissions information is
@@
 
-The \arxiv\ supports a protocol to access metadata for individual pre-prints, given their identifier.
-An identifier is defined by the year and month that the pre-print was posted, and the number of pre-prints already posted in that month across all fields.
+The \arxiv\ supports an Open Archive Initiative [REF] protocol to access metadata for individual pre-prints, given an identifier.
+A pre-print's identifier is defined by the year and month that the pre-print was posted, and the number of pre-prints already posted in that month (across all fields).
 We first retrieved the total number of pre-prints posted to the \arxiv\ each month [REF], and used this to generate all possible identifiers.
@@
 For pre-prints posted to \arxiv\ prior to 2007, the primary research category is needed to generate the \arxiv\ identifier.
-Since this is not readily available for all pre-prints, we chose to exclude pre-prints posted earlier than 2007.
+Since the primary research category is not readily available for all pre-prints \emph{a priori}, we chose to exclude pre-prints posted earlier than 2007.
 This data set includes metadata for 1,379,332 pre-prints posted between 2007-03-30 and 2020-12-30, inclusive.
@@
 
-Use a gaussian process.
+\todo{Model the long-term paper counts with a Poisson process.}
 
@@
 
-The metadata available to us does not include institutional affiliations.
-In many cases the author names are abbreviated by the submitting author, making it impossible to discern whether `A.~B.~Smith' and `A.~Smith' refer to the same person, even if they are publishing in the same field.
+The metadata available to us does not include institutional affiliations, or identifiers that would uniquely identify an author.
+For these reasons, we have taken steps to minimise the effects of name confusion.
+There are two primary ways that name confusion could impact our inferences.
+In the first scenario, two people with the same name are amalgamated and treated as a single author that is on average twice as productive (or more, for very common names) as other authors.
+In the second scenario, an author will sometimes publish as `A.~B.~Smith', and other times publish as `A.~Smith'.
+A careful exploration of the data shows that this is a very frequent scenario, and if left uncorrected, would appear as many `unique' authors with half as many publications on average.
+
+We have taken a simple approach to address name confusion.
+We first define a unique author by family name and the initial of the first given name (`Family-name, I.'), such that we intentionally group together authors that may share the same initial of their second given name.
+While our approach to name confusion is grossly simple, it is unlikely that these choices have any substantial impact on our inferences.
+Any common name is likely to appear in the literature early in the data set, and will not impact the conclusions we draw about how publishing changed in 2020.
+While common names will appear between different fields (e.g., quantitative biology and physics), by showing this as a function of time we get an intrinsic measure of the name ambiguity between fields, and we see the increase in overlap when a spike in quantitative biology papers occurs.
+Nevertheless, we manually inspected \todo{hundreds} of quantitative biology pre-prints that entirely consisted of `new authors' and cross-referenced these with Google Scholar to confirm that these authors were established researchers in other fields, and not biologists posting pre-prints for the first time.
+
+\section*{Segmenting by research field}
+
+Pre-prints posted to the \arxiv\ can be cross-listed in multiple fields of research, and multiple sub-fields.
+For example, one pre-print may have a primary field of research as stellar astrophysics (astro-ph.SA) and be cross-listed in computer science.
+The corresponding author must always specify the primary and additional fields of research.
+Throughout this work when we segment by primary research field, we take the primary parent research field provided (in this example, astro-ph).
+We ignore the cross-listings, as these are not always provided for pre-prints, and not always appropriate for appropriating research productivity between different fields.
+When we segment pre-prints by sub-field, we similarly take the primary sub-field provided and ignore any cross-listings.
 
 \section*{Code availability}
-Sure.
+The software developed to retrieve and analyse these data is publicly accessible online\footnote{https://github.com/andycasey/arxiv-covid}. All data was sourced from the \arxiv.
 

### 2021-06-03 article/arxiv-ms.tex
new format

\documentclass[a4paper,12pt]{article}
\usepackage[labelfont=bf,font=small]{caption}

\usepackage{geometry}
\geometry{a4paper,left=25mm,top=25mm,right=25mm,bottom=25mm}

\usepackage[pdftex]{graphicx}
\usepackage{url}
% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\let\iint\relax
\let\iiint\relax
\let\iiiint\relax
\let\idotsint\relax
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{pdflscape}
\usepackage[utf8]{inputenc}
\usepackage{xfrac}
\usepackage{color}
\usepackage{rotating}
\usepackage{tabularx}

\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\definecolor{mygray}{gray}{0.6}
\newcommand{\todo}[1]{\textcolor{red}{#1}}
\newcommand{\arxiv}{arXiv}

\usepackage[symbol]{footmisc}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\usepackage{setspace}

\usepackage{lineno}
\linenumbers

\usepackage{amsmath}
%\usepackage[superscript,biblabel]{cite}

\usepackage[resetlabels,labeled]{multibib}
\newcites{main,method}{{},{}} 
\usepackage[superscript,nomove]{cite}

\usepackage{xpatch}
\newcommand\removebibheader
  {\@{}
    {\section*{}%
     \@{}{}%
    }{}{}{}%
  }

\title{The impact on academic productivity owing to the COVID-19 pandemic}

\author{Andrew R. Casey$^{1,2}$\footnote{andrew.casey@monash.edu},
        Ilya Mandel$^{1,2}$,
        Prasun K. Ray$^{3}$\\
\normalsize{$^{1}$School of Physics \& Astronomy, Monash University, Clayton 3800, Victoria, Australia}\\
\normalsize{$^{2}$Center of Excellence for Astrophysics in Three Dimensions (ASTRO-3D), Australia}\\
\normalsize{$^{3}$Department of Mathematics, Imperial College London, London, United Kingdom}
}

\renewcommand*{\thefootnote}{\arabic{footnote}}

% [X] Any figures missing?
% [X] Write about drop in physics.comp-ph and physics.soc-ph.
% [X] All figure captions, update any figures.
% [X] Quantify difference from model.
% [ ] Abstract
% [ ] Read-through and edit.
% [ ] Distribute for comment.

\date{}
\doublespacing
\begin{document}
\newpage
\setcounter{page}{1}
\resetlinenumber[1]

\maketitle

% Conclusions
% -> Most fields show no impact.
% -> Biggest negative impact is in fields where most research is presented at conferences, or experimental sub-fields.
% -> Condensed matter shows up-tick, where pre-prints have focussed on experimental design instead of experimental results.
% -> Biology has a peak, seems to be a lot of biologists posting to arxiv for the first time, or people contributing from other fields.

\noindent \textbf{`\emph{Publish or perish}' is an expression describing the pressure on academics to consistently publish research to ensure a successful career in academia. With a global pandemic that has changed the world, how has it changed academic productivity? Here we show that academics are collectively posting just as many publications as if there were no pandemic.
	% Here we show.
%	Here we show that academics are collectively posting just as many publications as if there were no pandemic. 
%	This remains true across nearly all fields of research, including those focussed on theoretical and experimental research.
%	The most significant change in academic publishing is seen in biology, where there was a doubling of pre-prints posted in 2020. 
%	However, most of these extra pre-prints are not written by, or with, biologists.
%	\todo{At least half} of the excess biology pre-prints are entirely authored by researchers from other fields, primarily physicists.
%	No broad impact is found between theoretical and experimental research.
%	Reduced access to experimental laboratories seems to have motivated researchers to publish more papers (overall) based on work that was already in preparation. The full impact of the pandemic on academic publishing may not yet be fully realised.	
%	Many fields undeterred; posting just as many as before.
%	Some fields show a large increase (q-bio),
%	Others, mostly experimental fields, show a marked decline in preprints.
}\\

%----------------------------------------------------------------------------------------
%	MAIN MATTER OF THE DOCUMENT
%----------------------------------------------------------------------------------------

\noindent Peer-reviewed publications are the most common measure of productivity in academia. The \arxiv\cite{Ginsparg:2011} ({https://arxiv.org}) is a distribution service for research publications before they are printed in a journal (i.e., a pre-print). A pre-print on \arxiv\ does not ensure that the contents have passed peer-review, but most material on \arxiv\ eventually goes through peer-review because it is now standard in many research fields to post to \arxiv\ either during or after the peer-review process \cite{Lariviere:2014}. For this reason, the number of pre-prints posted to \arxiv\ approximates the number of peer-reviewed publications written at any time.

Here we make quantitative comparisons on the number of pre-prints posted to \arxiv\ before and during the pandemic. While we investigate the impact that the COVID-19 pandemic has had on different fields of research, we highlight that we are unable to identify or differentiate between authors and communities whose productivity has been significantly harmed by the pandemic, and those who were largely unscathed. The COVID-19 pandemic has impacted the population in unequal ways\cite{Nicola:2020,Chu:2020,IbnMohammed:2021}. Generational inequality, career stage, personal circumstances, carer responsibilities, work environments, places of employment, and many other factors, all significantly contribute to the disproportionate and unequal impact the pandemic may have had on a scientist's capacity to conduct research. 
%These factors are more important determinants to research capacity than a scientist's field of research. 
While this limited analysis does not address these issues, it is important to consider that while a community as a whole has not yet demonstrated a significant change in productivity, many researchers have faced significant challenges and suffered physically, mentally, emotionally, and professionally.

We retrieved metadata for 1,475,914 pre-prints posted on the \arxiv\ between 1 April 2007 and 31 May 2021. The metadata includes the creation date, research field(s), title, author name(s), abstract, and other miscellaneous information\cite{Clement:2019}. There is an increasing number of pre-prints posted to \arxiv\ each year in nearly every field (Figure~1). These long-term trends are relatively predictable from year to year, allowing us to quantify any change in academic productivity due to the COVID-19 pandemic. We used the number of publications from January 2008 to December 2019 to predict the expected number of monthly pre-prints in 2020 in each field. We modelled the number of monthly pre-prints in each field using a Gaussian Process with a quasi-periodic kernel function\cite{Rasmussen:2006,Ambikasaran:2014}, allowing us to accurately predict long-term trends and periodicity in given fields of research (Figure~1). In nearly all research fields the number of pre-prints posted since January 2020 agree excellently with the model predictions, indicating no collective impact (positive or negative) on academic productivity due to the COVID-19 pandemic.

\noindent\todo{Include a table with each field where we quantify the change?}
% field     & 2020 pre-prints & YoY change from 2019 to 2020 & Predicted pre-prints in 2020 & Number of sigma deviations away from model
%astro-ph    & 14,893    &  +3.4 & 14,948 \pm 655    & -0.1 
%cond-mat    & 16,188    &  +7.0 & 15,098 \pm 544    & +2.0 
%cs          & 54,808    & +24.3 & 50,567 \pm 3153   & +1.3 
%gr-qc       & 3,405     & +12.4 & 3,178 \pm 202     & +1.1 
%hep-ex      & 615       & -19.8 & 797 \pm 131       & -1.4 
%hep-lat     & 394       & -19.3 & 474 \pm 185       & -0.4 
%hep-ph      & 4,586     &  +1.3 & 4,563 \pm 285     & +0.1 
%hep-th      & 3,869     &  +0.9 & 3,971 \pm 262     & -0.4 
%math        & 35,367    &  +5.5 & 35,409 \pm 1974   & -0.0 
%math-ph     & 1,378     &  -6.3 & 1,445 \pm 131     & -0.5 
%nlin        & 885       &  +6.1 & 820 \pm 64        & +1.0 
%nucl-ex     & 499       &  +2.9 & 479 \pm 62        & +0.3 
%nucl-th     & 1,212     &  -3.2 & 1,248 \pm 97      & -0.4 
%physics     & 15,241    & +14.3 & 14,612 \pm 959    & +0.7 
%q-bio       & 2,790     & +49.0 & 1,944 \pm 191     & +4.4 
%q-fin       & 979       & +23.1 & 812 \pm 74        & +2.3 
%quant-ph    & 6,341     & +13.4 & 6,210 \pm 269     & +0.5 
%stat        & 5,180     & +20.2 & 4,719 \pm 509     & +0.9

Border closures and travel restrictions have forced many academic conferences to be held online-only, or to be cancelled. The immediate impact of cancelling a conference is readily apparent in pre-print counts in lattice physics (Figure~1; \arxiv\ subject code \texttt{hep-lat}, see Methods for explanation), where most pre-prints are posted around December each year as conference proceedings from the International Symposium on Lattice Field Theory. In 2020 the conference was cancelled\cite{LatticeConferenceWebsite} and no accompanying pre-prints exist. However, the impact on research from travel restrictions is likely to be much longer than what is represented by the drop in lattice physics pre-prints. Discussions at conferences or collaborative visits frequently spark new research ideas that might lead to a publication many months or years later. 
% TODO: Consider moving this paragraph elsewhere to better align with the long timelines of research.

Experimental research projects often require specialised laboratories, or data to be collected over many years. The pandemic has forced many laboratories to close or operate with restricted access, which could lead to long-lasting delays in ongoing experiments or immediate drops in publication rates in part due to difficulty in accessing completed (or nearly completed) experiment data. There is some evidence of this in pre-print numbers already, with {25\%} fewer pre-prints in high energy physics (\texttt{hep-ex}) in 2020 than expected by our model (observed 615; predicted 797 $\pm$ 131). But declines are not ubiquitous across experimental fields. The closest comparable field of phenomenology in high energy physics (\texttt{hep-ph}) showed no decline. And throughout 2020, experimental research in condensed matter (\texttt{cond-mat}) saw a 2$\sigma$ increase above our model predictions (observed 16,188; predicted 15,098 $\pm$ 544). Segmenting condensed matter pre-prints by research topic shows that most of this increase (in \texttt{cond-mat}) was driven by a 30\% increase in material science pre-prints (\texttt{cond-mat.mtrl-sci}), a sub-field of condensed matter that focusses on laboratory methods and techniques. Despite the COVID-19 pandemic restricting access to essential laboratory equipment, this shift to publish methods and experimental design work -- which might otherwise be neglected due to ongoing laboratory work -- produced an immediate boost in condensed matter research. This is likely an effect with a limited lifespan: eventually, condensed matter (and other experimental) research will require laboratory access to continue publishing.

The field of quantitative biology (\texttt{q-bio}) research showed the largest increase in pre-prints in 2020 above what is expected by our model. There were {2,790} quantitative biology pre-prints in 2020, {50\%} above the previous year, representing a $+4.4\sigma$ deviation from our model predictions (predicted 1,944 $\pm$ 191). This increase is explainable by an increase in COVID-19 related research, as there were 844 quantitative biology pre-prints in 2020 with pandemic-related terms in their title or abstract (see Methods), and just 43 in the decade prior. Indeed, in {April 2020} nearly 60\% of the \arxiv\ pre-prints in quantitative biology were related to the pandemic (Figure~2). Pandemic-related pre-prints also appeared in other fields (computer science, physics, statistics, and quantitative finance, all peaking around {April 2020}, but pandemic-related pre-prints constituted less than 10\% of the pre-prints in these fields. As of mid-2021 the number of monthly pre-prints in quantitative biology has dropped to pre-pandemic levels, although approximately 20\% of these pre-prints are related to the ongoing pandemic, representing a strong shift in ongoing research focus.

An increase in quantitative biology\footnote{Quantitative biology is only a sub-field of biology, and most biology pre-prints are posted to bioRxiv ({https://biorxiv.org}).} research during a global pandemic is unsurprising. However, the \arxiv\ data shows that this increase in pre-prints was not driven by biology researchers posting twice as many pre-prints than usual. In 2020 there was a peak in the number of new authors appearing for the first time in the quantitative biology literature (Figure~3) while the number of new authors in all other fields remained steady. The sudden influx of new authors cannot be explained by large ($>$10) newly-formed collaborations working together to tackle the impending pandemic (Figure~4). The increase in pre-prints (and new authors) is driven by small (1-4) groups of authors, where many had never posted pre-prints to quantitative biology before.

Some of these new authors in quantitative biology have never before appeared in any other \arxiv\ fields, either.
A close examination of these pre-prints reveals that many are established biologists who have not used the \arxiv\ before, and are now doing so presumably to help ensure that their COVID-19 research is more widely available. This represents a genuine increase in new researchers to the \arxiv, and a boost to making quantitative biology research more widely accessible. Of those new authors in quantitative biology who also appear in other \arxiv\ fields, some of this will be due to name confusion (see Methods): where two researchers in different fields share the same publishing name. However, a careful examination of pandemic-related pre-prints in quantitative biology that were posted in 2020 shows that many were indeed written by researchers from other fields (primarily physicists and mathematicians), who had never before posted about quantitative biology. These pre-prints tend to focus on modelling the COVID-19 outbreak using public data sets, rather than quantitative biology research that requires more expert domain knowledge.

%Any contributions by physicists or mathematicians to quantitative biology did not lead to significant drops in pre-prints in physics or maths. Throughout 2020 there were just as many pre-prints posted in physics as our model predicts, but an unusual asymmetry is present: with more pre-prints posted in the first half of 2020 than those in the later half. Other fields show strong seasonality in pre-prints (e.g., \texttt{hep-lat}), but physics does not. This seasonality in 2020 physics pre-prints is also conflated by a some physics sub-fields that saw a short burst of increased productivity early in 2020, before the effects of the pandemic were widespread (Figure~5).

%drastic drop in productivity (Figure~5). Specifically, pre-prints in computational physics (\texttt{physics.comp-ph}) and \emph{Physics and Society} (\texttt{physics.soc-ph}) peaked to their highest level in \todo{March 2020} before declining rapidly to levels not seen since 2018.

% -> timescales.

The COVID-19 pandemic has had disparate impacts on research productivity in different fields. Some fields show a drop in research outputs. Few show an increase, either driven by pandemic-related research or by shifts away from experimental research to topics that do not require ongoing laboratory access. In this study we have only focussed on quantity of publications as a measure for productivity, and not their quality. While citations are the most commonly used measure of impact of academic publications, that metric becomes a more biased statistic when many related pre-prints are all being posted nearly at the same time. 
The relatively long timescales of academic research would suggest that the full impact on academic productivity due to the COVID-19 pandemic is yet to be seen. For those fields without a measurable drop in research outputs yet, it's plausible that many of the pre-prints being posted are research projects that were well underway before the pandemic, with the full impact of academic research yet to be seen.

%Even among the measurable changes that we have discussed, none have fully returned to normality: even while quantitative biology pre-prints have returned to pre-pandemic levels, 20\% of the research in that field is now related to the ongoing pandemic.

%----------------------------------------------------------------------------------------
%	References
%----------------------------------------------------------------------------------------

%\bibliographystylemain{myieeetr2}

%\vspace{-4cm}
%\bibliographyNew{references_new.bib}
%{\removebibheader
 %\bibliographymain{references_new.bib}
%}

%\bibitem{Papics2017}
%{P{\'a}pics}, P.~I., {\em et~al.}, {Signatures of internal rotation discovered
%  in the Kepler data of five slowly pulsating B stars}, {\em \textit{Astron.
%  Astrophys.}}, {\bfseries 598}, A74, 2017.

\begin{thebibliography}{10}

\bibitem{Ginsparg:2011}
{Ginsparg}. P., {ArXiv at 20}, \emph{Nature}, \textbf{476}, 7359, doi:10.1038/476145a \newblock (2011).

\bibitem{Clement:2019}
{Clement, C.~B., Bierbaum, M., O'Keeffe, K.~P., Alemi, A.~A.,}
{On the Use of ArXiv as a Dataset},
Preprint at https://arxiv.org/abs/1905.00075 \newblock (2019).

\bibitem{Lariviere:2014}
%{Larivi{\'e}re, V., Sugimoto, C.~R., Macaluso, B., Milojevi{\'c}, S., Cronin, B., Thelwall, M.,} 
{Larivi{\'e}re, V., et al.,} {arXiv E-prints and the journal of record: An analysis of roles and relationships}, \emph{Journal of the Association for Information Science and Technology}, \textbf{65}, 6, doi:10.1002/asi.23044 \newblock (2014).

\bibitem{IbnMohammed:2021}
{Ibn-Mohammed, T., et al.,} {A critical analysis of the impacts of {COVID}-19 on the global economy and ecosystems and opportunities for circular economy strategies}, \emph{Resources, Conservation and Recycling}, \textbf{164}, 105169, doi:1016/j.resconrec.2020.105169 \newblock (2021).

\bibitem{Nicola:2020}
{Nicola, M., et al.,} {The socio-economic implications of the coronavirus pandemic ({COVID}-19): A review}, \emph{International Journal of Surgery}, \textbf{78}, 185--193, doi:10.1016/j.ijsu.2020.04.018 \newblock (2020).

\bibitem{Chu:2020}
{Yen-Hao Chu, I., Prima, A., Larson, H.~J., Leesa, L.}
{Social consequences of mass quarantine during epidemics: a systematic review with implications for the {COVID}-19 response}, \emph{Journal of Travel Medicine}, \textbf{27}, 7, doi:10.1093/jtm/taaa192 \newblock (2020).

\bibitem{Rasmussen:2006}
{Rasmussen, C.~E., Williams, C.~K.~I.,}
{Gaussian Processes for Machine Learning}, \emph{MIT Press}, ISBN 026218253X \newblock (2016).

\bibitem{Ambikasaran:2014}
{Ambikasaran, S., Foreman-Mackey, D., Greengard, L., Hogg, D.~W., O'Neil, M.,}
{Fast Direct Methods for Gaussian Processes}, Preprint at http://arxiv.org/abs/1403.6015 \newblock (2014).

\bibitem{LatticeConferenceWebsite}
Helmholtz-Institut f\"ur Strahlen- und Kernphysik Indico Service (Indico). 2021. \emph{The 38th International Symposium on Lattice Field Theory (Lattice 2020)}. [online] Available at: {https://indico.hiskp.uni-bonn.de/event/1/} [Accessed 3 June 2021].

%\bibitem{kippenhahn2012}
%{Kippenhahn}, R., {Weigert}, A., and {Weiss}, A., {\em {Stellar Structure and
%  Evolution, Astronomy and Astrophysics Library, Springer Berlin Heidelberg}}.
%\newblock 2012.

\end{thebibliography}

%\end{document}
\newpage
%----------------------------------------------------------------------------------------
%	Figures
%----------------------------------------------------------------------------------------
\begin{center}
\includegraphics[width=0.95\linewidth]{pre-prints-segmented-by-field}
\end{center}

\noindent \textbf{Figure 1:} \textbf{Since the COVID-19 pandemic began there has been no significant change in monthly \arxiv\ pre-prints in most fields (black).} The expected number of pre-prints from our model is shown in red, and the red region shows the uncertainty in that expectation. Each panel shows a primary field of research in \arxiv. \todo{Blue is ARIMA model.}

\newpage

\begin{center}
 \includegraphics[width=\linewidth]{pandemic-related-preprints.pdf}
\end{center}

\noindent \textbf{Figure 2: Pre-prints related to the COVID-19 pandemic peaked in quantitative biology in 2020, accounting for nearly 60\% of quantitative biology pre-prints.} Shown are five other research fields with the highest number of pandemic-related pre-prints in 2020. Monthly pre-prints are shown on the left, and the percent of pandemic-related pre-prints in that research field is shown on the right.

\newpage

\begin{center}
	\includegraphics[width=0.95\linewidth]{new-authors-segmented-by-field}
\end{center}

\noindent \textbf{Figure 3:} \textbf{The number of new authors appearing in \arxiv\ pre-prints per month peaked in quantitative biology in 2020.} Fields well-established before 2007 (e.g., \texttt{astro-ph}) show an apparent influx of `new authors' at the time the data starts: these authors are already established academics. The slow change in new authors after 2012 approximates the net number of new authors joining the field.

\newpage

\begin{center}
     \includegraphics[width=\linewidth]{q-bio-pre-prints-segmented-by-author-count}
\end{center}

\noindent \textbf{Figure 4:} \textbf{The increase in quantitative biology pre-prints in 2020 cannot be attributed to large collaborations.} The coloured regions represent the different author group sizes for quantitative biology pre-prints per month with time.

\newpage

%\begin{center}
%     \includegraphics[width=\linewidth]{winners-and-losers-of-covid}
%\end{center}

%\noindent \textbf{Figure 5:} \textbf{Mat-sci}

\newpage
%----------------------------------------------------------------------------------------
%	Methods
%----------------------------------------------------------------------------------------
\section*{Methods}

\subsection*{Long-term modelling of monthly pre-print counts}

We use monthly pre-print counts per primary research field from January 2007 until December 2019 to make predictions for the number of pre-prints expected per month from January 2020 onwards. We use a Gaussian process with a quasi-periodic kernel covariance function to model long-term trends and seasonal periodicity\cite{Rasmussen:2006,Ambikasaran:2014}. We fix the periodicity hyper-parameter to one year, and optimise all other hyper-parameters. The predictions of the expectation and variance in the number of monthly pre-print counts are conditioned on the optimized hyper-parameters. We fit this model to monthly pre-print counts for every primary research field.

\todo{We also considered an ARIMA model. The results were very similar. The largest difference between predictions is in \texttt{quant-ph}. Conclusions remained the same. Happy to use either, although I think a Gaussian process (which here is similar enough to a Poisson process) is a closer approximation to the data generating process.}

\subsection*{Segmenting by research field}

Pre-prints posted to the \arxiv\ can be listed in a single field of research, or cross-listed in multiple fields. A field is annotated by \texttt{primary.SEC} where the primary field of research has the prefix, and the sub-field of research is represented by a short capitalised suffix. For example, one pre-print may have a primary field of research as stellar astrophysics (\texttt{astro-ph.SA}) and be cross-listed in machine learning (\texttt{stat.ML}). These field(s) of research are supplied by the corresponding author. It is a subjective decision whether to include more than one field of research, or what those fields of research would be. For this reason, throughout this work when we segment by research field we take the primary parent research field provided and ignore any cross-listed fields of research. When we segment pre-prints by sub-field, we similarly take the primary sub-field provided and ignore any cross-listings.

\subsection*{Uniquely identifying authors}

The \arxiv\ metadata available to us does not include institutional affiliations, or identifiers that would uniquely identify an author. For these reasons, we have taken steps to minimise the effects of name confusion. There are two primary ways that name confusion could impact our inferences. In the first scenario, two people with the same name are amalgamated and treated as a single author that is on average twice as productive (or more, for very common names) as other authors. In the second scenario, an author will sometimes publish as `A.~B.~Smith', and other times publish as `A.~Smith'. A careful exploration of the data shows that this is a very frequent scenario, and if left uncorrected, would appear as many `unique' authors with half as many publications on average.

We have taken a simple approach to address name confusion. We first define a unique author by family name and the initial of the first given name (`Family-name, I.'), such that we intentionally group together authors that may share the same initial of their second given name. While our approach to name confusion is grossly simple, it is unlikely that these choices have any substantial impact on our inferences. Any common name is likely to appear in the literature early in the data set, and will not impact the conclusions we draw about how publishing changed in 2020. 
%While common names will appear between different fields (e.g., quantitative biology and physics), by showing this as a function of time we get an intrinsic measure of the name ambiguity between fields, and we see the increase in overlap when a spike in quantitative biology papers occurs. 
%Nevertheless, we manually inspected horiz of quantitative biology pre-prints that entirely consisted of `new authors' and cross-referenced these with Google Scholar to confirm that these authors were established researchers in other fields, and not biologists posting pre-prints for the first time.

\subsection*{Pre-prints related to the COVID-19 pandemic}

We identified a pre-print as being related to the COVID-19 pandemic if the title or abstract contained any of the following four (case-insensitive) terms: `pandemic', `COVID', `SARS-CoV-2', or `lockdown'. Before 2020, these terms rarely appear in the title or abstract of any pre-print on \arxiv\ (Figure~2).

%----------------------------------------------------------------------------------------
%	Data availability
%----------------------------------------------------------------------------------------

\section*{Data availability}

The entire dataset used in this article is from \arxiv. The dataset is curated by Cornell University and hosted by Kaggle (https://www.kaggle.com/Cornell-University/arxiv). We accessed this dataset on 01 June 2020.

%This data set includes the identifier and metadata for all pre-prints posted to \arxiv, and is updated weekly. 
%A pre-print's identifier is defined by the year and month that the pre-print was posted, and the number of pre-prints already posted in that month (across all fields). For each pre-print the metadata includes  the title, author name(s), abstract, research field(s), and the date the pre-print was first posted. Pre-prints posted prior to 2007 use a different identifier scheme, and for this reason we have excluded pre-prints earlier than this date. The data set used for analysis includes \todo{1,379,332} pre-prints posted between \todo{2007-03-30} and \todo{2020-12-30}.

%All data was sourced from Kaggle (https://www.kaggle.com/Cornell-University/arxiv) on 1 June 2020.

\section*{Code availability}
The software developed to retrieve and analyse the data for this projectis available online (https://github.com/andycasey/arxiv-covid).

%----------------------------------------------------------------------------------------
%	Methods references
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	Acknowledgements
%----------------------------------------------------------------------------------------
\section*{Acknowledgements}
We thank Peter Skands (Monash University) 
	 and Ross Young (University of Adelaide) 
for comments on publication trends in high energy physics.

%----------------------------------------------------------------------------------------
%	Author contributions
%----------------------------------------------------------------------------------------

\section*{Author contributions}

All authors contributed to the discussions and have read and iterated upon the text of the final manuscript.

%----------------------------------------------------------------------------------------
%	Competing interests
%----------------------------------------------------------------------------------------
\section*{Competing interests}
The authors declare no competing interests.

%----------------------------------------------------------------------------------------
%	Materials and correspondence
%----------------------------------------------------------------------------------------
\section*{Materials \& Correspondence}
Correspondence and requests for materials should be addressed to the corresponding author Andrew Casey (andrew.casey@monash.edu).

%----------------------------------------------------------------------------------------
%	Additional information
%----------------------------------------------------------------------------------------
\section*{Additional information}

\textbf{Supplementary Information} is available for this paper.

\noindent\textbf{Correspondence and requests for materials} should be addressed to A.~R.~C.

\noindent\textbf{Reprints and permissions information} is available at {\tt\small www.nature.com/reprints}

\end{document}

### 2021-06-03 article/arxiv.tex
things

@@
 \graphicspath{{figures/}}
-%\usepackage{lineno}
-%\linenumbers
+\usepackage{lineno}
+\linenumbers
 
@@
 \makeatother
-
-% TODO:
-% [ ] Re-make number of pre-prints posted per sub-field of q-bio into one panel.
-% [ ] Re-make number of pre-prints in q-bio by author as linear.
-% [ ] Nice-ify all figures.
-% [ ] Poisson process model predictions for all primary research fields.
 
@@
 
-% I don't suggest this title; this is a placeholder.
-\title{What pandemic?}
+%\title{The differentiated impacts on academic productivity owing to the COVID-19 pandemic}
+\title{The impact on academic productivity owing to the COVID-19 pandemic}
 
@@
         Ilya Mandel$^{\ast,1,2}$,
-        Prashun?, 
+        Prasun K. Ray$^{3}$, 
         and friends?
@@
 \begin{document}
+
+% Conclusions
+% -> Most fields show no impact.
+% -> Biggest negative impact is in fields where most research is presented at conferences, or experimental sub-fields.
+% -> Condensed matter shows up-tick, where pre-prints have focussed on experimental design instead of experimental results.
+% -> Biology has a peak, seems to be a lot of biologists posting to arxiv for the first time, or people contributing from other fields.
 
@@
 	\item Center of Excellence for Astrophysics in Three Dimensions (ASTRO-3D), Australia
+	\item Department of Mathematics, Imperial College London, London, United Kingdom
 \end{affiliations}
@@
 	% Context.
-	`\emph{Publish or perish}' is an expression that refers to the pressure placed on academics to consistently publish research to ensure a successful career in academia.
+	`\emph{Publish or perish}' is an expression that refers to the pressure on academics to consistently publish research to ensure a successful career in academia.
 	% Define the gap, or question.
@@
 	% Here we show.
-	Here we show that academics are collectively posting just as many publications as if there were no pandemic. 
-	This remains true across nearly all fields of research, including those focussed on theoretical and experimental research.
-	The most significant change in academic publishing is seen in biology, where there was a doubling of pre-prints posted in 2020. 
-	However, most of these extra pre-prints are not written by, or with, biologists.
-	\todo{At least half} of the excess biology pre-prints are entirely authored by researchers from other fields, primarily physicists.
-	No broad impact is found between theoretical and experimental research.
-	Reduced access to experimental laboratories seems to have motivated researchers to publish more papers (overall) based on work that was already in preparation.
+%	Here we show that academics are collectively posting just as many publications as if there were no pandemic. 
+%	This remains true across nearly all fields of research, including those focussed on theoretical and experimental research.
+%	The most significant change in academic publishing is seen in biology, where there was a doubling of pre-prints posted in 2020. 
+%	However, most of these extra pre-prints are not written by, or with, biologists.
+%	\todo{At least half} of the excess biology pre-prints are entirely authored by researchers from other fields, primarily physicists.
+%	No broad impact is found between theoretical and experimental research.
+%	Reduced access to experimental laboratories seems to have motivated researchers to publish more papers (overall) based on work that was already in preparation.
 The full impact of the pandemic on academic publishing may not yet be fully realised.
+	Many fields undeterred; posting just as many as before.
+	Some fields show a large increase (q-bio),
+	Others, mostly experimental fields, show a marked decline in preprints.
+
 \end{abstract}
 
-Peer-reviewed publications are the primary measure of productivity in academia.
-The \arxiv\ is a distribution service for research publications before they are printed in a journal (i.e., a pre-print).
-A pre-print on the \arxiv\ does not ensure that the contents have already passed peer-review, but most material on the \arxiv\ eventually goes through peer-review because it is now standard in many research fields to post pre-prints to the \arxiv\ either during or after the peer-review process.
-For this reason, the number of pre-prints posted to the \arxiv\ reflects the number of peer-reviewed publications written at any time.
+%Some research fields have shown an explainable drop in research outputs due to the COVID-19 pandemic.
+Others have shown an increase, driven in part by shifts away from experimental research to topics that do not require ongoing access to laboratories.
+One thing in common between pre-prints
 
-We retrieved metadata for 1.38 million pre-prints posted on the \arxiv\ between April 2007 and December 2020.
-The metadata includes the creation date, research field(s), title, author name(s), abstract, and other miscellaneous information.
-These data show that there is an increasing number of publications each year in nearly every field  (Figure~\ref{fig:pre-prints-segmented-by-field}).
-These long-term trends are relatively predictable from year to year, allowing us to measure any change in academic productivity due to the current pandemic.
+%measure that is common among those fields where there is a change in pre-prints is the timescale:
 
-In most fields there was no impact on pre-print rates due to the COVID-19 pandemic.
-We used the number of publications from January 2008 to January 2020 to predict the expected number of publications in 2020 in a given field (see Methods).
-The listed number of publications in 2020 agree excellently with the model predictions in most fields.
+% Like many professions, in 2020 researchers worked remotely, in isolation, while balancing additional responsibilities like home-schooling children.
+Many academics rightfully report feeling their most unproductive.
+
+%Some sub-fields have already seen an explainable drop in research outputs due to the COVID-19 pandemic.
+For those fields without a measurable drop yet, it's plausible that either some academics are at their most productive while many are at their least, and it's plausible that many of the pre-prints being posted now are research projects that were well underway before the pandemic, with the full impact still to come.
+
+%Part of this is due to name confusion, where two researchers in different fields share the same publishing name.
+However, a careful examination of pandemic-related pre-prints in quantitative biology posted in 2020 shows that many were posted by researchers from other fields (primarily physicists and mathematicians), who had never before posted about quantitative biology.
+
+% Here
+
+% the field of quantitative biology peaked in 2020
+
+%established
+
+% much of this increase in pre-prints was not written by a few
+
+%However,\todo{at least half} of these extra biology pre-prints were not written by established biologists.
+The number of new authors in quantitative biology peaked in 2020 (Figure~\ref{fig:new-authors-segmented-by-field}), while the number of new authors in all other fields remained steady.
+
+%However, these authors do appear in other \arxiv\ fields: in 2020 the author names in common between quantitative biology and other \arxiv\ fields spiked (Figure~\ref{fig:q-bio-pre-prints-with-author-overlap}).
+%These authors do appear in other \arxiv\ fields.
+In 2020 many of the pre-prints  in quantitative biology that are \emph{entirely} written by `new' authors (non-biologists) are authors that appear in other fields, primarily physicists and mathematicians.
+A careful cross-match of biology pre-prints with Google Scholar profiles shows that this statement is not conflated by two researchers with the same name that work in different fields.
+While it is possible that the COVID-19 pandemic inspired biologists to use \arxiv\ for the first time, this can only explain some of the excess biology pre-prints in 2020.
+
+% -> How much? A doubling
+% -> Driven entirely by COVID-papers.
+% -> Author make-up.
+
+% -> People wrote papers about COVID in other fields.
+
+%Pre-prints in lattice physics (hep-lat) periodically peak around December each year as conference proceedings from the International Symposium on Lattice Field Theory are posted to \arxiv.
+In 2020 the conference was cancelled, so no accompanying pre-prints exist.
+The impact on research from travel restrictions and cancelled conferences is likely to be much longer term than what is represented by the drop in lattice physics pre-prints.
+Conference discussions or collaborative visits frequently spark new research ideas that might lead to a publication many months later.
+Some research ideas require laboratories or specialised equipment to be built, or data to be collected over years.
+Depending on the research, a single publication may represent the work of weeks, years, or even a decade.
+
+%In some fields there are significant deviations from the model predictions, indicative of the differentiated impact the COVID-19 pandemic has had on academic productivity.
 The most significant deviation is in quantitative biology (\arxiv\ code `q-bio'), where 2020 saw about double the number of pre-prints than expected (\todo{$X^{X}_{X}$} expected, \todo{$Y$} observed).
-The increase in quantitative biology is dominated by  pre-prints in the sub-field of \emph{Populations and Evolution} (q-bio.PE; Figure~\ref{fig:q-bio-pre-prints-with-entirely-new-authors}), and the most frequent abstract terms relate to the SARS-COV-2 disease, the COVID-19 virus, or the associated pandemic.
+The titles and abstracts of these pre-prints shows that the excess biology pre-prints is entirely due to research related to the SARS-CoV-2 disease or the COVID-19 pandemic (Figure~\ref{fig:pandemic-related-preprints}).
+In March 2020, nearly 60\% of pre-prints in quantitative biology were related to the COVID-19 pandemic.
 
-An increase in biology research during a global pandemic is not surprising.
-However, the data suggests that at \todo{least half} of these additional biology pre-prints were not written by established biologists.
-The number of new authors appearing in any field is representative of how quickly a field is growing.
+%, with most pre-prints appearing in the sub-field of \emph{Populations and Evolution} (q-bio.PE; Figure~\ref{fig:q-bio-pre-prints-with-entirely-new-authors}).
+
+%An increase in biology research during a global pandemic is not surprising.
+However,\todo{at least half} of these extra biology pre-prints were not written by established biologists.
 The number of new authors in quantitative biology peaked in 2020 (Figure~\ref{fig:new-authors-segmented-by-field}), while the number of new authors in all other fields remained steady.
 The sudden influx of new authors cannot be explained by large newly-formed collaborations working together to tackle the impending pandemic (Figure~\ref{fig:q-bio-pre-prints-segmented-by-author-count}).
+The increase in pre-prints (and new authors) is driven by small (1-4) groups of authors who had never before posted pre-prints to biology before, either as leading- or co-author (Figure~\ref{fig:pre-prints-with-entirely-new-authors}).
+%However, these authors do appear in other \arxiv\ fields: in 2020 the author names in common between quantitative biology and other \arxiv\ fields spiked (Figure~\ref{fig:q-bio-pre-prints-with-author-overlap}).
+%These authors do appear in other \arxiv\ fields.
+In 2020 many of the pre-prints  in quantitative biology that are \emph{entirely} written by `new' authors (non-biologists) are authors that appear in other fields, primarily physicists and mathematicians.
+A careful cross-match of biology pre-prints with Google Scholar profiles shows that this statement is not conflated by two researchers with the same name that work in different fields.
+While it is possible that the COVID-19 pandemic inspired biologists to use \arxiv\ for the first time, this can only explain some of the excess biology pre-prints in 2020.
 
-The increase in pre-prints (and new authors) is driven by small (1-4) groups of authors who had never before posted pre-prints to quantitative biology before, either as leading- or co-author  (Figure~\ref{fig:pre-prints-with-entirely-new-authors}).
-It is possible that the COVID-19 pandemic inspired hundreds of established biologists, who had never posted pre-prints to the \arxiv\ before, to quickly use these services disseminate their research.
-This can only explain some of the excess biology pre-prints in 2020.
-We examined the number of pre-prints posted in \emph{Populations and Evolution} (q-bio.PE) where those author names had never appeared in any quantitative biology (q-bio) field before, but those names do appear in other \arxiv\ fields (Figure~\ref{fig:q-bio-pre-prints-with-author-overlap}).
-This gives a measure of the number of common names between different fields before the pandemic, which is about 1-3\%.
-In 2020 most of the pre-prints in \emph{Populations and Evolution} that are \emph{entirely} written by `new' authors (non-biologists) are author names from other fields, primarily physicists and mathematicians.
-The fields with the largest number of unique names are, in order: computer science, physics, condensed matter, math, and astrophysics.
-While it is more likely to have names overlapping between biology and computer science (the largest field by unique names) by chance, the newly authored pre-prints in quantitative biology in 2020 are driven by physicists and mathematicians.
-Taking the number of pre-prints in \emph{Populations and Evolution} (q-bio.PE) where no author has appeared before in that sub-field, at least 40\% of these pre-prints are written by researchers from other fields.
-A careful cross-match of biology pre-prints with Google Scholar profiles shows that this statement is not conflated by two researchers with the same name that work in different fields.
-While some biologists continued to post to \arxiv\ for the first time in 2020, the sharp increase in biology pre-prints in 2020 is driven by researchers from other fields.
+%While some biologists continued to post to \arxiv\ for the first time in 2020, the sharp increase in biology pre-prints in 2020 is driven by researchers from other fields.
 
-This could be viewed as a good or bad thing, depending on your view.
+%Physicists and mathematicians devoting time to research topics related to COVID-19  could be good or bad.
+Here we are reporting only on the quantity of pre-prints, which has no bearing on their quality or impact.
+Many COVID-19 pre-prints posted in 2020 by physicists and mathematicians have no expert biologist as a co-author.
 Is it still considered multi-disciplinary research -- a research trait fertilised by many institutions worldwide -- to publish in another field without any subject domain experts?
@@
 
-While the number of pre-prints within primary research fields has remained largely unchanged during the pandemic, there have been noticeable changes within specific sub-fields.
+%in 2020 the author names in common between quantitative biology and other \arxiv\ fields spiked
+
+% rlap in common names between different fields spiked well above the pre-pandemic
+
+%We examined the number of pre-prints posted in \emph{Populations and Evolution} (q-bio.PE) where those author names had never appeared in any quantitative biology (q-bio) field before, but those names do appear in other \arxiv\ fields .
+
+% This gives a measure of the number of common names between different fields before the pandemic, which is about 1-3\%. In 2020 most of the pre-prints in 
+%The fields with the largest number of unique names are, in order: computer science, physics, condensed matter, math, and astrophysics.
+While it is more likely to have names overlapping between biology and computer science (the largest field by unique names) by chance, the newly authored pre-prints in quantitative biology in 2020 are driven by physicists and mathematicians.
+%Taking the number of pre-prints in \emph{Populations and Evolution} (q-bio.PE) where no author has appeared before in that sub-field, at least 40\% of these pre-prints are written by researchers from other fields.
+
+%With physicists and mathematicians seemingly having time to write research papers in quantitative biology, one might expect a corresponding drop in physics or math pre-prints.
+No drop is evident in maths.
+Physics showed no average drop over the year, but this is balanced by an uncharacteristic increase in pre-prints in the first half of 2020 (the same time the quantitative biology peak occurred), and a drop in the second half.
+However, some sub-fields of experimental physics were noticeably impacted by COVID-19.
+%Even limiting access to compute facilities may have had a substantial effect on theoretical research. 
+%The field of computational physics (physics.comp-ph) saw a severe decline in research outputs from June 2020, reducing pre-prints to a level not seen in five years.
+If this drop is a causal effect of the pandemic, it is the strongest relative decline seen in any sub-field.
+%High energy experimental physics pre-prints dropped sharply in January 2020 ($\sim$20\% compared to 2019 levels), and remained low throughout 2020, likely due in part to limited laboratory access.
+Access to laboratories is also critical for condensed matter research, but this field saw a slight increase in total pre-prints, in part driven by a 20\% increase in material science pre-prints (cond-mat.mtrl-sci), a sub-field of condensed matter that focusses on laboratory methods and techniques.
+Without access to laboratories, this increase may be a consequence of `writing up what we can': a motivation to publish existing work that would otherwise be neglected due to ongoing laboratory work.
+
+%While the number of pre-prints within primary research fields has remained largely unchanged during the pandemic, there have been noticeable changes within specific sub-fields.
 Border restrictions have obviously limited field work and the capacity to attend conferences.
@@
 
-Even limiting access to compute facilities may have had a substantial effect on theoretical research.
-Within the field of computational physics (physics.comp-ph) the first half of 2020 saw a slight increase in pre-prints that was consistent with the rise in recent years.
+% Within the field of computational physics (physics.comp-ph) the first half of 2020 saw a slight increase in pre-prints that was consistent with the rise in recent years.
 By June 2020 that peak began to fall, and continued to do so until the extent of the data (January 2021), where new pre-prints were at the lowest level seen in about five years.
@@
 
-Experimental laboratories have been forced to close, or remain open with reduced time and capacity.
-Access to laboratories is critical for condensed matter research, so it is reasonable to expect a sharp decline in pre-prints, but the exact opposite is observed: there were 20\% more pre-prints posted on material science (cond-mat.mtrl-sci) than in previous years (Figure~\ref{fig:winners-and-losers-of-covid}).
-Materials science is a sub-field of condensed matter focusses on laboratory methods and techniques.
-Usually these techniques are developed as part of an ongoing experiment, and then published separately once the primary (`more interesting') research is complete.
-Without access to laboratories, this increase may be a consequence of `writing up what we can': a motivation to publish existing work that would otherwise be neglected due to ongoing laboratory work.
-However, there were no accompanying sharp declines in any other categories of materials science: overall there were more publications across all fields of condensed matter, not fewer.
-Many research projects take months or years to complete.
+%We hypothesise that this is in part attributed to the long timescales of academic research, and (as we discuss below) the differentiated impact the COVID-19 pandemic has had.
+Collectively academics may have published just as many research papers, but this may be driven by some individuals writing more papers, and many individuals writing fewer.
+
+%Depending on the research, a single publication may represent the work of months or years.
+Conference discussions or collaborative visits frequently spark new research ideas that lead to publication a year or more later.
+With travel restrictions introduced early- to mid-2020, any decline in preprints may not be visible in a field for some time.
+
+%Experimental physics fields either showed a drop in total preprint counts, or a shift towards writing methods-based publications that no longer require laboratory access.
+
+%Usually these techniques are developed as part of an ongoing experiment, and then published separately once the primary (`more interesting') research is complete.
+
+% However, there were no accompanying sharp declines in any other categories of materials science: overall there were more publications across all fields of condensed matter, not fewer.
+
+%The impact of reduced travel is most clearly seen in high energy lattice physics preprints, where a yearly conference is held each year and research is presented.
+Instead of running this conference virtually, in 2020 it was cancelled, and no associated pre-prints were posted.
+
+%The field of high energy physics saw an almost immediate reduction of about 20\% fewer pre-prints posted per month, a drop that persisted throughout 2020.
+
+% Many research projects take months or years to complete.
 It is reasonable to suspect that the full effect of the pandemic has yet to be realised on pre-print counts, while academics publish what they can within ongoing restrictions.
 
-\noindent\todo{physics.soc-ph saw a bump due to COVID-related research}
+% because full impact not known yet
+
+%hep ex
+%physics
+%q-fin
+% biology stsuff
+
+%While the number of pre-prints within primary research fields has remained largely unchanged during the pandemic, there have been noticeable changes within specific sub-fields.
+Border restrictions have obviously limited field work and the capacity to attend conferences.
+In the field of high energy physics (lattice), pre-prints tend to peak around December each year as research presented at the International Symposium on Lattice Field Theory conference is posted on \arxiv.
+In 2020 no Lattice Field Theory conference was held\footnote{The last Lattice Field Theory conference was coincidentally held in Wuhan, Hubei, China.}, so no accompanying pre-prints were posted (Figure~\ref{fig:new-authors-segmented-by-field}).
+
+%\noindent\todo{physics.soc-ph saw a bump due to COVID-related research}
 
  \begin{figure}
-     \includegraphics[width=\textwidth]{pre-prints-segmented-by-field}
+ 	\centering
+     \includegraphics[width=0.95\textwidth]{pre-prints-segmented-by-field}
      \caption{Most fields saw no change in the number of pre-prints posted due to the COVID-19 pandemic.
-The number of pre-prints continued to increase each year in most fields.
 \todo{The model.} The exception is quantitative biology (q-bio), where the spike in 2020 is in part caused by COVID-19 related pre-prints authored by people who are non-established biologists.}
@@
 
-\begin{figure}
-	\includegraphics[width=\textwidth]{q-bio-pre-prints-with-entirely-new-authors}
-	\caption{The number of pre-prints posted in sub-fields of quantitative biology where all authors have never before appeared in the biology literature. \todo{Re-make to single panel.}}
-	\label{fig:q-bio-pre-prints-with-entirely-new-authors}
-\end{figure}
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.65\textwidth]{q-bio-pre-prints-with-entirely-new-authors}
+%	\caption{In 2020 there was a sharp increase in \emph{Populations and Evolution} (q-bio.PE; blue) pre-prints that were seemingly written without a biologist.
+Here we show the number of \arxiv\ pre-prints posted per month where \emph{no author} had ever before appeared on a biology pre-print before, grouped by the primary category.
+The number of pre-prints per category remains stable with time, except for \emph{Populations and Evolution} (q-bio.PE) in 2020, which saw an influx of pre-prints where no author had ever appeared on a biology pre-print before.}
+%	\label{fig:q-bio-pre-prints-with-entirely-new-authors}
+%\end{figure}
 
@@
   \begin{figure}
-     \includegraphics[width=\textwidth]{q-bio-pre-prints-segmented-by-author-count}
+  	\centering
+     \includegraphics[width=\textwidth]{pandemic-related-preprints.pdf}
+     \caption{}
+     \label{fig:pandemic-related-preprints}
+       \end{figure}
+
+  \begin{figure}
+  	\centering
+     \includegraphics[width=0.5\textwidth]{q-bio-pre-prints-segmented-by-author-count}
      \caption{The increase in quantitative biology pre-prints in 2020 cannot be attributed to large collaborations.
@@
 
-\begin{figure}
-	\includegraphics[width=\textwidth]{pre-prints-with-entirely-new-authors}
-	\caption{The number of pre-prints appearing per field where all authors have never appeared in the literature for that field before.
+% ARC: no longer needed as covered by Figure 2
+%\begin{figure}
+%	\includegraphics[width=\textwidth]{pre-prints-with-entirely-new-authors}
+%	\caption{The number of pre-prints appearing per field where all authors have never appeared in the literature for that field before.
 A sharp increase remains present for quantitative biology (q-bio), representing an influx of pre-prints entirely written by authors who had never posted to \arxiv/q-bio before.}
-		\label{fig:pre-prints-with-entirely-new-authors}
-\end{figure}
+%		\label{fig:pre-prints-with-entirely-new-authors}
+%\end{figure}
 
-  \begin{figure}
-     \includegraphics[width=\textwidth]{q-bio-pre-prints-with-author-overlap}
-     \caption{Many physicists and mathematicians posted pre-prints to quantitative biology in 2020 for the first time.
+%  \begin{figure}
+%     \includegraphics[width=\textwidth]{q-bio-pre-prints-with-author-overlap}
+%%     \caption{Many physicists and mathematicians posted pre-prints to quantitative biology in 2020 for the first time.
 Pre-prints in \emph{Populations and Evolution} (q-bio.PE) that are written entirely by authors who have never before appeared in quantitative biology (q-bio).
 The area shows the number of pre-prints where those authors have appeared in other fields.
-     }
-   	\label{fig:q-bio-pre-prints-with-author-overlap}
-	\end{figure}
+%     }
+%   	\label{fig:q-bio-pre-prints-with-author-overlap}
+%	\end{figure}
 
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{winners-and-losers-of-covid}
+	\centering
+	\includegraphics[width=0.5\textwidth]{winners-and-losers-of-covid}
 	\caption{The number of pre-prints in selected sub-fields where the pandemic may have caused a change in publication rates.}
@@
 
-The \arxiv\ supports an Open Archive Initiative [REF] protocol to access metadata for individual pre-prints, given an identifier.
+We use the \arxiv\ data set\cite{1905.00075} submitted by Cornell University to the Kaggle website. % https://www.kaggle.com/Cornell-University/arxiv
+This data set includes the identifier and metadata for all pre-prints posted to \arxiv, and is updated weekly.
 A pre-print's identifier is defined by the year and month that the pre-print was posted, and the number of pre-prints already posted in that month (across all fields).
-We first retrieved the total number of pre-prints posted to the \arxiv\ each month [REF], and used this to generate all possible identifiers.
-For each identifier we retrieved the title, author name(s), abstract, research field(s), and the date the pre-print was first posted.
-For pre-prints posted to \arxiv\ prior to 2007, the primary research category is needed to generate the \arxiv\ identifier.
-Since the primary research category is not readily available for all pre-prints \emph{a priori}, we chose to exclude pre-prints posted earlier than 2007.
-This data set includes metadata for 1,379,332 pre-prints posted between 2007-03-30 and 2020-12-30, inclusive.
+For each pre-print the metadata includes  the title, author name(s), abstract, research field(s), and the date the pre-print was first posted.
+Pre-prints posted prior to 2007 use a different identifier scheme, and for this reason we have excluded pre-prints earlier than this date.
+The data set used for analysis includes \todo{1,379,332} pre-prints posted between \todo{2007-03-30} and \todo{2020-12-30}.
+
+\section*{Segmenting by research field}
+
+Pre-prints posted to the \arxiv\ can be cross-listed in multiple fields of research, and multiple sub-fields.
+For example, one pre-print may have a primary field of research as stellar astrophysics (astro-ph.SA) and be cross-listed in computer science.
+These field(s) of research are supplied by the corresponding author.
+It is a subjective decision whether to include more than one field of research, or what those fields of research would be.
+For this reason, throughout this work when we segment by research field we take the primary parent research field provided and ignore any cross-listed fields of research.
+When we segment pre-prints by sub-field, we similarly take the primary sub-field provided and ignore any cross-listings.
 
@@
 
-\todo{Model the long-term paper counts with a Poisson process.}
+We use monthly pre-print counts per primary research field from 2007-01 until 2019-12 to make predictions for the  number of pre-prints expected in 2020.
+We use a Gaussian process with a quasi-periodic kernel () to model.
+We fix the period of the kernel to be 1 year.
 
@@
 
-\section*{Segmenting by research field}
-
-Pre-prints posted to the \arxiv\ can be cross-listed in multiple fields of research, and multiple sub-fields.
-For example, one pre-print may have a primary field of research as stellar astrophysics (astro-ph.SA) and be cross-listed in computer science.
-The corresponding author must always specify the primary and additional fields of research.
-Throughout this work when we segment by primary research field, we take the primary parent research field provided (in this example, astro-ph).
-We ignore the cross-listings, as these are not always provided for pre-prints, and not always appropriate for appropriating research productivity between different fields.
-When we segment pre-prints by sub-field, we similarly take the primary sub-field provided and ignore any cross-listings.
-
 \section*{Code availability}

### 2021-06-04 article/arxiv-ms.tex
updates to preprint

@@
 
-% [X] Any figures missing?
-% [X] Write about drop in physics.comp-ph and physics.soc-ph.
-% [X] All figure captions, update any figures.
-% [X] Quantify difference from model.
-% [ ] Abstract
-% [ ] Read-through and edit.
-% [ ] Distribute for comment.
+% Responses to comments:
+% From Ilya: where fields are defined by primary arXiv category, i.e., chosen by the author(s) at submission (?)
+% --> More description given in Methods
+
+% Outstanding issues to address/close:
+% 1.
+From Ilya: Since ADS indexes arxiv (at least for astro), is it worth checking the fraction of papers from, say, 3 years ago that have only arXiv numbers (never published)?
+As well as those which are published in peer-review journals with no arXiv numbers, for completeness in the other direction?
+
+% 2. From Ilya: "All authors new to arXiv vs. at least one author previously on arXiv but all new to q-bio vs. at least one author previously on q-bio vs. all previously on q-bio."
 
@@
 
-% Conclusions
-% -> Most fields show no impact.
-% -> Biggest negative impact is in fields where most research is presented at conferences, or experimental sub-fields.
-% -> Condensed matter shows up-tick, where pre-prints have focussed on experimental design instead of experimental results.
-% -> Biology has a peak, seems to be a lot of biologists posting to arxiv for the first time, or people contributing from other fields.
-
 \noindent \textbf{`\emph{Publish or perish}' is an expression describing the pressure on academics to consistently publish research to ensure a successful career in academia. 
 With a global pandemic that has changed the world, how has it changed academic productivity? 
-Here we show that academics are collectively posting just as many publications as if there were no pandemic.
-	% Here we show.
-%	Here we show that academics are collectively posting just as many publications as if there were no pandemic. 
-%	This remains true across nearly all fields of research, including those focussed on theoretical and experimental research.
-%	The most significant change in academic publishing is seen in biology, where there was a doubling of pre-prints posted in 2020. 
-%	However, most of these extra pre-prints are not written by, or with, biologists.
-%	\todo{At least half} of the excess biology pre-prints are entirely authored by researchers from other fields, primarily physicists.
-%	No broad impact is found between theoretical and experimental research.
-%	Reduced access to experimental laboratories seems to have motivated researchers to publish more papers (overall) based on work that was already in preparation.
-The full impact of the pandemic on academic publishing may not yet be fully realised.
-%	Many fields undeterred; posting just as many as before.
-%	Some fields show a large increase (q-bio),
-%	Others, mostly experimental fields, show a marked decline in preprints.
+Here we show that academics are posting just as many publications on the \arxiv\ pre-print server as if there were no pandemic.
+However, some immediate impacts are visible in individual research fields.
+Conference cancellations have led to sharp drops in pre-prints, but laboratory closures have had mixed effects.
+While some experimental fields show mild declines in outputs, other fields have increased by writing more method-based papers that are less dependent on ongoing laboratory access.
+The most significant change is a 50\% increase in quantitative biology research, all related to the COVID-19 pandemic.
+Some of these publications are by biologists using \arxiv\ for the first time, and some are written by researchers from other fields (e.g., physicists, mathematicians).
+While quantitative biology pre-prints have returned to pre-pandemic levels, 20\% of the research in this field is now focussed on the COVID-19 pandemic, demonstrating a strong shift in research focus.
 }\\
-
-%----------------------------------------------------------------------------------------
-%	MAIN MATTER OF THE DOCUMENT
-%----------------------------------------------------------------------------------------
 
@@
 %These factors are more important determinants to research capacity than a scientist's field of research. 
-While this limited analysis does not address these issues, it is important to consider that while a community as a whole has not yet demonstrated a significant change in productivity, many researchers have faced significant challenges and suffered physically, mentally, emotionally, and professionally.
+While this limited analysis does not address these issues, it is important to consider that if a community as a whole has not yet demonstrated a significant change in productivity, many researchers have faced significant challenges and suffered physically, mentally, emotionally, and professionally.
 
@@
 We used the number of publications from January 2008 to December 2019 to predict the expected number of monthly pre-prints in 2020 in each field.
-We modelled the number of monthly pre-prints in each field using a Gaussian Process with a quasi-periodic kernel function\cite{Rasmussen:2006,Ambikasaran:2014}, allowing us to accurately predict long-term trends and periodicity in given fields of research (Figure~1).
-In nearly all research fields the number of pre-prints posted since January 2020 agree excellently with the model predictions, indicating no collective impact (positive or negative) on academic productivity due to the COVID-19 pandemic.
+We modelled the number of monthly pre-prints in each field using a Gaussian Process with a quasi-periodic kernel function\cite{Rasmussen:2006,Ambikasaran:2014}, allowing us to accurately predict long-term trends and seasonal variations (Figure~1).
+In nearly all research fields the number of pre-prints posted since January 2020 agree excellently with the model predictions (Table~1), indicating no collective impact (positive or negative) on academic productivity due to the COVID-19 pandemic.
 
-\noindent\todo{Include a table with each field where we quantify the change?}
-% field     & 2020 pre-prints & YoY change from 2019 to 2020 & Predicted pre-prints in 2020 & Number of sigma deviations away from model
-%astro-ph    & 14,893    &  +3.4 & 14,948 \pm 655    & -0.1 
-%cond-mat    & 16,188    &  +7.0 & 15,098 \pm 544    & +2.0 
-%cs          & 54,808    & +24.3 & 50,567 \pm 3153   & +1.3 
-%gr-qc       & 3,405     & +12.4 & 3,178 \pm 202     & +1.1 
-%hep-ex      & 615       & -19.8 & 797 \pm 131       & -1.4 
-%hep-lat     & 394       & -19.3 & 474 \pm 185       & -0.4 
-%hep-ph      & 4,586     &  +1.3 & 4,563 \pm 285     & +0.1 
-%hep-th      & 3,869     &  +0.9 & 3,971 \pm 262     & -0.4 
-%math        & 35,367    &  +5.5 & 35,409 \pm 1974   & -0.0 
-%math-ph     & 1,378     &  -6.3 & 1,445 \pm 131     & -0.5 
-%nlin        & 885       &  +6.1 & 820 \pm 64        & +1.0 
-%nucl-ex     & 499       &  +2.9 & 479 \pm 62        & +0.3 
-%nucl-th     & 1,212     &  -3.2 & 1,248 \pm 97      & -0.4 
-%physics     & 15,241    & +14.3 & 14,612 \pm 959    & +0.7 
-%q-bio       & 2,790     & +49.0 & 1,944 \pm 191     & +4.4 
-%q-fin       & 979       & +23.1 & 812 \pm 74        & +2.3 
-%quant-ph    & 6,341     & +13.4 & 6,210 \pm 269     & +0.5 
-%stat        & 5,180     & +20.2 & 4,719 \pm 509     & +0.9
+\begin{table}
+\begin{center}
+\small
+    \label{tab:table1}
+    \begin{tabular}{|c|l|c|c|c|c|} 
+    \hline
+    	\arxiv\ & Field of research & Pre-prints & Yearly & Predicted & Significance \\
+	code &&(2020)& change (\%)&(2020)&($\sigma$)\\
+      \hline
+\texttt{astro-ph}    & Astrophysics           & 14,893    & $ +3.4$ & 14,948 $\pm$ 655    & $-0.1$ \\
+\texttt{cond-mat}    & Condensed Matter       & 16,188    & $ +7.0$ & 15,098 $\pm$ 544    & $+2.0$ \\
+\texttt{cs      }    & Computer Science       & 54,808    & $+24.3$ & 50,567 $\pm$ 3,153  & $+1.3$ \\
+\texttt{gr-qc   }    & General Relativity and & 3,405     & $+12.4$ & 3,178 $\pm$ 202     & $+1.1$ \\
+					 & Quantum Cosmology      &&&&\\
+\texttt{hep-ex  }    & High Energy Physics    & 615       & $-19.8$ & 797 $\pm$ 131       & $-1.4$ \\
+                     & (Experiment)           &&&&\\
+\texttt{hep-lat }    & High Energy Physics    & 394       & $-19.3$ & 474 $\pm$ 185       & $-0.4$ \\
+                     & (Lattice)              &&&&\\
+\texttt{hep-ph  }    & High Energy Physics    & 4,586     & $ +1.3$ & 4,563 $\pm$ 285     & $+0.1$ \\
+                     & (Phenomenology)        &&&&\\
+\texttt{hep-th  }    & High Energy Physics    & 3,869     & $ +0.9$ & 3,971 $\pm$ 262     & $-0.4$ \\
+                     & (Theory)               &&&&\\
+\texttt{math    }    & Mathematics            & 35,367    & $ +5.5$ & 35,409 $\pm$ 1,974  & $-0.0$ \\
+\texttt{math-ph }    & Mathematical Physics   & 1,378     & $ -6.3$ & 1,445 $\pm$ 131     & $-0.5$ \\
+\texttt{nlin    }    & Nonlinear Sciences     & 885       & $ +6.1$ & 820 $\pm$ 64        & $+1.0$ \\
+\texttt{nucl-ex }    & Nuclear Experiment     & 499       & $ +2.9$ & 479 $\pm$ 62        & $+0.3$ \\
+\texttt{nucl-th }    & Nuclear Theory         & 1,212     & $ -3.2$ & 1,248 $\pm$ 97      & $-0.4$ \\
+\texttt{physics }    & Physics                & 15,241    & $+14.3$ & 14,612 $\pm$ 959    & $+0.7$ \\
+\texttt{q-bio   }    & Quantitative Biology   & 2,790     & $+49.0$ & 1,944 $\pm$ 191     & $+4.4$ \\
+\texttt{q-fin   }    & Quantitative Finance   & 979       & $+23.1$ & 812 $\pm$ 74        & $+2.3$ \\
+\texttt{quant-ph}    & Quantum Physics        & 6,341     & $+13.4$ & 6,210 $\pm$ 269     & $+0.5$ \\
+\texttt{stat    }    & Statistics             & 5,180     & $+20.2$ & 4,719 $\pm$ 509     & $+0.9$ \\
+\hline   
+    \end{tabular}
+       \caption{The number of \arxiv\ pre-prints per primary research field in 2020 compared to model predictions.}
+  \end{center}
+\end{table}
 
@@
 Discussions at conferences or collaborative visits frequently spark new research ideas that might lead to a publication many months or years later.
-% TODO: Consider moving this paragraph elsewhere to better align with the long timelines of research.
 
 Experimental research projects often require specialised laboratories, or data to be collected over many years.
-The pandemic has forced many laboratories to close or operate with restricted access, which could lead to long-lasting delays in ongoing experiments or immediate drops in publication rates in part due to difficulty in accessing completed (or nearly completed) experiment data.
-There is some evidence of this in pre-print numbers already, with {25\%} fewer pre-prints in high energy physics (\texttt{hep-ex}) in 2020 than expected by our model (observed 615; predicted 797 $\pm$ 131).
+The pandemic has forced many laboratories to close or operate with restricted access\cite{PhysicsWorld,ScienceMag}, which could lead to long-lasting delays in ongoing experiments or immediate drops in publication rates in part due to difficulty in accessing completed (or nearly completed) experiment data.
+There is some evidence of this in pre-print numbers already, with {20\%} fewer pre-prints in high energy physics (\texttt{hep-ex}) in 2020 than expected by our model (observed 615; predicted 797 $\pm$ 131).
 But declines are not ubiquitous across experimental fields.
@@
 And throughout 2020, experimental research in condensed matter (\texttt{cond-mat}) saw a 2$\sigma$ increase above our model predictions (observed 16,188; predicted 15,098 $\pm$ 544).
-Segmenting condensed matter pre-prints by research topic shows that most of this increase (in \texttt{cond-mat}) was driven by a 30\% increase in material science pre-prints (\texttt{cond-mat.mtrl-sci}), a sub-field of condensed matter that focusses on laboratory methods and techniques.
+Segmenting condensed matter pre-prints by research topic shows that most of this increase (in \texttt{cond-mat}) was driven by a 30\% increase in material science pre-prints (\texttt{cond-mat.mtrl-sci}), a sub-field of condensed matter focussed on laboratory methods and techniques.
 Despite the COVID-19 pandemic restricting access to essential laboratory equipment, this shift to publish methods and experimental design work -- which might otherwise be neglected due to ongoing laboratory work -- produced an immediate boost in condensed matter research.
@@
 The field of quantitative biology (\texttt{q-bio}) research showed the largest increase in pre-prints in 2020 above what is expected by our model.
-There were {2,790} quantitative biology pre-prints in 2020, {50\%} above the previous year, representing a $+4.4\sigma$ deviation from our model predictions (predicted 1,944 $\pm$ 191).
+There were {2,790} quantitative biology pre-prints in 2020, {50\%} above the previous year, representing a $+4.4\sigma$ deviation from our model predictions (predicted 1,944 $\pm$ 191, 846 fewer than observed).
 This increase is explainable by an increase in COVID-19 related research, as there were 844 quantitative biology pre-prints in 2020 with pandemic-related terms in their title or abstract (see Methods), and just 43 in the decade prior.
 Indeed, in {April 2020} nearly 60\% of the \arxiv\ pre-prints in quantitative biology were related to the pandemic (Figure~2).
-Pandemic-related pre-prints also appeared in other fields (computer science, physics, statistics, and quantitative finance, all peaking around {April 2020}, but pandemic-related pre-prints constituted less than 10\% of the pre-prints in these fields.
-As of mid-2021 the number of monthly pre-prints in quantitative biology has dropped to pre-pandemic levels, although approximately 20\% of these pre-prints are related to the ongoing pandemic, representing a strong shift in ongoing research focus.
+Pandemic-related pre-prints also appeared in other fields (computer science, physics, statistics, and quantitative finance, all peaking around {April 2020}), but pandemic-related pre-prints constituted less than 10\% of the total pre-prints in these fields.
+As of mid-2021 the number of monthly pre-prints in quantitative biology has dropped to pre-pandemic levels.
+Approximately 20\% of current pre-prints in quantitative biology are related to the ongoing pandemic, representing a strong shift in ongoing research focus.
 
 An increase in quantitative biology\footnote{Quantitative biology is only a sub-field of biology, and most biology pre-prints are posted to bioRxiv ({https://biorxiv.org}).} research during a global pandemic is unsurprising.
-However, the \arxiv\ data shows that this increase in pre-prints was not driven by biology researchers posting twice as many pre-prints than usual.
+However, the \arxiv\ data shows that this increase in pre-prints was not driven by existing researchers posting 50\% more pre-prints than usual.
 In 2020 there was a peak in the number of new authors appearing for the first time in the quantitative biology literature (Figure~3) while the number of new authors in all other fields remained steady.
@@
 
-%Any contributions by physicists or mathematicians to quantitative biology did not lead to significant drops in pre-prints in physics or maths.
-Throughout 2020 there were just as many pre-prints posted in physics as our model predicts, but an unusual asymmetry is present: with more pre-prints posted in the first half of 2020 than those in the later half.
-Other fields show strong seasonality in pre-prints (e.g., \texttt{hep-lat}), but physics does not.
-This seasonality in 2020 physics pre-prints is also conflated by a some physics sub-fields that saw a short burst of increased productivity early in 2020, before the effects of the pandemic were widespread (Figure~5).
-
-%drastic drop in productivity (Figure~5).
-Specifically, pre-prints in computational physics (\texttt{physics.comp-ph}) and \emph{Physics and Society} (\texttt{physics.soc-ph}) peaked to their highest level in \todo{March 2020} before declining rapidly to levels not seen since 2018.
-
-% -> timescales.
-
 The COVID-19 pandemic has had disparate impacts on research productivity in different fields.
@@
 
-%Even among the measurable changes that we have discussed, none have fully returned to normality: even while quantitative biology pre-prints have returned to pre-pandemic levels, 20\% of the research in that field is now related to the ongoing pandemic.
-
-%----------------------------------------------------------------------------------------
-%	References
-%----------------------------------------------------------------------------------------
-
-%\bibliographystylemain{myieeetr2}
-
-%\vspace{-4cm}
-%\bibliographyNew{references_new.bib}
-%{\removebibheader
- %\bibliographymain{references_new.bib}
-%}
-
-%\bibitem{Papics2017}
-%{P{\'a}pics}, P.~I., {\em et~al.}, {Signatures of internal rotation discovered
-%  in the Kepler data of five slowly pulsating B stars}, {\em \textit{Astron.
-%  Astrophys.}}, {\bfseries 598}, A74, 2017.
-
 \begin{thebibliography}{10}
@@
 
+\bibitem{Lariviere:2014}
+{Larivi{\'e}re, V., et al.,} {arXiv E-prints and the journal of record: An analysis of roles and relationships}, \emph{Journal of the Association for Information Science and Technology}, \textbf{65}, 6, doi:10.1002/asi.23044 \newblock (2014).
+
 \bibitem{Clement:2019}
@@
 Preprint at https://arxiv.org/abs/1905.00075 \newblock (2019).
-
-\bibitem{Lariviere:2014}
-%{Larivi{\'e}re, V., Sugimoto, C.~R., Macaluso, B., Milojevi{\'c}, S., Cronin, B., Thelwall, M.,} 
-{Larivi{\'e}re, V., et al.,} {arXiv E-prints and the journal of record: An analysis of roles and relationships}, \emph{Journal of the Association for Information Science and Technology}, \textbf{65}, 6, doi:10.1002/asi.23044 \newblock (2014).
 
@@
 
-%\bibitem{kippenhahn2012}
-%{Kippenhahn}, R., {Weigert}, A., and {Weiss}, A., {\em {Stellar Structure and
-%  Evolution, Astronomy and Astrophysics Library, Springer Berlin Heidelberg}}.
-%\newblock 2012.
+\bibitem{PhysicsWorld}
+physicsworld (IOP Publishing).
+2021.
+\emph{Critical research hit as COVID-19 forces physics labs to close}.
+[online] Available at {https://physicsworld.com/a/critical-research-hit-as-covid-19-forces-physics-labs-to-close/} [Accessed 4 June 2021].
+
+\bibitem{ScienceMag}
+Science Magazine (AAAS).
+2021.
+\emph{Amid pandemic, Energy Department labs close to tens of thousands of users.} [online] Available at {https://www.sciencemag.org/news/2020/03/amid-pandemic-energy-department-labs-close-tens-thousands-users} [Accessed 4 June 2021].
 
@@
 
-%\end{document}
 \newpage
-%----------------------------------------------------------------------------------------
-%	Figures
-%----------------------------------------------------------------------------------------
+
 \begin{center}
@@
 
+% Some notable peaks:
+% gr-qc 2013: https://arxiv.org/abs/1304.0670, GW
+
+\newpage
+
+\begin{center}
+	\includegraphics[width=0.95\linewidth]{new-authors-segmented-by-field-normalised}
+\end{center}
+
+\noindent \todo{\textbf{Figure 3(b):} The same as Figure 3, except normalised by number of pre-prints in that field. I wonder whether Figure 3 is useful at all!}
+% Some notable peaks:
+% gr-qc 2013: https://arxiv.org/abs/1304.0670, GW
+
 \newpage
@@
 
-\newpage
-%----------------------------------------------------------------------------------------
-%	Methods
-%----------------------------------------------------------------------------------------
+%\newpage
+
 \section*{Methods}
+
+\subsection*{Grouping by research field}
+
+Pre-prints posted to the \arxiv\ can be listed in a single field of research, or cross-listed in multiple fields.
+A field is annotated by \texttt{primary.SEC} where the primary field of research has the prefix, and the sub-field of research is represented by a short capitalised suffix.
+For example, one pre-print may have a primary field of research as stellar astrophysics (\texttt{astro-ph.SA}) and be cross-listed in machine learning (\texttt{stat.ML}).
+These field(s) of research are supplied by the corresponding author.
+It is a subjective decision whether to include more than one field of research, or what those fields of research would be.
+For this reason, throughout this work when we segment by research field we take the primary parent research field provided and ignore any cross-listed fields of research.
+When we segment pre-prints by sub-field, we similarly take the primary sub-field provided and ignore any cross-listings.
+
+A pre-print's identifier is defined by the year and month that the pre-print was posted, and the number of pre-prints already posted in that month (across all fields).
+Pre-prints posted prior to April 2007 use a different identifier scheme, and a different research field definition.
+For these reasons, we have excluded pre-prints earlier than this date.
 
@@
 Conclusions remained the same.
-Happy to use either, although I think a Gaussian process (which here is similar enough to a Poisson process) is a closer approximation to the data generating process.}
-
-\subsection*{Segmenting by research field}
-
-Pre-prints posted to the \arxiv\ can be listed in a single field of research, or cross-listed in multiple fields.
-A field is annotated by \texttt{primary.SEC} where the primary field of research has the prefix, and the sub-field of research is represented by a short capitalised suffix.
-For example, one pre-print may have a primary field of research as stellar astrophysics (\texttt{astro-ph.SA}) and be cross-listed in machine learning (\texttt{stat.ML}).
-These field(s) of research are supplied by the corresponding author.
-It is a subjective decision whether to include more than one field of research, or what those fields of research would be.
-For this reason, throughout this work when we segment by research field we take the primary parent research field provided and ignore any cross-listed fields of research.
-When we segment pre-prints by sub-field, we similarly take the primary sub-field provided and ignore any cross-listings.
+Happy to use either, although I think a Gaussian process (which here is similar enough to a Poisson process) is a closer approximation to the data generating process.
+If we use the ARIMA model then the change in q-bio research remains just as significant, but the drop in hep-lat pre-prints becomes more significant, as the GP predicts a shallower peak in pre-prints, following a gradual decline over many years, whereas the ARIMA model predicts a peak roughly equal to the previous year.}
 
@@
 We have taken a simple approach to address name confusion.
-We first define a unique author by family name and the initial of the first given name (`Family-name, I.'), such that we intentionally group together authors that may share the same initial of their second given name.
+We first define a unique author by family name and the initial of the first given name (`Family-name, I.'), such that we intentionally group together authors that may share the same initial of their first given name.
 While our approach to name confusion is grossly simple, it is unlikely that these choices have any substantial impact on our inferences.
+In general this errs on the side of name conflation and over-productivity, except for the relatively rare authors who change their last names over their career.
 Any common name is likely to appear in the literature early in the data set, and will not impact the conclusions we draw about how publishing changed in 2020.
-%While common names will appear between different fields (e.g., quantitative biology and physics), by showing this as a function of time we get an intrinsic measure of the name ambiguity between fields, and we see the increase in overlap when a spike in quantitative biology papers occurs.
-%Nevertheless, we manually inspected horiz of quantitative biology pre-prints that entirely consisted of `new authors' and cross-referenced these with Google Scholar to confirm that these authors were established researchers in other fields, and not biologists posting pre-prints for the first time.
 
@@
 
-%----------------------------------------------------------------------------------------
-%	Data availability
-%----------------------------------------------------------------------------------------
-
 \section*{Data availability}
@@
 The dataset is curated by Cornell University and hosted by Kaggle (https://www.kaggle.com/Cornell-University/arxiv).
+The dataset is updated weekly.
 We accessed this dataset on 01 June 2020.
-
-%This data set includes the identifier and metadata for all pre-prints posted to \arxiv, and is updated weekly. 
-%A pre-print's identifier is defined by the year and month that the pre-print was posted, and the number of pre-prints already posted in that month (across all fields).
-For each pre-print the metadata includes  the title, author name(s), abstract, research field(s), and the date the pre-print was first posted.
-Pre-prints posted prior to 2007 use a different identifier scheme, and for this reason we have excluded pre-prints earlier than this date.
-The data set used for analysis includes \todo{1,379,332} pre-prints posted between \todo{2007-03-30} and \todo{2020-12-30}.
-
-%All data was sourced from Kaggle (https://www.kaggle.com/Cornell-University/arxiv) on 1 June 2020.
 
@@
 
-%----------------------------------------------------------------------------------------
-%	Acknowledgements
-%----------------------------------------------------------------------------------------
 \section*{Acknowledgements}
@@
 
-%----------------------------------------------------------------------------------------
-%	Author contributions
-%----------------------------------------------------------------------------------------
-
 \section*{Author contributions}
@@
 
-%----------------------------------------------------------------------------------------
-%	Competing interests
-%----------------------------------------------------------------------------------------
 \section*{Competing interests}
@@
 
-%----------------------------------------------------------------------------------------
-%	Materials and correspondence
-%----------------------------------------------------------------------------------------
 \section*{Materials \& Correspondence}
@@
 
-%----------------------------------------------------------------------------------------
-%	Additional information
-%----------------------------------------------------------------------------------------
 \section*{Additional information}

### 2021-06-18 article/arxiv-ms.tex
incorporate comments from prasun

@@
 \usepackage{color}
-\usepackage[dvipsnames]{xcolor}
+\usepackage[table,dvipsnames]{xcolor}
 \definecolor{mygray}{gray}{0.6}
@@
 
-\usepackage[symbol]{footmisc}
-\renewcommand{\thefootnote}{\fnsymbol{footnote}}
 \usepackage{setspace}
@@
 
-\title{The impact on academic productivity owing to the COVID-19 pandemic}
+\title{The impact of the COVID-19 pandemic on academic productivity}
 
 \author{Andrew R. Casey$^{1,2}$\footnote{andrew.casey@monash.edu},
-        Ilya Mandel$^{1,2}$,
-        Prasun K. Ray$^{3}$\\
+        Ilya Mandel$^{1,2,3,4}$,
+        Prasun K. Ray$^{5}$\\
 \normalsize{$^{1}$School of Physics \& Astronomy, Monash University, Clayton 3800, Victoria, Australia}\\
 \normalsize{$^{2}$Center of Excellence for Astrophysics in Three Dimensions (ASTRO-3D), Australia}\\
-\normalsize{$^{3}$Department of Mathematics, Imperial College London, London, United Kingdom}
+\normalsize{$^{3}$OzGrav, Australian Research Council Centre of Excellence}\\
+\normalsize{for Gravitational Wave Discovery, Australia}\\
+\normalsize{$^{4}$Institute of Gravitational Wave Astronomy and School of Physics and Astronomy,}\\
+\normalsize{University of Birmingham, Birmingham, B15 2TT, United Kingdom}\\
+\normalsize{$^{5}$Department of Mathematics, Imperial College London, London, United Kingdom}
 }
@@
 \renewcommand*{\thefootnote}{\arabic{footnote}}
-
-% Responses to comments:
-% From Ilya: where fields are defined by primary arXiv category, i.e., chosen by the author(s) at submission (?)
-% --> More description given in Methods
 
@@
 With a global pandemic that has changed the world, how has it changed academic productivity? 
-Here we show that academics are posting just as many publications on the \arxiv\ pre-print server as if there were no pandemic.
-However, some immediate impacts are visible in individual research fields.
+Here we show that academics are posting just as many publications on the \arxiv\ pre-print server as if there were no pandemic: 168,630 were posted in 2020, a $\mathbf{+12.6\%}$ change from 2019 and a $\mathbf{+1.4\sigma}$ deviation above the predicted 162,577 $\mathbf{\pm}$ 4,393.
+However, some immediate negative impacts are visible in individual research fields.
 Conference cancellations have led to sharp drops in pre-prints, but laboratory closures have had mixed effects.
-While some experimental fields show mild declines in outputs, other fields have increased by writing more method-based papers that are less dependent on ongoing laboratory access.
-The most significant change is a 50\% increase in quantitative biology research, all related to the COVID-19 pandemic.
+Only some experimental fields show mild declines in outputs, with most being consistent on previous years or even increasing above model expectations.
+The most significant change is a 50\% increase ($\mathbf{+8\sigma}$) in quantitative biology research, all related to the COVID-19 pandemic.
 Some of these publications are by biologists using \arxiv\ for the first time, and some are written by researchers from other fields (e.g., physicists, mathematicians).
@@
 A pre-print on \arxiv\ does not ensure that the contents have passed peer-review, but most material on \arxiv\ eventually goes through peer-review because it is now standard in many research fields to post to \arxiv\ either during or after the peer-review process \cite{Lariviere:2014}.
-For this reason, the number of pre-prints posted to \arxiv\ approximates the number of peer-reviewed publications written at any time.
+For this reason, the number of pre-prints posted to \arxiv\ approximates the number of peer-reviewed publications written in these fields.
 
@@
 While we investigate the impact that the COVID-19 pandemic has had on different fields of research, we highlight that we are unable to identify or differentiate between authors and communities whose productivity has been significantly harmed by the pandemic, and those who were largely unscathed.
-The COVID-19 pandemic has impacted the population in unequal ways\cite{Nicola:2020,Chu:2020,IbnMohammed:2021}.
+The COVID-19 pandemic has impacted the population in unequal ways\cite{Nicola:2020,Chu:2020,Viglione:2020,Gewen:2020,IbnMohammed:2021,King:2021}.
 Generational inequality, career stage, personal circumstances, carer responsibilities, work environments, places of employment, and many other factors, all significantly contribute to the disproportionate and unequal impact the pandemic may have had on a scientist's capacity to conduct research.
@@
 
-We retrieved metadata for 1,475,914 pre-prints posted on the \arxiv\ between 1 April 2007 and 31 May 2021.
+We retrieved metadata for 1,475,914 pre-prints posted to \arxiv\ between 1 April 2007 and 31 May 2021.
 The metadata includes the creation date, research field(s), title, author name(s), abstract, and other miscellaneous information\cite{Clement:2019}.
@@
 These long-term trends are relatively predictable from year to year, allowing us to quantify any change in academic productivity due to the COVID-19 pandemic.
-We used the number of publications from January 2008 to December 2019 to predict the expected number of monthly pre-prints in 2020 in each field.
-We modelled the number of monthly pre-prints in each field using a Gaussian Process with a quasi-periodic kernel function\cite{Rasmussen:2006,Ambikasaran:2014}, allowing us to accurately predict long-term trends and seasonal variations (Figure~1).
-In nearly all research fields the number of pre-prints posted since January 2020 agree excellently with the model predictions (Table~1), indicating no collective impact (positive or negative) on academic productivity due to the COVID-19 pandemic.
+We used the number of publications from January 2015 to December 2019 to predict the expected number of monthly pre-prints in each field after January 2020.
+We modelled the number of monthly pre-prints in each field with a straight line to capture long-term (linear) trends, and used a Gaussian Process with a quasi-periodic kernel function\cite{Rasmussen:2006,Ambikasaran:2014} to capture seasonal variations (Figure~1).
+In nearly all research fields the number of pre-prints posted since January 2020 agree excellently with the model predictions (Table~1), indicating no collective impact on academic productivity due to the COVID-19 pandemic.
+The total number of \arxiv\ pre-prints in 2020 across all fields (168,630) exceeds our model predictions ($162{,}577 \pm 4{,}393$; $+1.4\sigma$).
 
@@
       \hline
-\texttt{astro-ph}    & Astrophysics           & 14,893    & $ +3.4$ & 14,948 $\pm$ 655    & $-0.1$ \\
-\texttt{cond-mat}    & Condensed Matter       & 16,188    & $ +7.0$ & 15,098 $\pm$ 544    & $+2.0$ \\
-\texttt{cs      }    & Computer Science       & 54,808    & $+24.3$ & 50,567 $\pm$ 3,153  & $+1.3$ \\
-\texttt{gr-qc   }    & General Relativity and & 3,405     & $+12.4$ & 3,178 $\pm$ 202     & $+1.1$ \\
+\texttt{astro-ph}    & Astrophysics           & 14,893    & $ +3.4$ & 14,990 $\pm$ 653      & $-0.1$ \\
+\texttt{cond-mat}    & Condensed Matter       & 16,188    & $ +7.0$ & 15,325 $\pm$ 492      & $+1.8$ \\
+\texttt{cs      }    & Computer Science       & 54,808    & $+24.3$ & 51,673 $\pm$ 4,047    & $+0.8$ \\
+\texttt{gr-qc   }    & General Relativity and & 3,405     & $+12.4$ & 3,166 $\pm$ 154       & $+1.6$ \\
 					 & Quantum Cosmology      &&&&\\
-\texttt{hep-ex  }    & High Energy Physics    & 615       & $-19.8$ & 797 $\pm$ 131       & $-1.4$ \\
+\texttt{hep-ex  }    & High Energy Physics    & 615       & $-19.8$ & 818 $\pm$ 94          & $-2.2$ \\
                      & (Experiment)           &&&&\\
-\texttt{hep-lat }    & High Energy Physics    & 394       & $-19.3$ & 474 $\pm$ 185       & $-0.4$ \\
+\texttt{hep-lat }    & High Energy Physics    & 394       & $-19.3$ & 475 $\pm$ 170         & $-0.5$ \\
                      & (Lattice)              &&&&\\
-\texttt{hep-ph  }    & High Energy Physics    & 4,586     & $ +1.3$ & 4,563 $\pm$ 285     & $+0.1$ \\
+\texttt{hep-ph  }    & High Energy Physics    & 4,586     & $ +1.3$ & 4,615 $\pm$ 329       & $-0.1$ \\
                      & (Phenomenology)        &&&&\\
-\texttt{hep-th  }    & High Energy Physics    & 3,869     & $ +0.9$ & 3,971 $\pm$ 262     & $-0.4$ \\
+\texttt{hep-th  }    & High Energy Physics    & 3,869     & $ +0.9$ & 4,000 $\pm$ 291       & $-0.4$ \\
                      & (Theory)               &&&&\\
-\texttt{math    }    & Mathematics            & 35,367    & $ +5.5$ & 35,409 $\pm$ 1,974  & $-0.0$ \\
-\texttt{math-ph }    & Mathematical Physics   & 1,378     & $ -6.3$ & 1,445 $\pm$ 131     & $-0.5$ \\
-\texttt{nlin    }    & Nonlinear Sciences     & 885       & $ +6.1$ & 820 $\pm$ 64        & $+1.0$ \\
-\texttt{nucl-ex }    & Nuclear Experiment     & 499       & $ +2.9$ & 479 $\pm$ 62        & $+0.3$ \\
-\texttt{nucl-th }    & Nuclear Theory         & 1,212     & $ -3.2$ & 1,248 $\pm$ 97      & $-0.4$ \\
-\texttt{physics }    & Physics                & 15,241    & $+14.3$ & 14,612 $\pm$ 959    & $+0.7$ \\
-\texttt{q-bio   }    & Quantitative Biology   & 2,790     & $+49.0$ & 1,944 $\pm$ 191     & $+4.4$ \\
-\texttt{q-fin   }    & Quantitative Finance   & 979       & $+23.1$ & 812 $\pm$ 74        & $+2.3$ \\
-\texttt{quant-ph}    & Quantum Physics        & 6,341     & $+13.4$ & 6,210 $\pm$ 269     & $+0.5$ \\
-\texttt{stat    }    & Statistics             & 5,180     & $+20.2$ & 4,719 $\pm$ 509     & $+0.9$ \\
+\texttt{math    }    & Mathematics            & 35,367    & $ +5.5$ & 35,362 $\pm$ 1,177    & $+0.0$ \\
+\texttt{math-ph }    & Mathematical Physics   & 1,378     & $ -6.3$ & 1,460 $\pm$ 95        & $-0.9$ \\
+\texttt{nlin    }    & Nonlinear Sciences     & 885       & $ +6.1$ & 822 $\pm$ 63          & $+1.0$ \\
+\texttt{nucl-ex }    & Nuclear Experiment     & 499       & $ +2.9$ & 490 $\pm$ 61          & $+0.1$ \\
+\texttt{nucl-th }    & Nuclear Theory         & 1,212     & $ -3.2$ & 1,234 $\pm$ 111       & $-0.2$ \\
+\texttt{physics }    & Physics                & 15,241    & $+14.3$ & 14,746 $\pm$ 506      & $+1.0$ \\
+\texttt{q-bio   }    & Quantitative Biology   & 2,790     & $+49.0$ & 1,914 $\pm$ 109       & $+8.0$ \\
+\texttt{q-fin   }    & Quantitative Finance   & 979       & $+23.1$ & 805 $\pm$ 55          & $+3.2$ \\
+\texttt{quant-ph}    & Quantum Physics        & 6,341     & $+13.4$ & 5,858 $\pm$ 266       & $+1.8$ \\
+\texttt{stat    }    & Statistics             & 5,180     & $+20.2$ & 4,824 $\pm$ 490       & $+0.7$ \\
+\hline
+&\cellcolor{gray!25}\textbf{Total} & \cellcolor{gray!25}$\mathbf{168{,}630}$ & \cellcolor{gray!25}$\mathbf{+12.6}$ & \cellcolor{gray!25}$\mathbf{162{,}577 \pm 4{,}393}$ \cellcolor{gray!25}& \cellcolor{gray!25}$\mathbf{+1.4}$ \\
 \hline 
     \end{tabular}
-       \caption{The number of \arxiv\ pre-prints per primary research field in 2020 compared to model predictions.}
+       \caption{The number of \arxiv\ pre-prints per primary research field in 2020 compared to 2019, and model predictions (see Methods).}
   \end{center}
@@
 Border closures and travel restrictions have forced many academic conferences to be held online-only, or to be cancelled.
+Hosting conferences online creates a mix of effects, forgoing in-person experiences while eliminating visa issues and disadvantages introduced by travel costs\cite{Guinnessy:2021}.
 The immediate impact of cancelling a conference is readily apparent in pre-print counts in lattice physics (Figure~1; \arxiv\ subject code \texttt{hep-lat}, see Methods for explanation), where most pre-prints are posted around December each year as conference proceedings from the International Symposium on Lattice Field Theory.
@@
 The pandemic has forced many laboratories to close or operate with restricted access\cite{PhysicsWorld,ScienceMag}, which could lead to long-lasting delays in ongoing experiments or immediate drops in publication rates in part due to difficulty in accessing completed (or nearly completed) experiment data.
-There is some evidence of this in pre-print numbers already, with {20\%} fewer pre-prints in high energy physics (\texttt{hep-ex}) in 2020 than expected by our model (observed 615; predicted 797 $\pm$ 131).
+There is some evidence of this in pre-print numbers already, with {20\%} fewer pre-prints in high energy physics (\texttt{hep-ex}) in 2020 than expected by our model (observed 615; predicted 818 $\pm$ 94).
 But declines are not ubiquitous across experimental fields.
-The closest comparable field of phenomenology in high energy physics (\texttt{hep-ph}) showed no decline.
-And throughout 2020, experimental research in condensed matter (\texttt{cond-mat}) saw a 2$\sigma$ increase above our model predictions (observed 16,188; predicted 15,098 $\pm$ 544).
-Segmenting condensed matter pre-prints by research topic shows that most of this increase (in \texttt{cond-mat}) was driven by a 30\% increase in material science pre-prints (\texttt{cond-mat.mtrl-sci}), a sub-field of condensed matter focussed on laboratory methods and techniques.
-Despite the COVID-19 pandemic restricting access to essential laboratory equipment, this shift to publish methods and experimental design work -- which might otherwise be neglected due to ongoing laboratory work -- produced an immediate boost in condensed matter research.
+The closest comparable field of phenomenology in high energy physics (\texttt{hep-ph}) showed no substantial decline.
+Neither did nuclear experimental research (\texttt{nucl-ex}).
+And throughout 2020, experimental research in condensed matter (\texttt{cond-mat}) saw a $1.8\sigma$ increase above our model predictions (observed 16,188; predicted 15,325 $\pm$ 492).
+Segmenting condensed matter pre-prints by research topic shows that most of this increase (in \texttt{cond-mat}) was driven by a 30\% increase in material science pre-prints (\texttt{cond-mat.mtrl-sci}).
+%focussed on laboratory methods and techniques.
+
+%Despite the COVID-19 pandemic restricting access to essential laboratory equipment, 
+%we speculate that 
+%this shift to publish methods and experimental design work -- which might otherwise be neglected due to ongoing laboratory work -- produced an immediate boost in condensed matter research.
 This is likely an effect with a limited lifespan: eventually, condensed matter (and other experimental) research will require laboratory access to continue publishing.
 
+%We speculate that this increase relies on experimental data taken prior to 2020, since the COVID-19 pandemic has restricted access to laboratory equipment.
+
 The field of quantitative biology (\texttt{q-bio}) research showed the largest increase in pre-prints in 2020 above what is expected by our model.
-There were {2,790} quantitative biology pre-prints in 2020, {50\%} above the previous year, representing a $+4.4\sigma$ deviation from our model predictions (predicted 1,944 $\pm$ 191, 846 fewer than observed).
+There were {2,790} quantitative biology pre-prints in 2020, {50\%} above the previous year, representing a $+8\sigma$ deviation from our model predictions (predicted 1,914 $\pm$ 109, 876 fewer than observed).
 This increase is explainable by an increase in COVID-19 related research, as there were 844 quantitative biology pre-prints in 2020 with pandemic-related terms in their title or abstract (see Methods), and just 43 in the decade prior.
 Indeed, in {April 2020} nearly 60\% of the \arxiv\ pre-prints in quantitative biology were related to the pandemic (Figure~2).
-Pandemic-related pre-prints also appeared in other fields (computer science, physics, statistics, and quantitative finance, all peaking around {April 2020}), but pandemic-related pre-prints constituted less than 10\% of the total pre-prints in these fields.
+Pandemic-related pre-prints also appeared in other fields (computer science, physics, statistics, and quantitative finance, all peaking around {April 2020}), but pandemic-related pre-prints constituted fewer than 10\% of the total pre-prints in these fields.
 As of mid-2021 the number of monthly pre-prints in quantitative biology has dropped to pre-pandemic levels.
@@
 
+\renewcommand{\thefootnote}{$\dagger$}
+
 An increase in quantitative biology\footnote{Quantitative biology is only a sub-field of biology, and most biology pre-prints are posted to bioRxiv ({https://biorxiv.org}).} research during a global pandemic is unsurprising.
-However, the \arxiv\ data shows that this increase in pre-prints was not driven by existing researchers posting 50\% more pre-prints than usual.
+However, the question arises if this increase was driven by existing researchers in the field, or by authors from other fields entering \texttt{q-bio} due to the pandemic.
+%However, the \arxiv\ data shows that this increase in pre-prints was not driven by existing researchers posting 50\% more pre-prints than usual. 
 In 2020 there was a peak in the number of new authors appearing for the first time in the quantitative biology literature (Figure~3) while the number of new authors in all other fields remained steady.
-The sudden influx of new authors cannot be explained by large ($>$10) newly-formed collaborations working together to tackle the impending pandemic (Figure~4).
+%The sudden influx of new authors cannot be explained by large ($>$10) newly-formed collaborations working together to tackle the impending pandemic (Figure~4).
 The increase in pre-prints (and new authors) is driven by small (1-4) groups of authors, where many had never posted pre-prints to quantitative biology before.
@@
 A close examination of these pre-prints reveals that many are established biologists who have not used the \arxiv\ before, and are now doing so presumably to help ensure that their COVID-19 research is more widely available.
-This represents a genuine increase in new researchers to the \arxiv, and a boost to making quantitative biology research more widely accessible.
+This represents a genuine increase in new researchers using the \arxiv, and a boost to making quantitative biology research more widely accessible.
 Of those new authors in quantitative biology who also appear in other \arxiv\ fields, some of this will be due to name confusion (see Methods): where two researchers in different fields share the same publishing name.
@@
 
-The COVID-19 pandemic has had disparate impacts on research productivity in different fields.
-Some fields show a drop in research outputs.
+%The COVID-19 pandemic has had disparate impacts on research productivity in different fields.
+
+%\todo{Some fields show a drop in research outputs.
 Few show an increase, either driven by pandemic-related research or by shifts away from experimental research to topics that do not require ongoing laboratory access.
-In this study we have only focussed on quantity of publications as a measure for productivity, and not their quality.
-While citations are the most commonly used measure of impact of academic publications, that metric becomes a more biased statistic when many related pre-prints are all being posted nearly at the same time.
+}
+
+The COVID-19 pandemic appears to have had minimal impact on collective research productivity to date. 
+Amongst individual fields the largest decline is explainable by a cancelled conference, and the few other drops in pre-prints are statistically insignificant. 
+In this study we have only focussed on the quantity of publications as a measure for productivity, and not their quality.
+While citations are the most commonly used measure of impact of academic publications, that metric becomes a more biased statistic when many related pre-prints are all being posted nearly at the same time\cite{Fassin:2021}.
 The relatively long timescales of academic research would suggest that the full impact on academic productivity due to the COVID-19 pandemic is yet to be seen.
-For those fields without a measurable drop in research outputs yet, it's plausible that many of the pre-prints being posted are research projects that were well underway before the pandemic, with the full impact of academic research yet to be seen.
+For those fields without a measurable drop in research outputs, it's plausible that many of the pre-prints being posted are research projects that were well underway before the pandemic, with the full impact of academic research yet to be seen.
 
@@
 
-\bibitem{Clement:2019}
-{Clement, C.~B., Bierbaum, M., O'Keeffe, K.~P., Alemi, A.~A.,}
-{On the Use of ArXiv as a Dataset},
-Preprint at https://arxiv.org/abs/1905.00075 \newblock (2019).
-
-\bibitem{IbnMohammed:2021}
-{Ibn-Mohammed, T., et al.,} {A critical analysis of the impacts of {COVID}-19 on the global economy and ecosystems and opportunities for circular economy strategies}, \emph{Resources, Conservation and Recycling}, \textbf{164}, 105169, doi:1016/j.resconrec.2020.105169 \newblock (2021).
-
 \bibitem{Nicola:2020}
@@
 {Social consequences of mass quarantine during epidemics: a systematic review with implications for the {COVID}-19 response}, \emph{Journal of Travel Medicine}, \textbf{27}, 7, doi:10.1093/jtm/taaa192 \newblock (2020).
+
+\bibitem{Viglione:2020}
+{Viglione, G.}, {Are women publishing less during the pandemic? Here?s what the data say}. \emph{Nature}, \textbf{581}, 356-366, doi:10.1038/d41586-020-01294-9 \newblock (2020).
+
+\bibitem{Gewen:2020},
+{Gewen, V.}, {The career cost of COVID-19 to female researchers, and how science should respond}. \emph{Nature}, \textbf{583}, 867--869, doi:10.1038/d41586-020-02183-x \newblock (2020).
+
+\bibitem{IbnMohammed:2021}
+{Ibn-Mohammed, T., et al.,} {A critical analysis of the impacts of {COVID}-19 on the global economy and ecosystems and opportunities for circular economy strategies}, \emph{Resources, Conservation and Recycling}, \textbf{164}, 105169, doi:1016/j.resconrec.2020.105169 \newblock (2021).
+
+\bibitem{King:2021}
+{King, M.~M., Frederickson M.~E.}, {The Pandemic Penalty: The Gendered Effects of COVID-19 on Scientific Productivity}. \emph{Socius}. January 2021. doi:10.1177/23780231211006977 \newblock (2021).
+
+\bibitem{Clement:2019}
+{Clement, C.~B., Bierbaum, M., O'Keeffe, K.~P., Alemi, A.~A.,}
+{On the Use of ArXiv as a Dataset},
+Preprint at https://arxiv.org/abs/1905.00075 \newblock (2019).
 
@@
 {Fast Direct Methods for Gaussian Processes}, Preprint at http://arxiv.org/abs/1403.6015 \newblock (2014).
+
+\bibitem{Guinnessy:2021}
+Guinnessy, P.
+(Physics Today).
+2021.
+\emph{Physicists gathered virtually for APS March Meeting}.
+[online] Available at {https://physicstoday.scitation.org/do/10.1063/PT.6.3.20210405a/full} [Accessed 17 June 2021].
 
@@
 
+\bibitem{Fassin:2021}
+{Fassin, Y.} {Research on Covid-19: a disruptive phenomenon for bibliometrics}, \emph{Scientometrics}, \textbf{126}, 5305--5319, doi:10.1007/s11192-021-03989-w \newblock (2021).
+
 \end{thebibliography}
@@
 
-\noindent \textbf{Figure 1:} \textbf{Since the COVID-19 pandemic began there has been no significant change in monthly \arxiv\ pre-prints in most fields (black).} The expected number of pre-prints from our model is shown in red, and the red region shows the uncertainty in that expectation.
-Each panel shows a primary field of research in \arxiv.
-\todo{Blue is ARIMA model.}
+\noindent \textbf{Figure 1:} \textbf{There has been no significant change in monthly \arxiv\ pre-prints (black) since the COVID-19 pandemic began.} The expected number of pre-prints from our Gaussian Process model is shown in red, and the red region shows the uncertainty in that expectation.
+The dashed light grey line shows predictions from an autoregressive integrated moving average model.
 
@@
 
-\noindent \textbf{Figure 2: Pre-prints related to the COVID-19 pandemic peaked in quantitative biology in 2020, accounting for nearly 60\% of quantitative biology pre-prints.} Shown are five other research fields with the highest number of pandemic-related pre-prints in 2020.
+\noindent \textbf{Figure 2: Pre-prints related to the COVID-19 pandemic peaked in April 2020, accounting for nearly 60\% the quantitative biology (\texttt{q-bio}) pre-prints in that month.} Shown are five other research fields with the highest number of pandemic-related pre-prints in 2020.
 Monthly pre-prints are shown on the left, and the percent of pandemic-related pre-prints in that research field is shown on the right.
@@
 \begin{center}
-	\includegraphics[width=0.95\linewidth]{new-authors-segmented-by-field}
+	\includegraphics[width=0.95\linewidth]{new-authors-segmented-by-field-combined}
 \end{center}
 
-\noindent \textbf{Figure 3:} \textbf{The number of new authors appearing in \arxiv\ pre-prints per month peaked in quantitative biology in 2020.} Fields well-established before 2007 (e.g., \texttt{astro-ph}) show an apparent influx of `new authors' at the time the data starts: these authors are already established academics.
-The slow change in new authors after 2012 approximates the net number of new authors joining the field.
+\noindent \textbf{Figure 3:} \textbf{The number of new authors appearing in \arxiv\ pre-prints per month (black) peaked in quantitative biology in 2020.} Here we show new authors since 2010, but we used data since 2007 to determine if an author was new.
+The slow change in new authors  approximates the net number of new authors joining the field.
+The number of new authors appearing per number of pre-prints is shown in light blue, with identical limits on the right-hand y-axis.
+This highlights authorship differences between fields (e.g., \texttt{hep-ex} and \texttt{hep-lat}) for large experiments.
 
@@
 
-\newpage
+%\newpage
 
-\begin{center}
-	\includegraphics[width=0.95\linewidth]{new-authors-segmented-by-field-normalised}
-\end{center}
+%\begin{center}
+%     \includegraphics[width=\linewidth]{q-bio-pre-prints-segmented-by-author-count}
+%\end{center}
 
-\noindent \todo{\textbf{Figure 3(b):} The same as Figure 3, except normalised by number of pre-prints in that field. I wonder whether Figure 3 is useful at all!}
-% Some notable peaks:
-% gr-qc 2013: https://arxiv.org/abs/1304.0670, GW
-
-\newpage
-
-\begin{center}
-     \includegraphics[width=\linewidth]{q-bio-pre-prints-segmented-by-author-count}
-\end{center}
-
-\noindent \textbf{Figure 4:} \textbf{The increase in quantitative biology pre-prints in 2020 cannot be attributed to large collaborations.} The coloured regions represent the different author group sizes for quantitative biology pre-prints per month with time.
+%\noindent \textbf{Figure 4:} \textbf{The increase in quantitative biology (\texttt{q-bio}) pre-prints in 2020 cannot be attributed to large collaborations.} The coloured regions represent the different author group sizes for quantitative biology pre-prints per month with time.
 
@@
 Pre-prints posted to the \arxiv\ can be listed in a single field of research, or cross-listed in multiple fields.
-A field is annotated by \texttt{primary.SEC} where the primary field of research has the prefix, and the sub-field of research is represented by a short capitalised suffix.
+A field is annotated by \texttt{primary.SEC} where the primary field of research has the prefix, and the sub-field of research is represented by a short suffix.
 For example, one pre-print may have a primary field of research as stellar astrophysics (\texttt{astro-ph.SA}) and be cross-listed in machine learning (\texttt{stat.ML}).
@@
 
-We use monthly pre-print counts per primary research field from January 2007 until December 2019 to make predictions for the number of pre-prints expected per month from January 2020 onwards.
-We use a Gaussian process with a quasi-periodic kernel covariance function to model long-term trends and seasonal periodicity\cite{Rasmussen:2006,Ambikasaran:2014}.
-We fix the periodicity hyper-parameter to one year, and optimise all other hyper-parameters.
+We use monthly pre-print counts per primary research field from January 2015 until December 2019 to make predictions for the number of pre-prints expected per month from January 2020 onwards.
+We use a straight line to model the mean pre-prints over time, and a Gaussian Process with a quasi-periodic kernel covariance function to capture seasonal periodicity\cite{Rasmussen:2006,Ambikasaran:2014}, and a white noise component.
+We fit the parameters of the mean model and the kernel hyper-parameters simultaneously, but kept the periodicity hyper-parameter fixed to one year.
 The predictions of the expectation and variance in the number of monthly pre-print counts are conditioned on the optimized hyper-parameters.
 We fit this model to monthly pre-print counts for every primary research field.
+These predictions are listed in Table~1.
+The expected total number of pre-prints, and the uncertainty in that number, are summed from the models used in individual research fields.
 
-\todo{We also considered an ARIMA model.
-The results were very similar.
-The largest difference between predictions is in \texttt{quant-ph}.
-Conclusions remained the same.
-Happy to use either, although I think a Gaussian process (which here is similar enough to a Poisson process) is a closer approximation to the data generating process.
-If we use the ARIMA model then the change in q-bio research remains just as significant, but the drop in hep-lat pre-prints becomes more significant, as the GP predicts a shallower peak in pre-prints, following a gradual decline over many years, whereas the ARIMA model predicts a peak roughly equal to the previous year.}
+Our conclusions are reasonably insensitive to the choice of model.
+As an alternative model we considered an autoregressive integrated moving average (ARIMA) model\cite{BoxJenkins:1970} with seasonal terms ($p=d=q=1$ for the non-seasonal component, and $P=D=Q=1$ with $m=12$ for the seasonal component).
+The predictions from the ARIMA model agree excellently with the mean predictions from the Gaussian Process model.
+The largest differences between these models is in \texttt{hep-lat}, where the ARIMA model predicts a higher peak number of pre-prints than the Gaussian Process model.
+This is not unexpected: the peak height of lattice physics pre-prints has been dropping for the last three years, and this gradual decline is better captured by the Gaussian Process model.
+If we adopted the ARIMA model then the drop in \texttt{hep-lat} pre-prints would become more significant.
+Throughout this work we adopt the predictions from the Gaussian Process model, unless otherwise specified.
 
@@
 
-The \arxiv\ metadata available to us does not include institutional affiliations, or identifiers that would uniquely identify an author.
-For these reasons, we have taken steps to minimise the effects of name confusion.
+The \arxiv\ metadata available to us do not include institutional affiliations, or identifiers that would uniquely identify an author. 
+%For these reasons, we have taken steps to minimise the effects of name confusion. 
 There are two primary ways that name confusion could impact our inferences.
@@
 
-We identified a pre-print as being related to the COVID-19 pandemic if the title or abstract contained any of the following four (case-insensitive) terms: `pandemic', `COVID', `SARS-CoV-2', or `lockdown'.
+We identified a pre-print as being related to the COVID-19 pandemic if the title or abstract contained any of the following four (case-insensitive) terms: `pandemic', `COVID', `SARS-CoV-2', `lockdown', or `coronavirus'.
 Before 2020, these terms rarely appear in the title or abstract of any pre-print on \arxiv\ (Figure~2).
@@
 The dataset is updated weekly.
-We accessed this dataset on 01 June 2020.
+We accessed this dataset on 01 June 2021.
 
 \section*{Code availability}
-The software developed to retrieve and analyse the data for this projectis available online (https://github.com/andycasey/arxiv-covid).
+The software developed to retrieve and analyse the data for this project is available online (https://github.com/andycasey/arxiv-covid).
 
@@
 
+% \let\oldthebibliography=\thebibliography
+% \renewenvironment{thebibliography}[1]{%
+%    \oldthebibliography{#1}%
+%    \setcounter{enumiv}{36}%
+% }
+
+\begin{thebibliography}{10}
+
+\setcounter{enumiv}{16}
+
+\bibitem{BoxJenkins:1970}
+{Box, G.~E.~P., Jenkins, G. M.} Time series analysis: Forecasting and control. San Francisco: Holden-Day. \newblock (1970).
+
+\end{thebibliography}
+
 \section*{Acknowledgements}
@@
 for comments on publication trends in high energy physics.
+% Funding acknowledgements.
+A.~R.~C. is supported in part by the Australian Research Council through a Discovery Early Career Researcher Award (DE190100656). 
+Parts of this research were supported by the Australian Research Council Centre of Excellence for All Sky Astrophysics in 3 Dimensions (ASTRO 3D), through project number CE170100013.
+I.~M. is a recipient of the Australian Research Council Future Fellowship FT190100574.
 
@@
 
-\textbf{Supplementary Information} is available for this paper.
+%\textbf{Supplementary Information} is available for this paper.
 

### 2021-06-22 article/arxiv-ms.tex
formatting etc

@@
 \usepackage{lineno}
-\linenumbers
+%\linenumbers
 
@@
 \normalsize{$^{1}$School of Physics \& Astronomy, Monash University, Clayton 3800, Victoria, Australia}\\
-\normalsize{$^{2}$Center of Excellence for Astrophysics in Three Dimensions (ASTRO-3D), Australia}\\
+\normalsize{$^{2}$ARC Centre of Excellence for Astrophysics in Three Dimensions (ASTRO-3D), Australia}\\
 \normalsize{$^{3}$OzGrav, Australian Research Council Centre of Excellence}\\
@@
 \renewcommand*{\thefootnote}{\arabic{footnote}}
-
-% Outstanding issues to address/close:
-% 1.
-From Ilya: Since ADS indexes arxiv (at least for astro), is it worth checking the fraction of papers from, say, 3 years ago that have only arXiv numbers (never published)?
-As well as those which are published in peer-review journals with no arXiv numbers, for completeness in the other direction?
-
-% 2. From Ilya: "All authors new to arXiv vs. at least one author previously on arXiv but all new to q-bio vs. at least one author previously on q-bio vs. all previously on q-bio."
 
@@
 Here we show that academics are posting just as many publications on the \arxiv\ pre-print server as if there were no pandemic: 168,630 were posted in 2020, a $\mathbf{+12.6\%}$ change from 2019 and a $\mathbf{+1.4\sigma}$ deviation above the predicted 162,577 $\mathbf{\pm}$ 4,393.
-However, some immediate negative impacts are visible in individual research fields.
+However, some immediate impacts are visible in individual research fields.
 Conference cancellations have led to sharp drops in pre-prints, but laboratory closures have had mixed effects.
@@
 Generational inequality, career stage, personal circumstances, carer responsibilities, work environments, places of employment, and many other factors, all significantly contribute to the disproportionate and unequal impact the pandemic may have had on a scientist's capacity to conduct research.
-%These factors are more important determinants to research capacity than a scientist's field of research. 
-While this limited analysis does not address these issues, it is important to consider that if a community as a whole has not yet demonstrated a significant change in productivity, many researchers have faced significant challenges and suffered physically, mentally, emotionally, and professionally.
+While this limited analysis does not address these issues, it is important to consider that if a community has not yet demonstrated a significant change in productivity, many researchers have faced significant challenges and suffered physically, mentally, emotionally, and professionally.
 
@@
 Border closures and travel restrictions have forced many academic conferences to be held online-only, or to be cancelled.
-Hosting conferences online creates a mix of effects, forgoing in-person experiences while eliminating visa issues and disadvantages introduced by travel costs\cite{Guinnessy:2021}.
+Hosting conferences online creates a mix of effects, as forgoing in-person experiences benefits those who could not otherwise participate due to visa issues or travel costs\cite{Guinnessy:2021}.
 The immediate impact of cancelling a conference is readily apparent in pre-print counts in lattice physics (Figure~1; \arxiv\ subject code \texttt{hep-lat}, see Methods for explanation), where most pre-prints are posted around December each year as conference proceedings from the International Symposium on Lattice Field Theory.
@@
 Segmenting condensed matter pre-prints by research topic shows that most of this increase (in \texttt{cond-mat}) was driven by a 30\% increase in material science pre-prints (\texttt{cond-mat.mtrl-sci}).
-%focussed on laboratory methods and techniques.
-
-%Despite the COVID-19 pandemic restricting access to essential laboratory equipment, 
-%we speculate that 
-%this shift to publish methods and experimental design work -- which might otherwise be neglected due to ongoing laboratory work -- produced an immediate boost in condensed matter research.
-This is likely an effect with a limited lifespan: eventually, condensed matter (and other experimental) research will require laboratory access to continue publishing.
-
-%We speculate that this increase relies on experimental data taken prior to 2020, since the COVID-19 pandemic has restricted access to laboratory equipment.
 
@@
 There were {2,790} quantitative biology pre-prints in 2020, {50\%} above the previous year, representing a $+8\sigma$ deviation from our model predictions (predicted 1,914 $\pm$ 109, 876 fewer than observed).
-This increase is explainable by an increase in COVID-19 related research, as there were 844 quantitative biology pre-prints in 2020 with pandemic-related terms in their title or abstract (see Methods), and just 43 in the decade prior.
+This increase is explainable by an increase in COVID-19 related research, as there were 864 quantitative biology pre-prints in 2020 with pandemic-related terms in their title or abstract (see Methods), and just 39 in the decade prior.
 Indeed, in {April 2020} nearly 60\% of the \arxiv\ pre-prints in quantitative biology were related to the pandemic (Figure~2).
@@
 An increase in quantitative biology\footnote{Quantitative biology is only a sub-field of biology, and most biology pre-prints are posted to bioRxiv ({https://biorxiv.org}).} research during a global pandemic is unsurprising.
-However, the question arises if this increase was driven by existing researchers in the field, or by authors from other fields entering \texttt{q-bio} due to the pandemic.
-%However, the \arxiv\ data shows that this increase in pre-prints was not driven by existing researchers posting 50\% more pre-prints than usual. 
+However, the question arises if this increase was driven by existing researchers in the field, or by authors from other fields entering quantitative biology due to the pandemic.
 In 2020 there was a peak in the number of new authors appearing for the first time in the quantitative biology literature (Figure~3) while the number of new authors in all other fields remained steady.
-%The sudden influx of new authors cannot be explained by large ($>$10) newly-formed collaborations working together to tackle the impending pandemic (Figure~4).
 The increase in pre-prints (and new authors) is driven by small (1-4) groups of authors, where many had never posted pre-prints to quantitative biology before.
 
-Some of these new authors in quantitative biology have never before appeared in any other \arxiv\ fields, either.
-A close examination of these pre-prints reveals that many are established biologists who have not used the \arxiv\ before, and are now doing so presumably to help ensure that their COVID-19 research is more widely available.
+Some of these new authors in quantitative biology have never appeared in any other \arxiv\ fields before, either.
+A close examination of these pre-prints reveals that many are established biologists who have not used the \arxiv\ before and are now doing so presumably to help ensure that their COVID-19 research is more widely available.
 This represents a genuine increase in new researchers using the \arxiv, and a boost to making quantitative biology research more widely accessible.
 Of those new authors in quantitative biology who also appear in other \arxiv\ fields, some of this will be due to name confusion (see Methods): where two researchers in different fields share the same publishing name.
-However, a careful examination of pandemic-related pre-prints in quantitative biology that were posted in 2020 shows that many were indeed written by researchers from other fields (primarily physicists and mathematicians), who had never before posted about quantitative biology.
+However, a careful examination of pandemic-related pre-prints in quantitative biology posted in 2020 shows that many were indeed written by researchers from other fields (primarily physicists and mathematicians), who had never posted about quantitative biology.
 These pre-prints tend to focus on modelling the COVID-19 outbreak using public data sets, rather than quantitative biology research that requires more expert domain knowledge.
 
-%The COVID-19 pandemic has had disparate impacts on research productivity in different fields.
-
-%\todo{Some fields show a drop in research outputs.
-Few show an increase, either driven by pandemic-related research or by shifts away from experimental research to topics that do not require ongoing laboratory access.
-}
-
 The COVID-19 pandemic appears to have had minimal impact on collective research productivity to date.
+In this study we have only focussed on the quantity of publications as a measure for productivity, and not their quality.
+While citations are the most used measure of impact of academic publications, that metric becomes a more biased statistic when many related pre-prints are all being posted nearly at the same time\cite{Fassin:2021}.
 Amongst individual fields the largest decline is explainable by a cancelled conference, and the few other drops in pre-prints are statistically insignificant.
-In this study we have only focussed on the quantity of publications as a measure for productivity, and not their quality.
-While citations are the most commonly used measure of impact of academic publications, that metric becomes a more biased statistic when many related pre-prints are all being posted nearly at the same time\cite{Fassin:2021}.
+It's plausible that many of the pre-prints already posted are research projects that were well underway before the pandemic.
 The relatively long timescales of academic research would suggest that the full impact on academic productivity due to the COVID-19 pandemic is yet to be seen.
-For those fields without a measurable drop in research outputs, it's plausible that many of the pre-prints being posted are research projects that were well underway before the pandemic, with the full impact of academic research yet to be seen.
+The approach applied here will be a useful quantitative approach to return to in subsequent years to explore the longer-term impact of the pandemic.
 
@@
 \begin{center}
-\includegraphics[width=0.95\linewidth]{pre-prints-segmented-by-field}
+\includegraphics[width=0.95\linewidth]{Casey_Figure_1} %pre-prints-segmented-by-field}
 \end{center}
@@
 \begin{center}
- \includegraphics[width=\linewidth]{pandemic-related-preprints.pdf}
+ \includegraphics[width=\linewidth]{Casey_Figure_2} %pandemic-related-preprints.pdf}
 \end{center}
@@
 \begin{center}
-	\includegraphics[width=0.95\linewidth]{new-authors-segmented-by-field-combined}
+	\includegraphics[width=0.95\linewidth]{Casey_Figure_3} %new-authors-segmented-by-field-combined
 \end{center}
@@
 
-%\newpage
-
-%\begin{center}
-%     \includegraphics[width=\linewidth]{q-bio-pre-prints-segmented-by-author-count}
-%\end{center}
-
-%\noindent \textbf{Figure 4:} \textbf{The increase in quantitative biology (\texttt{q-bio}) pre-prints in 2020 cannot be attributed to large collaborations.} The coloured regions represent the different author group sizes for quantitative biology pre-prints per month with time.
-
 \newpage
-
-%\begin{center}
-%     \includegraphics[width=\linewidth]{winners-and-losers-of-covid}
-%\end{center}
-
-%\noindent \textbf{Figure 5:} \textbf{Mat-sci}
-
-%\newpage
 
@@
 
+The \arxiv\ team has parsed given and family names from author strings for all pre-prints.
+In general the author names have been appropriately parsed, even given a wide set of input formats.
+We made only one correction to these parsed names.
+When the author name was provided in the format of `Family-name First-given-name I.', without any commas, this was interpreted by the \arxiv\ parser such that the family name was `I.' and the given names were `Family-name First-given-name'.
+We identified situations like this by checking if the parsed family name was just an initial, and if two given names were parsed.
+In these situations, we correctly ordered the author name.
+
 The \arxiv\ metadata available to us do not include institutional affiliations, or identifiers that would uniquely identify an author. 
-%For these reasons, we have taken steps to minimise the effects of name confusion. 
 There are two primary ways that name confusion could impact our inferences.
@@
 While our approach to name confusion is grossly simple, it is unlikely that these choices have any substantial impact on our inferences.
-In general this errs on the side of name conflation and over-productivity, except for the relatively rare authors who change their last names over their career.
-Any common name is likely to appear in the literature early in the data set, and will not impact the conclusions we draw about how publishing changed in 2020.
+In general, this errs on the side of name conflation and over-productivity, except for the relatively rare authors who change their last names over their career.
+Any common name is likely to appear in the literature early in the data set and will not impact the conclusions we draw about how publishing changed in 2020.
+Indeed, we re-produced the analysis and figures in this paper by taking all initials given and it made no change to our conclusions.
+This is due, in part, because most authors do not provide middle initials.
 

### 2021-09-14 article/arxiv-ms.tex
old upload eeek

@@
 \noindent \textbf{`\emph{Publish or perish}' is an expression describing the pressure on academics to consistently publish research to ensure a successful career in academia. 
-With a global pandemic that has changed the world, how has it changed academic productivity? 
+A global pandemic has changed the world, but how has it changed academic productivity?
 Here we show that academics are posting just as many publications on the \arxiv\ pre-print server as if there were no pandemic: 168,630 were posted in 2020, a $\mathbf{+12.6\%}$ change from 2019 and a $\mathbf{+1.4\sigma}$ deviation above the predicted 162,577 $\mathbf{\pm}$ 4,393.
@@
 Neither did nuclear experimental research (\texttt{nucl-ex}).
-And throughout 2020, experimental research in condensed matter (\texttt{cond-mat}) saw a $1.8\sigma$ increase above our model predictions (observed 16,188; predicted 15,325 $\pm$ 492).
+And throughout 2020, research in condensed matter (\texttt{cond-mat}) saw a $1.8\sigma$ increase above our model predictions (observed 16,188; predicted 15,325 $\pm$ 492).
 Segmenting condensed matter pre-prints by research topic shows that most of this increase (in \texttt{cond-mat}) was driven by a 30\% increase in material science pre-prints (\texttt{cond-mat.mtrl-sci}).
@@
 
-Some of these new authors in quantitative biology have never appeared in any other \arxiv\ fields before, either.
-A close examination of these pre-prints reveals that many are established biologists who have not used the \arxiv\ before and are now doing so presumably to help ensure that their COVID-19 research is more widely available.
+One quarter of the excess pre-prints posted to quantitative biology during 2020 were related to the pandemic and all authors had never appeared in the quantitative biology literature (234 of the 876 excess pre-prints in \texttt{q-bio}).
+A careful examination reveals that about half of these (45\%; 107/234) were written by mathematicians or physicists, with another 26\% (61/234) led by engineers, computer scientists, or economists.
+These pre-prints tend to focus on modelling the COVID-19 outbreak using public data sets, rather than quantitative biology research that requires more expert domain knowledge.
+About 3\% (8/234) were led by established biologists who have not used the \arxiv\ before and are now doing so presumably to help ensure that their COVID-19 research is more widely available.
+This fraction is likely higher among pre-prints where at least one author has appeared before in the quantitative biology literature.
 This represents a genuine increase in new researchers using the \arxiv, and a boost to making quantitative biology research more widely accessible.
-Of those new authors in quantitative biology who also appear in other \arxiv\ fields, some of this will be due to name confusion (see Methods): where two researchers in different fields share the same publishing name.
-However, a careful examination of pandemic-related pre-prints in quantitative biology posted in 2020 shows that many were indeed written by researchers from other fields (primarily physicists and mathematicians), who had never posted about quantitative biology.
-These pre-prints tend to focus on modelling the COVID-19 outbreak using public data sets, rather than quantitative biology research that requires more expert domain knowledge.
+
+% What to say of the remaining 75%?
+
+%However, a careful examination of pandemic-related pre-prints in quantitative biology posted in 2020 shows that many were indeed written by researchers from other fields (primarily physicists and mathematicians), who had never posted about quantitative biology.
+
+%Some of these new authors in quantitative biology have never appeared in any other \arxiv\ fields before, either.
+A close examination of 234 pre-prints posted during 2020 to \texttt{q-bio.PE} where no authors had posted to \texttt{q-bio} before shows that
+
+%A close examination of these pre-prints reveals that many are
+
+% 234 papers in 2020 in q-bio.PE written about the pandemic where no author had ever appeared before in the q-bio literature.
+% among these, 8 were led by authors who were domain experts in quantitative biology or a related field, and were 'new' to posting on arxiv
+% among these 234, 107 (45%) were led by mathematicians or physicists, with another 61/234 (26%) led by engineers, computer scientists, or economists.
+
+% [('mathematics', 58),
+% ('physics', 41), + astrophysics, 8 ==> 49
+% ('engineering', 26),
+% ('computer science', 19),
+% ('economics', 16),
+% close examination of a random sample of 50 of these "new" authors and their academic affiliations found x to be established biologists who had not used the arXiv before while the rest were researchers from other fields (primarily physics and mathematics) who had not previously posted quantitative biology papers to the arXiv.
+
+The pandemic has also created distinct challenges when interpreting trends in \arxiv\ data.
+For example, there was a large spike in papers submitted to the \texttt{stat} category in June 2020 which one might assume was driven by COVID-related work given what we have observed in \texttt{q-bio}.
+However, the number of COVID-related papers posted in this period is modest, and the key factor appears to be delayed paper submission deadlines for two large conferences (NeurIPS and the Joint Statistical Meetings).
+Both conferences have had submission deadlines in mid-May in recent years, but in 2020, their deadlines were moved to early June.
 

### 2021-09-14 article/arxiv.bib .tex
old upload eeek

@article{Ginsparg:2011,
  doi = {10.1038/476145a},
  url = {https://doi.org/10.1038/476145a},
  year = {2011},
  month = aug,
  publisher = {Springer Science and Business Media {LLC}},
  volume = {476},
  number = {7359},
  pages = {145--147},
  author = {Paul Ginsparg},
  title = {{ArXiv} at 20},
  journal = {Nature}
}

@article{Lariviere:2014,
   title={arXiv E-prints and the journal of record: An analysis of roles and relationships},
   volume={65},
   ISSN={2330-1635},
   url={http://dx.doi.org/10.1002/asi.23044},
   DOI={10.1002/asi.23044},
   number={6},
   journal={Journal of the Association for Information Science and Technology},
   publisher={Wiley},
   author={Larivière, Vincent and Sugimoto, Cassidy R. and Macaluso, Benoit and Milojevi?, Sta¨a and Cronin, Blaise and Thelwall, Mike},
   year={2014},
   month={Jan},
   pages={1157?1169}
}

@misc{clement2019use,
      title={On the Use of ArXiv as a Dataset}, 
      author={Colin B. Clement and Matthew Bierbaum and Kevin P. O'Keeffe and Alexander A. Alemi},
      year={2019},
      eprint={1905.00075},
      archivePrefix={arXiv},
      primaryClass={cs.IR}
}
@article{IbnMohammed:2021,
  doi = {10.1016/j.resconrec.2020.105169},
  url = {https://doi.org/10.1016/j.resconrec.2020.105169},
  year = {2021},
  month = jan,
  publisher = {Elsevier {BV}},
  volume = {164},
  pages = {105169},
  author = {T. Ibn-Mohammed and K.B. Mustapha and J. Godsell and Z. Adamu and K.A. Babatunde and D.D. Akintade and A. Acquaye and H. Fujii and M.M. Ndiaye and F.A. Yamoah and S.C.L. Koh},
  title = {A critical analysis of the impacts of {COVID}-19 on the global economy and ecosystems and opportunities for circular economy strategies},
  journal = {Resources,  Conservation and Recycling}
}

@misc{LatticeConferenceWebsite, 
	title={The 38th International Symposium on Lattice Field Theory (Lattice 2020)}, 
	url={https://indico.hiskp.uni-bonn.de/event/1/}, 
	journal={Helmholtz-Institut für Strahlen- und Kernphysik Indico Service (Indico)}
} 
@book{rasmussen:williams:2006,
  added-at = {2009-03-05T08:49:50.000+0100},
  author = {Rasmussen, C. E. and Williams, C. K. I.},
  biburl = {https://www.bibsonomy.org/bibtex/26771eaebbee7d852934f29aa33dea971/bcao},
  interhash = {72c030472023000e0bdeeb06081c3764},
  intrahash = {6771eaebbee7d852934f29aa33dea971},
  keywords = {},
  publisher = {MIT Press},
  timestamp = {2009-03-05T08:49:50.000+0100},
  title = {Gaussian Processes for Machine Learning},
  year = 2006
}

@article{hodlr,
    author = {{Ambikasaran}, S. and {Foreman-Mackey}, D. and
              {Greengard}, L. and {Hogg}, D.~W. and {O'Neil}, M.},
     title = "{Fast Direct Methods for Gaussian Processes}",
      year = 2014,
     month = mar,
       url = http://arxiv.org/abs/1403.6015
}

@article{Nicola:2020,
  doi = {10.1016/j.ijsu.2020.04.018},
  url = {https://doi.org/10.1016/j.ijsu.2020.04.018},
  year = {2020},
  month = jun,
  publisher = {Elsevier {BV}},
  volume = {78},
  pages = {185--193},
  author = {Maria Nicola and Zaid Alsafi and Catrin Sohrabi and Ahmed Kerwan and Ahmed Al-Jabir and Christos Iosifidis and Maliha Agha and Riaz Agha},
  title = {The socio-economic implications of the coronavirus pandemic ({COVID}-19): A review},
  journal = {International Journal of Surgery}
}

@article{Chu:2020,
  doi = {10.1093/jtm/taaa192},
  url = {https://doi.org/10.1093/jtm/taaa192},
  year = {2020},
  month = oct,
  publisher = {Oxford University Press ({OUP})},
  volume = {27},
  number = {7},
  author = {Isaac Yen-Hao Chu and Prima Alam and Heidi J Larson and Leesa Lin},
  title = {Social consequences of mass quarantine during epidemics: a systematic review with implications for the {COVID}-19 response},
  journal = {Journal of Travel Medicine}
}

### 2021-09-14 article/leftovers.tex
old upload eeek

	Long-term trends over 14 years show that academics are writing more papers, with more co-authors, and those papers are getting longer. 
	However, the increased quantity does not reflect quality: we find evidence that the level of intellectual contribution needed to warrant co-authorship on a paper has dropped in astronomy, \todo{and elsewhere?}. 
	We find that more leading and co-author papers in the first few years of an academic's career increases their chances to stay in academia, affirming the `\emph{publish or perish}' mantra, but only to a point: after an early career researcher has published enough papers, their chances of continuing in academia remain flat. At that point, they should stop writing papers and do something else.

Academics are writing more papers, and with more co-authors (Figure~\ref{fig:pre-prints-segmented-by-author-count}), with a steady decline of single author papers in every field. All papers are getting longer, too: the mean abstract length is increasing, and even the shortest abstracts each year are getting longer (Figure~\ref{fig:pre-prints-abstract-length}).

\section*{Longevity and authorship valuation}

%	\item Is this increased quantity of publications, and length of publications, a measure of increased quality of research?

We find that the number of authors on a paper has tended to increase in most fields between 2007 and 2020 (Figure~\ref{fig:number-of-authors-with-time}). These trends have been explained by the increasing complexity of research projects that requires expertise from different areas [REF]. If the increasing collaboration size is entirely due to research expertise, then we would expect that the expanding network size would mostly be reflected in established researchers that are experts in their field. Instead, we find that the minimum amount of work necessary to warrant co-authorship has decreased in some fields, which \todo{mostly} accounts for the increase in collaboration sizes with time.

Consider a motivated PhD supervisor with good intentions. If the supervisor encourages their student to contribute to a paper where the supervisor is already involved (but the supervisor is not leading), then the supervisor will be helping to improve the student's visibility (and perceived productivity) in an already competitive landscape. This is a common scenario in academia: an established researcher will assign some task to a PhD student, and the outcome of that task will contribute to a paper in preparation. As a consequence, both the student and supervisor are rightfully listed as co-authors instead of just the supervisor. If the valuation of `intellectual contribution necessary to warrant publication' remained stable with time, then we should expect that the number of co-author publications that overlap with the PhD supervisor(s) will also remain stable with time.

We identified two cohorts of early career researchers (ECRs) in astronomy that are separated by their career start date by five years (see Methods). The distribution of first-author (leading) publications between these cohorts is indistinguishable (\todo{some metrics}, Figure~\ref{fig:longevity}). The relative number of researchers with \todo{up to 5} co-author publications is also the same between both cohorts, but the more recent cohort has a long-tail of ECRs with \todo{X or more} co-author publications. These researchers would appear to be more productive, and to have a more extensive network of collaborators. When we separate the co-author publications to those that do, or do not, include the inferred PhD supervisor(s), it is clear that the increased number of co-author publications is not due to network expansion: those ECRs are publishing with their PhD supervisors and not independently expanding their network. While it is impossible to reliably identify a causal link here, it would appear that those ECR students with increased co-author publications, that overlap heavily with their supervisors, is perhaps due to the supervisors good intention of helping the student expand their network. While individual intentions are good, the net behavioural effect of this as a research field is that the intellectual merit necessary for publication can decrease on average.

\todo{Check whether the increased collaboration size is driven by experts being co-authors, or ECRs as being co-authors}

The data suggest that those extra publications has little or no impact on an ECR's likelihood of remaining in academia. We identified astronomy ECRs that started between 2010-2012 and separated them into two cohorts: those that continued to publish in astronomy, and those that did not. For every first-author paper written by an ECR in the first \todo{X} years, their likelihood of continuing in academia increased. However, this effect flattens out at \todo{$N \approx 6$}: an ECR with 6 first author publications is just as likely to continue in academia as an ECR with 10 first author publications. A similar effect is seen in co-author publications, although the rate is lower. These two effects are related: an ECR with 10 first-author papers and no co-author papers may look as employable (or un-employable) as an ECR with 10 co-author papers and no first-author papers, but for different reasons. Matching these cohorts with citation, network, and institution metrics would likely reveal a more predictive model for whether an individual ECR will remain in academia. But as a cohort, the probability of continuing in astronomy research flattens out at about 50\%: for astronomy ECRs is the top echelon of productivity, half of them will still leave academia.

\section*{Authorship valuation analysis}

\section*{Longevity analysis}

### 2022-06-06 data/process.py
Commit on new branch

@@ -1,0 +1,92 @@
+
+import json
+import os
+from tqdm import tqdm
+from astropy.table import Table
+from subprocess import getoutput
+
+arxiv_dataset_path = "arxiv-metadata-oai-snapshot.json"
+# Don't process any preprints from 1 June 2022 onwards.
+end_date = (2022, 6)
+
+dirname = os.path.dirname(os.path.abspath(__file__))
+relative_path = lambda basename: os.path.join(dirname, basename)
+
+with open(relative_path(arxiv_dataset_path), "r") as fp:
+    preprints = map(json.loads, fp.readlines())
+
+output = getoutput(f"wc {relative_path(arxiv_dataset_path)}")
+total = int(output.strip().split()[0])
+
+records = []
+# Add dummy line.
+records.append({
+    "id": "dummy",
+    "primary_parent_category": "",
+    "primary_category": "",
+    "num_categories": 0, 
+    "categories": "",
+    "num_authors": 0,
+    "num_affiliations": 0,
+    "words_in_abstract": 0
+})
+
+metadata = {}
+authors = {}
+for preprint in tqdm(preprints, total=total):
+    if end_date is not None:
+        yymm = preprint["id"].split(".")[0]
+        yy, mm = (int(f"20{yymm[:2]}"), int(yymm[2:]))
+        end_year, end_month = end_date
+        if yy >= end_year and mm >= end_month:
+            print(f"Reached {preprint['id']} stopping because end date is {end_date}")
+            break
+
+    affiliations = [author[-1] for author in preprint["authors_parsed"]]
+    metadata[preprint["id"]] = {
+        "title": preprint["title"],
+        "abstract": preprint["abstract"],
+        "affiliations": affiliations
+    }
+
+    these_authors = []
+    for author in preprint["authors_parsed"]:
+        keep = [name for name in author if name != "" and " and " not in name and list(set(name)) != ["I"]]
+        these_authors.append(keep[:2])
+
+    authors[preprint["id"]] = these_authors
+
+    created = preprint['versions'][0]["created"]
+    categories = preprint["categories"].split()
+    primary_parent_category = categories[0].split(".")[0]
+
+    record = {
+        "id": preprint["id"],
+        "primary_parent_category": primary_parent_category,
+        "primary_category": categories[0],
+        "num_categories": len(categories),
+        "categories": preprint["categories"],
+        "num_authors": len(preprint["authors_parsed"]),
[…]

### 2022-06-06 plot-new-authors-segmented-by-field.py

@@ -1,0 +1,345 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.ticker import MaxNLocator
+from tqdm import tqdm
+
+from plot_utils import (load_records, load_authors, load_metadata, get_decimal_date, pad_edges, unique_ify)
+
+matplotlib.style.use("matplotlibrc.paper")
+
+try:
+    (records, authors, metadata)
+except NameError:
+    records = load_records()
+    authors = load_authors()
+    metadata = load_metadata()
+else:
+    print("WARNING: Using pre-loaded data.")
+
+
+# Hmm. Let's see how many 'new' names entered a field with time.
+# That is to say: the name had never appeared in the literature until then.
+
+bins = np.arange(2007, max(records["created_decimal_year"]) + 1/12, 1/12)
+
+xlim = (bins[0], bins[-1])
+xlim = (2010, bins[-1])
+
+
+def get_new_authors(arxiv_ids, authors, callback=None):
+    d = np.array(list(map(get_decimal_date, arxiv_ids)))
+    idx = np.digitize(d, bins)
+    
+    y = np.zeros(bins.size - 1)
+    z = np.zeros(bins.size - 1)
+    has_appeared = []
+    for i, bin_edge in enumerate(tqdm(bins[:-1])):
+        mask = (idx == (i + 1))
+        authors_in_bin = []
+        for arxiv_id in arxiv_ids[mask]:
+            authors_on_this = list(map(unique_ify, authors[arxiv_id]))
+            authors_in_bin.extend(authors_on_this)
+            z[np.digitize(get_decimal_date(arxiv_id), bins) - 1] += 1
+            if callback is not None:
+                N = len(set(authors_on_this).difference(has_appeared))
+                if callback(bins[i], bins[i+1], N):
+                    md = metadata[arxiv_id]
+                    print(arxiv_id, N, md["title"])
+
+        new_names = set(authors_in_bin).difference(has_appeared)
+        y[i] = len(new_names)
+        
+        has_appeared.extend(list(new_names))
+
+    return (y, z)
+
+
+ignore_ppcs = ("econ", "eess")
+ppcs = sorted(list(set(records["primary_parent_category"]).difference(ignore_ppcs)))
+
[…]

### 2022-06-06 plot-pandemic-related-preprints.py

@@ -1,0 +1,145 @@
+
+import numpy as np
+import json
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.ticker import MaxNLocator
+from tqdm import tqdm
+
+from plot_utils import (
+    load_records,
+    load_authors,
+    load_metadata,
+    get_decimal_date,
+    unique_ify,
+    in_a_not_appeared_before_in_b
+)
+
+matplotlib.style.use("matplotlibrc.paper")
+
+try:
+    records, authors, metadata
+except NameError:
+    records = load_records()
+    authors = load_authors()
+    metadata = load_metadata()
+
+else:
+    print("WARNING: Using pre-loaded records.")
+
+
+ignore_ppcs = [
+    "eess", "econ", # These don't go back far enough
+    # These have ~zero instances throughout time:
+    "astro-ph",
+    "cond-mat", "gr-qc", "hep-ex", "hep-lat", "hep-ph", "hep-th",
+    "math-ph", "nlin", "nucl-ex", "nucl-th",
+    "quant-ph",
+    "math"
+]
+
+ppcs = sorted(list(set(records["primary_parent_category"]).difference(ignore_ppcs)))
+
+bin_max = max(records["created_decimal_year"])
+
+bins = np.arange(2011, bin_max, 1/12)
+counts = {}
+totals = {}
+
+patterns = ("pandemic", "covid", "sars-cov-2", "lockdown", "coronavirus")
+
+q_bio_2020_ids = []
+
+for ppc in ppcs:
+    print(ppc)
+
+    counts[ppc] = np.zeros_like(bins)
+    totals[ppc] = np.zeros_like(bins)
+    
+    mask = (records["primary_parent_category"] == ppc)
+
+    for arxiv_id in tqdm(records["id"][mask]):
+        date = get_decimal_date(arxiv_id)
+        if date < bins[0] or date > bins[-1]:
+            continue
+
+        idx = np.digitize(date, bins)
+        totals[ppc][idx] += 1
+
+        md = metadata[arxiv_id]
+        context = f"{md['title']} {md['abstract']}".lower()
+        for pattern in patterns:
+            if pattern in context:
+                counts[ppc][idx] += 1
+                if ppc == "q-bio" and date > 2020 and date < 2021:
[…]

### 2022-06-06 plot-pre-prints-segmented-by-field.py

@@ -1,0 +1,351 @@
+
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+import george
+import scipy.optimize as op
+from george import kernels
+from george.modeling import Model
+from plot_utils import load_records, load_metadata, get_decimal_date, pad_edges
+from statsmodels.tsa.arima.model import ARIMA
+from matplotlib.ticker import MaxNLocator
+
+matplotlib.style.use("matplotlibrc.paper")
+
+try:
+    records, metadata
+except NameError:
+    records = load_records()
+    metadata = load_metadata()
+
+else:
+    print("WARNING: Using pre-loaded records.")
+
+
+show_arima = True
+
+bin_width = 1/12
+try:
+    counts
+except NameError:
+        
+    # Plot submissions by time for each field.
+    ignore_ppcs = ("econ", "eess")
+
+    ppcs = sorted(list(set(records["primary_parent_category"]).difference(ignore_ppcs)))
+
+    counts = {}
+    totals = {}
+
+    bins = np.arange(2010, max(records["created_decimal_year"]) + bin_width, bin_width)
+
+    xlim = (bins[0], int(np.ceil(bins[-1])))
+    patterns = ("pandemic", "covid", "sars-cov-2", "lockdown")
+
+    for ppc in ppcs:
+        print(ppc)
+
+        counts[ppc] = np.zeros_like(bins)
+        totals[ppc] = np.zeros_like(bins)
+        
+        mask = (records["primary_parent_category"] == ppc)
+
+        for arxiv_id in records["id"][mask]:
+            date = get_decimal_date(arxiv_id)
+            if date < bins[0] or date > bins[-1]:
+                continue
+
+            idx = np.digitize(date, bins)
+            totals[ppc][idx] += 1
+
+            md = metadata[arxiv_id]
+            context = f"{md['title']} {md['abstract']}".lower()
+            for pattern in patterns:
+                if pattern in context:
+                    counts[ppc][idx] += 1
+                    break
+
+else:
+    print("WARNING: Using pre-determined counts")
+
+
+xlim = (2010, 2023)
+
+class LineModel(Model):
+    parameter_names = ("m", "b")
[…]

### 2022-06-06 plot_utils.py

@@ -1,0 +1,210 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import json
+from astropy.table import Table
+from astropy.time import Time
+from collections import OrderedDict
+from tqdm import tqdm
+
+def load_metadata(path="data/metadata.json"):
+    with open(path, "r") as fp:
+        metadata = json.load(fp)
+    return metadata
+
+
+def load_records(path="data/records.csv"):
+
+    records = Table.read(path)
+    # Ignore the dummy line, if it exists.
+    if records["id"][0] == "dummy":
+        records = records[1:]
+
+    # Exclude old-style
+    #mask = np.array(["/" in id_ for id_ in records["id"]])
+    #records = records[~mask]
+
+    # Let's add some useful columns.
+    if "created" in records.dtype.names:
+        records["created_decimal_year"] = Time(records["created"]).decimalyear
+    else:
+        records["created_decimal_year"] = list(map(get_decimal_date, records["id"]))
+
+    return records
+
+
+def pad_edges(bin_edges, counts):
+    if bin_edges.size == (counts.size + 1):
+        x = np.hstack([
+            bin_edges[0],
+            bin_edges[:-1] + 0.5 * np.diff(bin_edges),
+            bin_edges[-1] 
+        ])
+        y = np.hstack([
+            counts[0],
+            counts,
+            counts[-1]
+        ])
+
+    else:
+        x = np.hstack([
+            bin_edges[0] - 0.5 * np.diff(bin_edges[:2]),
+            bin_edges,
+            bin_edges[-1] + 0.5 * np.diff(bin_edges[-2:])
+        ])
+        y = np.hstack([
+            counts[0],
+            counts,
+            counts[-1]
+        ])
+
+    return (x, y)
+
+
+def load_authors(path="data/authors.json"):
+    with open(path, "r") as fp:
+        authors = json.load(fp)
+    return authors
+    
+
+# How we will define a 'unique' author.
+def unique_ify(author_name, debug=False):
+    try:
+        last, given = author_name
+    except:
+        try:
+            response = author_name[0]
+        except:
+            response = ""
+
+        if debug: print(response)
+        return response
+
+    
+    if last.count(".") > 0 and given.count(" ") == 1:
+        if debug: print(author_name, given, last)
+        #0704.1276 Pavlenko Elena P.
+        
+        given, last = given.split()
[…]

### 2022-06-06 article/ms.tex
add article files

--- 2021-09-14 article/arxiv-ms.tex
@@
-\documentclass[a4paper,12pt]{article}
-\usepackage[labelfont=bf,font=small]{caption}
+%% Author_tex.tex
+%% V1.0
+%% 2012/13/12
+%% developed by Techset
+%%
+%% This file describes the coding for rsproca.cls
 
-\usepackage{geometry}
-\geometry{a4paper,left=25mm,top=25mm,right=25mm,bottom=25mm}
+\documentclass[]{rsos}%%%%where rsos is the template name
 
-\usepackage[pdftex]{graphicx}
-\usepackage{url}
-% Only include extra packages if you really need them. Common packages are:
-\usepackage{graphicx}	% Including figure files
-\let\iint\relax
-\let\iiint\relax
-\let\iiiint\relax
-\let\idotsint\relax
-\usepackage{amsmath}
-\usepackage{amssymb}
-
-\usepackage{pdflscape}
-\usepackage[utf8]{inputenc}
-\usepackage{xfrac}
-\usepackage{color}
-\usepackage{rotating}
-\usepackage{tabularx}
+%%%% *** Do not adjust lengths that control margins, column widths, etc. ***
 
@@
 \definecolor{mygray}{gray}{0.6}
-\newcommand{\todo}[1]{\textcolor{red}{#1}}
+
+%%%%%%%%%%% Defining Enunciations  %%%%%%%%%%%
 \newcommand{\arxiv}{arXiv}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\usepackage{setspace}
+\begin{document}
 
-\usepackage{lineno}
-%\linenumbers
+%%%% Article title to be placed here
+\title{The impact of the COVID-19 pandemic on academic productivity}
 
-\usepackage{amsmath}
-%\usepackage[superscript,biblabel]{cite}
+\author{%%%% Author details
+A. R. Casey$^{1,2}$,  I. Mandel$^{1,3,4}$ and P. K. Ray$^{5}$}
 
-\usepackage[resetlabels,labeled]{multibib}
-\newcites{main,method}{{},{}} 
-\usepackage[superscript,nomove]{cite}
-
-\usepackage{xpatch}
-\newcommand\removebibheader
-  {\@{}
-    {\section*{}%
-     \@{}{}%
-    }{}{}{}%
+%%%%%%%%% Insert author address here
+\address{$^1$School of Physics \& Astronomy, Monash University, Clayton 3800, Victoria, Australia\\
+$^2$ARC Centre of Excellence for Astrophysics in Three Dimensions (ASTRO-3D), Australia\\
+$^3$OzGrav, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Australia\\
+$^4$Institute of Gravitational Wave Astronomy and School of Physics and Astronomy, University of Birmingham, Birmingham, B15 2TT, United Kingdom\\
+$^5$Department of Mathematics, Imperial College London, London, United Kingdom
 }
 
-\title{The impact of the COVID-19 pandemic on academic productivity}
+%%%% Subject entries to be placed here %%%%
+\subject{xxxxx, xxxxx, xxxx}
 
-\author{Andrew R. Casey$^{1,2}$\footnote{andrew.casey@monash.edu},
-        Ilya Mandel$^{1,2,3,4}$,
-        Prasun K. Ray$^{5}$\\
-\normalsize{$^{1}$School of Physics \& Astronomy, Monash University, Clayton 3800, Victoria, Australia}\\
-\normalsize{$^{2}$ARC Centre of Excellence for Astrophysics in Three Dimensions (ASTRO-3D), Australia}\\
-\normalsize{$^{3}$OzGrav, Australian Research Council Centre of Excellence}\\
-\normalsize{for Gravitational Wave Discovery, Australia}\\
-\normalsize{$^{4}$Institute of Gravitational Wave Astronomy and School of Physics and Astronomy,}\\
-\normalsize{University of Birmingham, Birmingham, B15 2TT, United Kingdom}\\
-\normalsize{$^{5}$Department of Mathematics, Imperial College London, London, United Kingdom}
-}
+%%%% Keyword entries to be placed here %%%%
+\keywords{COVID, pre-print, academic productivity}
 
-\renewcommand*{\thefootnote}{\arabic{footnote}}
+%%%% Insert corresponding author and its email address}
+\corres{Andrew R. Casey\\
+\email{andrew.casey@monash.edu}}
 
-\date{}
-\doublespacing
-\begin{document}
-\newpage
-\setcounter{page}{1}
-\resetlinenumber[1]
+%%%% Abstract text to be placed here %%%%%%%%%%%%
+\begin{abstract}
+`Publish or perish’ is an expression describing the pressure on academics to consistently publish research to ensure a successful career in academia. 
+With a global pandemic that has changed the world, how has it changed academic productivity? 
+Here we show that academics are posting more papers on the arXiv pre-print server than would be expected had the pandemic not occurred: 168,630 were posted in 2020, a +12.6\% change from 2019 and $+1.4\sigma$ deviation above the 162,577 predicted by our Gaussian process model.
+Our results also indicate that most individual fields were not significantly affected by the pandemic, however we do observe large declines in output in areas of high-energy physics where laboratory closures and the cancellation of a key conference have played important roles.
+The most significant increase occurred in quantitative biology where a 50\%~increase ($+8\sigma$) in pre-prints was driven by research related to the COVID-19 pandemic.
+Some of these `COVID papers' are by biologists using arXiv for the first time, but many are written by researchers from other fields (e.g., physicists, mathematicians). 
+As of mid-2021, quantitative biology pre-prints have returned to pre-pandemic levels, but 20\% of the research in this field is now related to the COVID-19 pandemic, demonstrating a strong shift in research focus.
+\end{abstract}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\noindent \textbf{`\emph{Publish or perish}' is an expression describing the pressure on academics to consistently publish research to ensure a successful career in academia. 
-A global pandemic has changed the world, but how has it changed academic productivity?
-Here we show that academics are posting just as many publications on the \arxiv\ pre-print server as if there were no pandemic: 168,630 were posted in 2020, a $\mathbf{+12.6\%}$ change from 2019 and a $\mathbf{+1.4\sigma}$ deviation above the predicted 162,577 $\mathbf{\pm}$ 4,393.
-However, some immediate impacts are visible in individual research fields.
-Conference cancellations have led to sharp drops in pre-prints, but laboratory closures have had mixed effects.
-Only some experimental fields show mild declines in outputs, with most being consistent on previous years or even increasing above model expectations.
-The most significant change is a 50\% increase ($\mathbf{+8\sigma}$) in quantitative biology research, all related to the COVID-19 pandemic.
-Some of these publications are by biologists using \arxiv\ for the first time, and some are written by researchers from other fields (e.g., physicists, mathematicians).
-While quantitative biology pre-prints have returned to pre-pandemic levels, 20\% of the research in this field is now focussed on the COVID-19 pandemic, demonstrating a strong shift in research focus.
-}\\
+%%%%%%%%%% Insert the texts which can accomdate on firstpage in the tag "fmtext" %%%%%
 
 \noindent Peer-reviewed publications are the most common measure of productivity in academia.
-The \arxiv\cite{Ginsparg:2011} ({https://arxiv.org}) is a distribution service for research publications before they are printed in a journal (i.e., a pre-print).
-A pre-print on \arxiv\ does not ensure that the contents have passed peer-review, but most material on \arxiv\ eventually goes through peer-review because it is now standard in many research fields to post to \arxiv\ either during or after the peer-review process \cite{Lariviere:2014}.
+The \arxiv\cite{Ginsparg:2011} ({https://arxiv.org}) is a service for the distribution of research publications (pre-prints) prior to their appearance in a journal.
+The presence of a pre-print on \arxiv\ does not ensure that the contents have passed peer-review, but most material on \arxiv\ eventually goes through peer-review because it is now standard in many research fields to post to \arxiv\ either during or after the peer-review process \cite{Lariviere:2014}.
 For this reason, the number of pre-prints posted to \arxiv\ approximates the number of peer-reviewed publications written in these fields.
@@
 Here we make quantitative comparisons on the number of pre-prints posted to \arxiv\ before and during the pandemic.
-While we investigate the impact that the COVID-19 pandemic has had on different fields of research, we highlight that we are unable to identify or differentiate between authors and communities whose productivity has been significantly harmed by the pandemic, and those who were largely unscathed.
-The COVID-19 pandemic has impacted the population in unequal ways\cite{Nicola:2020,Chu:2020,Viglione:2020,Gewen:2020,IbnMohammed:2021,King:2021}.
-Generational inequality, career stage, personal circumstances, carer responsibilities, work environments, places of employment, and many other factors, all significantly contribute to the disproportionate and unequal impact the pandemic may have had on a scientist's capacity to conduct research.
-While this limited analysis does not address these issues, it is important to consider that if a community has not yet demonstrated a significant change in productivity, many researchers have faced significant challenges and suffered physically, mentally, emotionally, and professionally.
+It is important to recognise that the pandemic has affected both the general population\cite{Nicola:2020,Chu:2020,IbnMohammed:2021} and scientific community\cite{Viglione:2020,Gewen:2020,King:2021} in complex and unequal ways.
+Scientists have faced substantial challenges driven by a large number of factors including: generational inequality, carer responsibilities, modified work environments, and other personal and professional circumstances.
+Our analysis does not directly consider these important factors, nor does it account for groups or individuals disproportionately affected by the pandemic.
+We have instead adopted a broad perspective and present new quantitative results for entire scientific fields.
+The aim is to provide a “large-scale” view of scientific productivity which will serve as a foundation for future studies on the detailed impacts of COVID-19 on academic work.
 
@@
 The metadata includes the creation date, research field(s), title, author name(s), abstract, and other miscellaneous information\cite{Clement:2019}.
-There is an increasing number of pre-prints posted to \arxiv\ each year in nearly every field (Figure~1).
+There is an increasing number of pre-prints posted to \arxiv\ each year in nearly every field (Figure~\ref{fig:1}).
 These long-term trends are relatively predictable from year to year, allowing us to quantify any change in academic productivity due to the COVID-19 pandemic.
-We used the number of publications from January 2015 to December 2019 to predict the expected number of monthly pre-prints in each field after January 2020.
-We modelled the number of monthly pre-prints in each field with a straight line to capture long-term (linear) trends, and used a Gaussian Process with a quasi-periodic kernel function\cite{Rasmussen:2006,Ambikasaran:2014} to capture seasonal variations (Figure~1).
-In nearly all research fields the number of pre-prints posted since January 2020 agree excellently with the model predictions (Table~1), indicating no collective impact on academic productivity due to the COVID-19 pandemic.
-The total number of \arxiv\ pre-prints in 2020 across all fields (168,630) exceeds our model predictions ($162{,}577 \pm 4{,}393$; $+1.4\sigma$).
+We used the number of publications from January 2015 to December 2019 to train a predictive Gaussian process model\cite{Rasmussen:2006}.
+This model treats the number of monthly pre-prints in a research field as a sequence of normally distributed random variables.
+The expected number of pre-prints was assumed to increase linearly with time, and a quasi-periodic kernel function\cite{Rasmussen:2006,Ambikasaran:2014} was used to capture seasonal variations (Figure~\ref{fig:1}).
+In nearly all research fields, the number of pre-prints posted since January 2020 agree closely with the model predictions (Table~1) indicating no meaningful collective impact on academic productivity due to the COVID-19 pandemic.
+In fact, the total number of \arxiv\ pre-prints in 2020 across all fields (168,630) exceeds our model predictions ($162{,}577 \pm 4{,}393$; $+1.4\sigma$).
+
+\begin{figure}[!h]
+\includegraphics[width=0.95\linewidth]{pre-prints-segmented-by-field}
+\caption{There has been no significant change in monthly \arxiv\ pre-prints (black) since the COVID-19 pandemic began.
+The expected number of pre-prints from our Gaussian Process model is shown in red, and the red region shows the uncertainty in that expectation.
+The dashed light grey line shows predictions from an autoregressive integrated moving average model.}
+\label{fig:1}
+\end{figure}
+
+Border closures and travel restrictions have forced many academic conferences to be held online-only, or to be cancelled.
+Hosting conferences online creates a mix of effects, as forgoing in-person experiences benefits those who could not otherwise participate due to visa issues or travel costs\cite{Guinnessy:2021}.
+The immediate impact of cancelling a conference is readily apparent in pre-print counts in lattice physics (Figure~\ref{fig:1}; \arxiv\ subject code \texttt{hep-lat}, see Methods for explanation), where most pre-prints are posted around December each year as conference proceedings from the International Symposium on Lattice Field Theory.
+In 2020 the conference was cancelled\cite{LatticeConferenceWebsite}, and no accompanying pre-prints exist.
+The impact of conference cancellations and travel restrictions on academic research is likely to extend well beyond the output of a single field in a particular year since discussions at conferences spark new research ideas and collaborations that can lead to publications many months or years later\cite{Wang:2017}.
+
+Experimental research projects often require specialised laboratories, or data to be collected over many years.
+The pandemic has forced many laboratories to close or operate with restricted access\cite{PhysicsWorld,ScienceMag,Castelvecchi}, which could lead to long-lasting delays in ongoing experiments or immediate drops in publication rates in part due to difficulty in accessing completed (or nearly completed) experiment data.
+There is some evidence of this in pre-print numbers already, with {20\%} fewer pre-prints in high energy physics (\texttt{hep-ex}) in 2020 than expected by our model (observed 615; predicted 818 $\pm$ 94).
+But declines are not ubiquitous across experimental fields.
+The closest comparable field of phenomenology in high energy physics (\texttt{hep-ph}) showed no substantial decline.
+Neither did nuclear experimental research (\texttt{nucl-ex}).
+And throughout 2020, research in condensed matter (\texttt{cond-mat}) saw a $1.8\sigma$ increase above our model predictions (observed 16,188; predicted 15,325 $\pm$ 492).
+Segmenting condensed matter pre-prints by research topic shows that most of this increase (in \texttt{cond-mat}) was driven by a 30\% increase in material science pre-prints (\texttt{cond-mat.mtrl-sci}).
+
+The field of quantitative biology (\texttt{q-bio}) showed the largest increase in pre-prints in 2020 above what is expected by our model.
+There were {2,790} quantitative biology pre-prints in 2020, {50\%} above the previous year, representing a $+8\sigma$ deviation from our model predictions (predicted 1,914 $\pm$ 109, 876 fewer than observed).
+This increase is explainable by an increase in COVID-19 related research, as there were 864 quantitative biology pre-prints in 2020 with pandemic-related terms in their title or abstract (see Methods), and just 39 in the decade prior.
+Indeed, in {April 2020} nearly 60\% of the \arxiv\ pre-prints in quantitative biology were related to the pandemic (Figure~\ref{fig:2}).
+Pandemic-related pre-prints also appeared in other fields (computer science, physics, statistics, and quantitative finance, all peaking around {April 2020}), but these papers constituted fewer than 10\% of the total pre-prints in these fields.
+As of mid-2021 the number of monthly pre-prints in quantitative biology has dropped to pre-pandemic levels.
+Approximately 20\% of current pre-prints in quantitative biology are related to the ongoing pandemic, representing a strong shift in ongoing research focus.
+
+\begin{figure}
+ \includegraphics[width=\linewidth]{pandemic-related-preprints}
+\caption{Pre-prints related to the COVID-19 pandemic peaked in April 2020, accounting for nearly 60\% the quantitative biology (\texttt{q-bio}) pre-prints in that month.
+Shown are five other research fields with the highest number of pandemic-related pre-prints in 2020.
+Monthly pre-prints are shown on the left, and the percent of pandemic-related pre-prints in that research field is shown on the right.}
+\label{fig:2}
+\end{figure}
+
+\renewcommand{\thefootnote}{$\dagger$}
+
+An increase in quantitative biology\footnote{Quantitative biology is only a sub-field of biology, and most biology pre-prints are posted to bioRxiv ({https://biorxiv.org}).} research during a global pandemic is unsurprising.
+However, the question arises if this increase was driven by existing researchers in the field, or by authors from other fields entering quantitative biology due to the pandemic.
+In 2020 there was a peak in the number of new authors appearing for the first time in the quantitative biology literature (Figure~\ref{fig:3}) while the number of new authors in all other fields remained steady.
+The increase in pre-prints (and new authors) is driven by small (1-4) groups of authors, where many had never posted pre-prints to quantitative biology before.
+
+\begin{figure}[!h]
+\begin{center}
+	\includegraphics[width=0.95\linewidth]{new-authors-segmented-by-field-combined}
+\end{center}
+\caption{The number of new authors appearing in \arxiv\ pre-prints per month (grey) peaked in quantitative biology in 2020.
+Here we show new authors since 2010, but we used data since 2007 to determine if an author was new.
+The slow change in new authors  approximates the net number of new authors joining the field.
+The number of new authors appearing per number of pre-prints is shown in light blue, with identical limits on the right-hand y-axis.
+This highlights authorship differences between fields (e.g., \texttt{hep-ex} and \texttt{hep-lat}) for large experiments.}
+\label{fig:3}
+\end{figure}
+
+One quarter of the excess pre-prints posted to quantitative biology during 2020 were both related to the pandemic and written by authors who had never appeared in the quantitative biology literature (234 of the 876 excess pre-prints in \texttt{q-bio}).
+A careful examination reveals that about half of these (45\%; 107/234) were written by mathematicians or physicists, with another 26\% (61/234) led by engineers, computer scientists, or economists.
+These pre-prints tend to focus on modelling the COVID-19 outbreak using public data sets, rather than quantitative biology research that requires more expert domain knowledge.
+About 3\% (8/234) were led by established biologists who have not used the \arxiv\ before and are now doing so presumably to help ensure that their COVID-19 research is more widely available.
+This fraction is likely higher among pre-prints where at least one author has appeared before in the quantitative biology literature.
+This represents a genuine increase in new researchers using the \arxiv, and a boost to making quantitative biology research more widely accessible.
+
+% What to say of the remaining 75%?
+
+%However, a careful examination of pandemic-related pre-prints in quantitative biology posted in 2020 shows that many were indeed written by researchers from other fields (primarily physicists and mathematicians), who had never posted about quantitative biology.
+
+%Some of these new authors in quantitative biology have never appeared in any other \arxiv\ fields before, either.
+A close examination of 234 pre-prints posted during 2020 to \texttt{q-bio.PE} where no authors had posted to \texttt{q-bio} before shows that
+
+%A close examination of these pre-prints reveals that many are
+
+% 234 papers in 2020 in q-bio.PE written about the pandemic where no author had ever appeared before in the q-bio literature.
+% among these, 8 were led by authors who were domain experts in quantitative biology or a related field, and were 'new' to posting on arxiv
+% among these 234, 107 (45%) were led by mathematicians or physicists, with another 61/234 (26%) led by engineers, computer scientists, or economists.
+
+% [('mathematics', 58),
+% ('physics', 41), + astrophysics, 8 ==> 49
+% ('engineering', 26),
+% ('computer science', 19),
+% ('economics', 16),
+% close examination of a random sample of 50 of these "new" authors and their academic affiliations found x to be established biologists who had not used the arXiv before while the rest were researchers from other fields (primarily physics and mathematics) who had not previously posted quantitative biology papers to the arXiv.
 
 \begin{table}
+       \caption{The number of \arxiv\ pre-prints per primary research field in 2020 compared to 2019, and model predictions (see Methods).}
 \begin{center}
@@
     	\arxiv\ & Field of research & Pre-prints & Yearly & Predicted & Significance \\
-	code &&(2020)& change (\%)&(2020)&($\sigma$)\\
+	code &&(2020)& change (\%)&(2020)& ($\sigma$)\\
       \hline
@@
     \end{tabular}
-       \caption{The number of \arxiv\ pre-prints per primary research field in 2020 compared to 2019, and model predictions (see Methods).}
   \end{center}
@@
 
-Border closures and travel restrictions have forced many academic conferences to be held online-only, or to be cancelled.
-Hosting conferences online creates a mix of effects, as forgoing in-person experiences benefits those who could not otherwise participate due to visa issues or travel costs\cite{Guinnessy:2021}.
-The immediate impact of cancelling a conference is readily apparent in pre-print counts in lattice physics (Figure~1; \arxiv\ subject code \texttt{hep-lat}, see Methods for explanation), where most pre-prints are posted around December each year as conference proceedings from the International Symposium on Lattice Field Theory.
-In 2020 the conference was cancelled\cite{LatticeConferenceWebsite} and no accompanying pre-prints exist.
-However, the impact on research from travel restrictions is likely to be much longer than what is represented by the drop in lattice physics pre-prints.
-Discussions at conferences or collaborative visits frequently spark new research ideas that might lead to a publication many months or years later.
-
-Experimental research projects often require specialised laboratories, or data to be collected over many years.
-The pandemic has forced many laboratories to close or operate with restricted access\cite{PhysicsWorld,ScienceMag}, which could lead to long-lasting delays in ongoing experiments or immediate drops in publication rates in part due to difficulty in accessing completed (or nearly completed) experiment data.
-There is some evidence of this in pre-print numbers already, with {20\%} fewer pre-prints in high energy physics (\texttt{hep-ex}) in 2020 than expected by our model (observed 615; predicted 818 $\pm$ 94).
-But declines are not ubiquitous across experimental fields.
-The closest comparable field of phenomenology in high energy physics (\texttt{hep-ph}) showed no substantial decline.
-Neither did nuclear experimental research (\texttt{nucl-ex}).
-And throughout 2020, research in condensed matter (\texttt{cond-mat}) saw a $1.8\sigma$ increase above our model predictions (observed 16,188; predicted 15,325 $\pm$ 492).
-Segmenting condensed matter pre-prints by research topic shows that most of this increase (in \texttt{cond-mat}) was driven by a 30\% increase in material science pre-prints (\texttt{cond-mat.mtrl-sci}).
-
-The field of quantitative biology (\texttt{q-bio}) research showed the largest increase in pre-prints in 2020 above what is expected by our model.
-There were {2,790} quantitative biology pre-prints in 2020, {50\%} above the previous year, representing a $+8\sigma$ deviation from our model predictions (predicted 1,914 $\pm$ 109, 876 fewer than observed).
-This increase is explainable by an increase in COVID-19 related research, as there were 864 quantitative biology pre-prints in 2020 with pandemic-related terms in their title or abstract (see Methods), and just 39 in the decade prior.
-Indeed, in {April 2020} nearly 60\% of the \arxiv\ pre-prints in quantitative biology were related to the pandemic (Figure~2).
-Pandemic-related pre-prints also appeared in other fields (computer science, physics, statistics, and quantitative finance, all peaking around {April 2020}), but pandemic-related pre-prints constituted fewer than 10\% of the total pre-prints in these fields.
-As of mid-2021 the number of monthly pre-prints in quantitative biology has dropped to pre-pandemic levels.
-Approximately 20\% of current pre-prints in quantitative biology are related to the ongoing pandemic, representing a strong shift in ongoing research focus.
-
-\renewcommand{\thefootnote}{$\dagger$}
-
-An increase in quantitative biology\footnote{Quantitative biology is only a sub-field of biology, and most biology pre-prints are posted to bioRxiv ({https://biorxiv.org}).} research during a global pandemic is unsurprising.
-However, the question arises if this increase was driven by existing researchers in the field, or by authors from other fields entering quantitative biology due to the pandemic.
-In 2020 there was a peak in the number of new authors appearing for the first time in the quantitative biology literature (Figure~3) while the number of new authors in all other fields remained steady.
-The increase in pre-prints (and new authors) is driven by small (1-4) groups of authors, where many had never posted pre-prints to quantitative biology before.
-
-One quarter of the excess pre-prints posted to quantitative biology during 2020 were related to the pandemic and all authors had never appeared in the quantitative biology literature (234 of the 876 excess pre-prints in \texttt{q-bio}).
-A careful examination reveals that about half of these (45\%; 107/234) were written by mathematicians or physicists, with another 26\% (61/234) led by engineers, computer scientists, or economists.
-These pre-prints tend to focus on modelling the COVID-19 outbreak using public data sets, rather than quantitative biology research that requires more expert domain knowledge.
-About 3\% (8/234) were led by established biologists who have not used the \arxiv\ before and are now doing so presumably to help ensure that their COVID-19 research is more widely available.
-This fraction is likely higher among pre-prints where at least one author has appeared before in the quantitative biology literature.
-This represents a genuine increase in new researchers using the \arxiv, and a boost to making quantitative biology research more widely accessible.
-
-% What to say of the remaining 75%?
-
-%However, a careful examination of pandemic-related pre-prints in quantitative biology posted in 2020 shows that many were indeed written by researchers from other fields (primarily physicists and mathematicians), who had never posted about quantitative biology.
-
-%Some of these new authors in quantitative biology have never appeared in any other \arxiv\ fields before, either.
-A close examination of 234 pre-prints posted during 2020 to \texttt{q-bio.PE} where no authors had posted to \texttt{q-bio} before shows that
-
-%A close examination of these pre-prints reveals that many are
-
-% 234 papers in 2020 in q-bio.PE written about the pandemic where no author had ever appeared before in the q-bio literature.
-% among these, 8 were led by authors who were domain experts in quantitative biology or a related field, and were 'new' to posting on arxiv
-% among these 234, 107 (45%) were led by mathematicians or physicists, with another 61/234 (26%) led by engineers, computer scientists, or economists.
-
-% [('mathematics', 58),
-% ('physics', 41), + astrophysics, 8 ==> 49
-% ('engineering', 26),
-% ('computer science', 19),
-% ('economics', 16),
-% close examination of a random sample of 50 of these "new" authors and their academic affiliations found x to be established biologists who had not used the arXiv before while the rest were researchers from other fields (primarily physics and mathematics) who had not previously posted quantitative biology papers to the arXiv.
-
 The pandemic has also created distinct challenges when interpreting trends in \arxiv\ data.
 For example, there was a large spike in papers submitted to the \texttt{stat} category in June 2020 which one might assume was driven by COVID-related work given what we have observed in \texttt{q-bio}.
-However, the number of COVID-related papers posted in this period is modest, and the key factor appears to be delayed paper submission deadlines for two large conferences (NeurIPS and the Joint Statistical Meetings).
+However, the number of COVID-related papers posted in this period is modest (Figure~\ref{fig:2}), and the key factor appears to be delayed paper submission deadlines for two large conferences (NeurIPS and the Joint Statistical Meetings).
 Both conferences have had submission deadlines in mid-May in recent years, but in 2020, their deadlines were moved to early June.
@@
 The relatively long timescales of academic research would suggest that the full impact on academic productivity due to the COVID-19 pandemic is yet to be seen.
-The approach applied here will be a useful quantitative approach to return to in subsequent years to explore the longer-term impact of the pandemic.
-
-\begin{thebibliography}{10}
-
-\bibitem{Ginsparg:2011}
-{Ginsparg}. P., {ArXiv at 20}, \emph{Nature}, \textbf{476}, 7359, doi:10.1038/476145a \newblock (2011).
-
-\bibitem{Lariviere:2014}
-{Larivi{\'e}re, V., et al.,} {arXiv E-prints and the journal of record: An analysis of roles and relationships}, \emph{Journal of the Association for Information Science and Technology}, \textbf{65}, 6, doi:10.1002/asi.23044 \newblock (2014).
-
-\bibitem{Nicola:2020}
-{Nicola, M., et al.,} {The socio-economic implications of the coronavirus pandemic ({COVID}-19): A review}, \emph{International Journal of Surgery}, \textbf{78}, 185--193, doi:10.1016/j.ijsu.2020.04.018 \newblock (2020).
-
-\bibitem{Chu:2020}
-{Yen-Hao Chu, I., Prima, A., Larson, H.~J., Leesa, L.}
-{Social consequences of mass quarantine during epidemics: a systematic review with implications for the {COVID}-19 response}, \emph{Journal of Travel Medicine}, \textbf{27}, 7, doi:10.1093/jtm/taaa192 \newblock (2020).
-
-\bibitem{Viglione:2020}
-{Viglione, G.}, {Are women publishing less during the pandemic? Here?s what the data say}. \emph{Nature}, \textbf{581}, 356-366, doi:10.1038/d41586-020-01294-9 \newblock (2020).
-
-\bibitem{Gewen:2020},
-{Gewen, V.}, {The career cost of COVID-19 to female researchers, and how science should respond}. \emph{Nature}, \textbf{583}, 867--869, doi:10.1038/d41586-020-02183-x \newblock (2020).
-
-\bibitem{IbnMohammed:2021}
-{Ibn-Mohammed, T., et al.,} {A critical analysis of the impacts of {COVID}-19 on the global economy and ecosystems and opportunities for circular economy strategies}, \emph{Resources, Conservation and Recycling}, \textbf{164}, 105169, doi:1016/j.resconrec.2020.105169 \newblock (2021).
-
-\bibitem{King:2021}
-{King, M.~M., Frederickson M.~E.}, {The Pandemic Penalty: The Gendered Effects of COVID-19 on Scientific Productivity}. \emph{Socius}. January 2021. doi:10.1177/23780231211006977 \newblock (2021).
-
-\bibitem{Clement:2019}
-{Clement, C.~B., Bierbaum, M., O'Keeffe, K.~P., Alemi, A.~A.,}
-{On the Use of ArXiv as a Dataset},
-Preprint at https://arxiv.org/abs/1905.00075 \newblock (2019).
-
-\bibitem{Rasmussen:2006}
-{Rasmussen, C.~E., Williams, C.~K.~I.,}
-{Gaussian Processes for Machine Learning}, \emph{MIT Press}, ISBN 026218253X \newblock (2016).
-
-\bibitem{Ambikasaran:2014}
-{Ambikasaran, S., Foreman-Mackey, D., Greengard, L., Hogg, D.~W., O'Neil, M.,}
-{Fast Direct Methods for Gaussian Processes}, Preprint at http://arxiv.org/abs/1403.6015 \newblock (2014).
-
-\bibitem{Guinnessy:2021}
-Guinnessy, P.
-(Physics Today).
-2021.
-\emph{Physicists gathered virtually for APS March Meeting}.
-[online] Available at {https://physicstoday.scitation.org/do/10.1063/PT.6.3.20210405a/full} [Accessed 17 June 2021].
-
-\bibitem{LatticeConferenceWebsite}
-Helmholtz-Institut f\"ur Strahlen- und Kernphysik Indico Service (Indico).
-2021.
-\emph{The 38th International Symposium on Lattice Field Theory (Lattice 2020)}.
-[online] Available at: {https://indico.hiskp.uni-bonn.de/event/1/} [Accessed 3 June 2021].
-
-\bibitem{PhysicsWorld}
-physicsworld (IOP Publishing).
-2021.
-\emph{Critical research hit as COVID-19 forces physics labs to close}.
-[online] Available at {https://physicsworld.com/a/critical-research-hit-as-covid-19-forces-physics-labs-to-close/} [Accessed 4 June 2021].
-
-\bibitem{ScienceMag}
-Science Magazine (AAAS).
-2021.
-\emph{Amid pandemic, Energy Department labs close to tens of thousands of users.} [online] Available at {https://www.sciencemag.org/news/2020/03/amid-pandemic-energy-department-labs-close-tens-thousands-users} [Accessed 4 June 2021].
-
-\bibitem{Fassin:2021}
-{Fassin, Y.} {Research on Covid-19: a disruptive phenomenon for bibliometrics}, \emph{Scientometrics}, \textbf{126}, 5305--5319, doi:10.1007/s11192-021-03989-w \newblock (2021).
-
-\end{thebibliography}
-
-\newpage
-
-\begin{center}
-\includegraphics[width=0.95\linewidth]{Casey_Figure_1} %pre-prints-segmented-by-field}
-\end{center}
-
-\noindent \textbf{Figure 1:} \textbf{There has been no significant change in monthly \arxiv\ pre-prints (black) since the COVID-19 pandemic began.} The expected number of pre-prints from our Gaussian Process model is shown in red, and the red region shows the uncertainty in that expectation.
-The dashed light grey line shows predictions from an autoregressive integrated moving average model.
-
-\newpage
-
-\begin{center}
- \includegraphics[width=\linewidth]{Casey_Figure_2} %pandemic-related-preprints.pdf}
-\end{center}
-
-\noindent \textbf{Figure 2: Pre-prints related to the COVID-19 pandemic peaked in April 2020, accounting for nearly 60\% the quantitative biology (\texttt{q-bio}) pre-prints in that month.} Shown are five other research fields with the highest number of pandemic-related pre-prints in 2020.
-Monthly pre-prints are shown on the left, and the percent of pandemic-related pre-prints in that research field is shown on the right.
-
-\newpage
-
-\begin{center}
-	\includegraphics[width=0.95\linewidth]{Casey_Figure_3} %new-authors-segmented-by-field-combined
-\end{center}
-
-\noindent \textbf{Figure 3:} \textbf{The number of new authors appearing in \arxiv\ pre-prints per month (black) peaked in quantitative biology in 2020.} Here we show new authors since 2010, but we used data since 2007 to determine if an author was new.
-The slow change in new authors  approximates the net number of new authors joining the field.
-The number of new authors appearing per number of pre-prints is shown in light blue, with identical limits on the right-hand y-axis.
-This highlights authorship differences between fields (e.g., \texttt{hep-ex} and \texttt{hep-lat}) for large experiments.
-
-% Some notable peaks:
-% gr-qc 2013: https://arxiv.org/abs/1304.0670, GW
-
-\newpage
+The methodology and results presented here provide a quantitative framework for future investigations of the longer term impacts of the pandemic.
 
@@
 Indeed, we re-produced the analysis and figures in this paper by taking all initials given and it made no change to our conclusions.
-This is due, in part, because most authors do not provide middle initials.
+This is due, in part, to most authors not providing middle initials.
 
@@
 We identified a pre-print as being related to the COVID-19 pandemic if the title or abstract contained any of the following four (case-insensitive) terms: `pandemic', `COVID', `SARS-CoV-2', `lockdown', or `coronavirus'.
-Before 2020, these terms rarely appear in the title or abstract of any pre-print on \arxiv\ (Figure~2).
+Before 2020, these terms rarely appear in the title or abstract of any pre-print on \arxiv\ (Figure~\ref{fig:2}).
 
-\section*{Data availability}
+\vspace{\fill}
 
-The entire dataset used in this article is from \arxiv.
+\ethics{No ethics approval was required for this work.}
+
+\dataccess{The entire dataset used in this article is from \arxiv.
 The dataset is curated by Cornell University and hosted by Kaggle (https://www.kaggle.com/Cornell-University/arxiv).
 The dataset is updated weekly.
-We accessed this dataset on 01 June 2021.
+We accessed this dataset on 01 June 2021.}
 
-\section*{Code availability}
-The software developed to retrieve and analyse the data for this project is available online (https://github.com/andycasey/arxiv-covid).
+\aucontribute{All authors contributed to the discussions and have read and iterated upon the text of the final manuscript.}
 
-%----------------------------------------------------------------------------------------
-%	Methods references
-%----------------------------------------------------------------------------------------
+\competing{The authors declare no competing interests.}
 
-% \let\oldthebibliography=\thebibliography
-% \renewenvironment{thebibliography}[1]{%
-%    \oldthebibliography{#1}%
-%    \setcounter{enumiv}{36}%
-% }
+\funding{A.~R.~C. is supported in part by the Australian Research Council through a Discovery Early Career Researcher Award (DE190100656). 
+Parts of this research were supported by the Australian Research Council Centre of Excellence for All Sky Astrophysics in 3 Dimensions (ASTRO 3D), through project number CE170100013.
+I.~M. is a recipient of the Australian Research Council Future Fellowship FT190100574.}
+
+\ack{We thank Peter Skands (Monash University) 
+	 and Ross Young (University of Adelaide) 
+for comments on publication trends in high energy physics.
+% Funding acknowledgements.
+}
+
+\disclaimer{The paper reflects the authors' views and may not reflect those of their employers, funding agencies, or affiliated institutions.}
+
+\pagebreak
+
+%%%%%%%%%% Insert bibliography here %%%%%%%%%%%%%%
 
@@
 
-\setcounter{enumiv}{16}
+\bibitem{Ginsparg:2011}
+{Ginsparg}. P., {ArXiv at 20}, \emph{Nature}, \textbf{476}, 7359, doi:10.1038/476145a \newblock (2011).
+
+\bibitem{Lariviere:2014}
+{Larivi{\'e}re, V., et al.,} {arXiv E-prints and the journal of record: An analysis of roles and relationships}, \emph{Journal of the Association for Information Science and Technology}, \textbf{65}, 6, doi:10.1002/asi.23044 \newblock (2014).
+
+\bibitem{Nicola:2020}
+{Nicola, M., et al.,} {The socio-economic implications of the coronavirus pandemic ({COVID}-19): A review}, \emph{International Journal of Surgery}, \textbf{78}, 185--193, doi:10.1016/j.ijsu.2020.04.018 \newblock (2020).
+
+\bibitem{Chu:2020}
+{Yen-Hao Chu, I., Prima, A., Larson, H.~J., Leesa, L.}
+{Social consequences of mass quarantine during epidemics: a systematic review with implications for the {COVID}-19 response}, \emph{Journal of Travel Medicine}, \textbf{27}, 7, doi:10.1093/jtm/taaa192 \newblock (2020).
+
+\bibitem{IbnMohammed:2021}
+{Ibn-Mohammed, T., et al.,} {A critical analysis of the impacts of {COVID}-19 on the global economy and ecosystems and opportunities for circular economy strategies}, \emph{Resources, Conservation and Recycling}, \textbf{164}, 105169, doi:1016/j.resconrec.2020.105169 \newblock (2021).
+
+\bibitem{Viglione:2020}
+{Viglione, G.}, {Are women publishing less during the pandemic? Here?s what the data say}. \emph{Nature}, \textbf{581}, 356-366, doi:10.1038/d41586-020-01294-9 \newblock (2020).
+
+\bibitem{Gewen:2020},
+{Gewen, V.}, {The career cost of COVID-19 to female researchers, and how science should respond}. \emph{Nature}, \textbf{583}, 867--869, doi:10.1038/d41586-020-02183-x \newblock (2020).
+
+\bibitem{King:2021}
+{King, M.~M., Frederickson M.~E.}, {The Pandemic Penalty: The Gendered Effects of COVID-19 on Scientific Productivity}. \emph{Socius}. January 2021. doi:10.1177/23780231211006977 \newblock (2021).
+
+\bibitem{Clement:2019}
+{Clement, C.~B., Bierbaum, M., O'Keeffe, K.~P., Alemi, A.~A.,}
+{On the Use of ArXiv as a Dataset},
+Preprint at https://arxiv.org/abs/1905.00075 \newblock (2019).
+
+\bibitem{Rasmussen:2006}
+{Rasmussen, C.~E., Williams, C.~K.~I.,}
+{Gaussian Processes for Machine Learning}, \emph{MIT Press}, ISBN 026218253X \newblock (2016).
+
+\bibitem{Ambikasaran:2014}
+{Ambikasaran, S., Foreman-Mackey, D., Greengard, L., Hogg, D.~W., O'Neil, M.,}
+{Fast Direct Methods for Gaussian Processes}, Preprint at http://arxiv.org/abs/1403.6015 \newblock (2014).
+
+\bibitem{Guinnessy:2021}
+Guinnessy, P.
+(Physics Today).
+2021.
+\emph{Physicists gathered virtually for APS March Meeting}.
+[online] Available at {https://physicstoday.scitation.org/do/10.1063/PT.6.3.20210405a/full} [Accessed 17 June 2021].
+
+\bibitem{LatticeConferenceWebsite}
+Helmholtz-Institut f\"ur Strahlen- und Kernphysik Indico Service (Indico).
+2021.
+\emph{The 38th International Symposium on Lattice Field Theory (Lattice 2020)}.
+[online] Available at: {https://indico.hiskp.uni-bonn.de/event/1/} [Accessed 3 June 2021].
+
+\bibitem{Wang:2017}
+{Wang, W., Bai, X., Xia, F.
+et al.}, {From triadic closure to conference closure: the role of academic conferences in promoting scientific collaborations.}, \emph{Scientometrics}, \textbf{113}, 177–193  doi:10.1007/s11192-017-2468-x \newblock (2017).]
+
+\bibitem{PhysicsWorld}
+physicsworld (IOP Publishing).
+2021.
+\emph{Critical research hit as COVID-19 forces physics labs to close}.
+[online] Available at {https://physicsworld.com/a/critical-research-hit-as-covid-19-forces-physics-labs-to-close/} [Accessed 4 June 2021].
+
+\bibitem{ScienceMag}
+Science Magazine (AAAS).
+2021.
+\emph{Amid pandemic, Energy Department labs close to tens of thousands of users.} [online] Available at {https://www.sciencemag.org/news/2020/03/amid-pandemic-energy-department-labs-close-tens-thousands-users} [Accessed 4 June 2021].
+
+\bibitem{Castelvecchi}
+Castelvecchi, D.
+(Nature), {How the coronavirus pandemic is affecting the world's biggest physics experiments.} [online] Available at {https://www.nature.com/articles/d41586-020-00943-3} doi:10.1038/d41586-020-00943-3 [Accessed 6 June 2022].
+
+\bibitem{Fassin:2021}
+{Fassin, Y.} {Research on Covid-19: a disruptive phenomenon for bibliometrics}, \emph{Scientometrics}, \textbf{126}, 5305--5319, doi:10.1007/s11192-021-03989-w \newblock (2021).
 
@@
 
-\section*{Acknowledgements}
-We thank Peter Skands (Monash University) 
-	 and Ross Young (University of Adelaide) 
-for comments on publication trends in high energy physics.
-% Funding acknowledgements.
-A.~R.~C. is supported in part by the Australian Research Council through a Discovery Early Career Researcher Award (DE190100656). 
-Parts of this research were supported by the Australian Research Council Centre of Excellence for All Sky Astrophysics in 3 Dimensions (ASTRO 3D), through project number CE170100013.
-I.~M. is a recipient of the Australian Research Council Future Fellowship FT190100574.
-
-\section*{Author contributions}
-
-All authors contributed to the discussions and have read and iterated upon the text of the final manuscript.
-
-\section*{Competing interests}
-The authors declare no competing interests.
-
-\section*{Materials \& Correspondence}
-Correspondence and requests for materials should be addressed to the corresponding author Andrew Casey (andrew.casey@monash.edu).
-
-\section*{Additional information}
-
-%\textbf{Supplementary Information} is available for this paper.
-
-\noindent\textbf{Correspondence and requests for materials} should be addressed to A.~R.~C.
-
-\noindent\textbf{Reprints and permissions information} is available at {\tt\small www.nature.com/reprints}
-
 \end{document}

### 2022-06-06 article/ms.tex
minor

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage[normalem]{ulem}
+\newcommand{\del}[1]{\textcolor{red}{\sout{#1}}}
+\newcommand{\add}[1]{\textcolor{blue}{#1}}
 
@@
 Here we show that academics are posting more papers on the arXiv pre-print server than would be expected had the pandemic not occurred: 168,630 were posted in 2020, a +12.6\% change from 2019 and $+1.4\sigma$ deviation above the 162,577 predicted by our Gaussian process model.
-Our results also indicate that most individual fields were not significantly affected by the pandemic, however we do observe large declines in output in areas of high-energy physics where laboratory closures and the cancellation of a key conference have played important roles.
-The most significant increase occurred in quantitative biology where a 50\%~increase ($+8\sigma$) in pre-prints was driven by research related to the COVID-19 pandemic.
+We find that that most individual fields were not measurably affected by the pandemic.
+However, we do observe \add{a significant and ongoing decline in physics pre-prints, as well as temporary declines}
+\del{large declines in output} in areas of high-energy physics, where laboratory closures and the cancellation of a key conference have played important roles. 
+The most significant increase occurred \add{during 2020} in quantitative biology where a 50\%~increase ($+8\sigma$) in pre-prints was driven by research related to the COVID-19 pandemic.
 Some of these `COVID papers' are by biologists using arXiv for the first time, but many are written by researchers from other fields (e.g., physicists, mathematicians). 
-As of mid-2021, quantitative biology pre-prints have returned to pre-pandemic levels, but 20\% of the research in this field is now related to the COVID-19 pandemic, demonstrating a strong shift in research focus.
+\del{As of mid-2021, quantitative biology pre-prints have returned to pre-pandemic levels, but 20\% of the research in this field is now related to the COVID-19 pandemic, demonstrating a strong shift in research focus.}
 \end{abstract}
@@
 
-We retrieved metadata for 1,475,914 pre-prints posted to \arxiv\ between 1 April 2007 and 31 May 2021.
+We retrieved metadata for 1,660,423 pre-prints posted to \arxiv\ between 1 April 2007 and 31 May 2022.
 The metadata includes the creation date, research field(s), title, author name(s), abstract, and other miscellaneous information\cite{Clement:2019}.
@@
 
-Border closures and travel restrictions have forced many academic conferences to be held online-only, or to be cancelled.
+Border closures and travel restrictions forced many academic conferences to be held online-only, or to be cancelled.
 Hosting conferences online creates a mix of effects, as forgoing in-person experiences benefits those who could not otherwise participate due to visa issues or travel costs\cite{Guinnessy:2021}.
@@
 In 2020 the conference was cancelled\cite{LatticeConferenceWebsite}, and no accompanying pre-prints exist.
-The impact of conference cancellations and travel restrictions on academic research is likely to extend well beyond the output of a single field in a particular year since discussions at conferences spark new research ideas and collaborations that can lead to publications many months or years later\cite{Wang:2017}.
+\add{The conference was held in 2021, but t}\del{T}he impact of conference cancellations and travel restrictions on academic research is likely to extend well beyond the output of a single field in a particular year, since discussions at conferences spark new research ideas and collaborations that can lead to publications many months or years later\cite{Wang:2017}.
 
@@
 The pandemic has forced many laboratories to close or operate with restricted access\cite{PhysicsWorld,ScienceMag,Castelvecchi}, which could lead to long-lasting delays in ongoing experiments or immediate drops in publication rates in part due to difficulty in accessing completed (or nearly completed) experiment data.
-There is some evidence of this in pre-print numbers already, with {20\%} fewer pre-prints in high energy physics (\texttt{hep-ex}) in 2020 than expected by our model (observed 615; predicted 818 $\pm$ 94).
-But declines are not ubiquitous across experimental fields.
+\del{There is some evidence of this in pre-print numbers, with {20\%} fewer pre-prints in high energy physics (\texttt{hep-ex}) in 2020 than expected by our model (observed 615; predicted 818 $\pm$ 94)}
+\add{The number of pre-prints posted to \arxiv\ in the \texttt{physics} category spiked early in 2020, but the total number of physics pre-prints posted that year was only $1\sigma$ higher than the prediction from our model, because this spike was counter-balanced by a decline in the second half of 2020.
+This decline has steadily continued throughout 2021 and 2022, with physics showing the most significant decline in pre-prints since the pandemic began.}
+But declines are not ubiquitous across all experimental fields.
 The closest comparable field of phenomenology in high energy physics (\texttt{hep-ph}) showed no substantial decline.
@@
 The field of quantitative biology (\texttt{q-bio}) showed the largest increase in pre-prints in 2020 above what is expected by our model.
-There were {2,790} quantitative biology pre-prints in 2020, {50\%} above the previous year, representing a $+8\sigma$ deviation from our model predictions (predicted 1,914 $\pm$ 109, 876 fewer than observed).
+There were {2,790} quantitative biology pre-prints in 2020, {50\%} above the previous year, representing a $+8\sigma$ deviation from our model predictions (predicted 1,914 $\pm$ 109; 876 fewer than observed).
 This increase is explainable by an increase in COVID-19 related research, as there were 864 quantitative biology pre-prints in 2020 with pandemic-related terms in their title or abstract (see Methods), and just 39 in the decade prior.
@@
 As of mid-2021 the number of monthly pre-prints in quantitative biology has dropped to pre-pandemic levels.
-Approximately 20\% of current pre-prints in quantitative biology are related to the ongoing pandemic, representing a strong shift in ongoing research focus.
+Approximately 10--15\% of current pre-prints in quantitative biology are related to the ongoing pandemic.\del{, representing a strong shift in ongoing research focus.} \add{This represents a strong shift in research focus, but this fraction is declining with time.}
 
@@
 While citations are the most used measure of impact of academic publications, that metric becomes a more biased statistic when many related pre-prints are all being posted nearly at the same time\cite{Fassin:2021}.
-Amongst individual fields the largest decline is explainable by a cancelled conference, and the few other drops in pre-prints are statistically insignificant.
-It's plausible that many of the pre-prints already posted are research projects that were well underway before the pandemic.
-The relatively long timescales of academic research would suggest that the full impact on academic productivity due to the COVID-19 pandemic is yet to be seen.
+\add{Amongst individual fields the largest \emph{relative} decline in a single field is explanable by a high energy lattice physics conference that was cancelled in 2020.}
+\add{The largest persistent decline in pre-prints has occurred in physics, a field where many researchers rely on unfettered laboratory access to conduct careful experiments.
+As of mid-2022, this steady decline in physics pre-prints continues.}
+\del{Amongst individual fields the largest fractional decline is explainable by a cancelled conference, and the few other drops in pre-prints are statistically insignificant.}
+\del{It's plausible that many of the pre-prints already posted are research projects that were well underway before the pandemic.} The relatively long timescales of academic research would suggest that the full impact on academic productivity due to the COVID-19 pandemic is yet to be seen.
 The methodology and results presented here provide a quantitative framework for future investigations of the longer term impacts of the pandemic.
@@
 
-\dataccess{The entire dataset used in this article is from \arxiv.
-The dataset is curated by Cornell University and hosted by Kaggle (https://www.kaggle.com/Cornell-University/arxiv).
+\dataccess{\add{All of the data and software used in this work are publicly available.
+The dataset used here is from \arxiv.
+The original dataset is curated by Cornell University and hosted by Kaggle (https://www.kaggle.com/Cornell-University/arxiv).
 The dataset is updated weekly.
-We accessed this dataset on 01 June 2021.}
+We accessed this dataset on 06 June 2022.
+We have stored a time-stamped version of this dataset, as well as the code used to process the data, as Zenodo record \cite{ZenodoRecord}.
+A public GitHub repository (https://github.com/andycasey/arxiv) stores all software used to preprocess and analyse these data, including the scripts to produce the figures shown here, and to compile this manuscript.
+The repository includes a \texttt{README.md} file that describes the software environment used to perform the analysis, and steps on how to reproduce our work in full.}}
 
-\aucontribute{All authors contributed to the discussions and have read and iterated upon the text of the final manuscript.}
+\aucontribute{\add{AC and IM conceived the idea to examine changes in academic productivity due to the pandemic.
+PK and IM helped guide the investigation and provided ideas on new avenues to explore and tests to perform.
+AC conducted the analysis.
+AC led the writing of the manuscript, with all authors contributing significantly.} All authors read and iterated upon the text of the final manuscript.}
 
@@
 
-\bibitem{Gewen:2020},
+\bibitem{Gewen:2020}
 {Gewen, V.}, {The career cost of COVID-19 to female researchers, and how science should respond}. \emph{Nature}, \textbf{583}, 867--869, doi:10.1038/d41586-020-02183-x \newblock (2020).

### 2022-06-20 article/ms.tex
penultimate?

@@
 \newcommand{\add}[1]{\textcolor{blue}{#1}}
+\newcommand{\todo}[1]{\textcolor{green}{#1}}
+
+%\newcommand{\del}[1]{}
+%\newcommand{\add}[1]{#1}
 
@@
 With a global pandemic that has changed the world, how has it changed academic productivity? 
-Here we show that academics are posting more papers on the arXiv pre-print server than would be expected had the pandemic not occurred: 168,630 were posted in 2020, a +12.6\% change from 2019 and $+1.4\sigma$ deviation above the 162,577 predicted by our Gaussian process model.
+Here we show that academics are posting more papers on the arXiv pre-print server than would be expected had the pandemic not occurred: 341,474 pre-prints were posted between 2020-2021, up 14\% from 2019 and above the predictions of our Gaussian process model.
+%168,630 were posted in 2020 , a +12.6\% change from 2019 and $+1.4\sigma$ deviation above the 162,577 predicted by our Gaussian process model. % todo: reword
 We find that that most individual fields were not measurably affected by the pandemic.
-However, we do observe \add{a significant and ongoing decline in physics pre-prints, as well as temporary declines}
-\del{large declines in output} in areas of high-energy physics, where laboratory closures and the cancellation of a key conference have played important roles. 
-The most significant increase occurred \add{during 2020} in quantitative biology where a 50\%~increase ($+8\sigma$) in pre-prints was driven by research related to the COVID-19 pandemic.
+However, we do observe a significant ongoing decline in physics pre-prints, as well as temporary declines in areas of high-energy physics, where the cancellation of a key conference played a key role.
+The most significant increase occurred in quantitative biology where a 50\%~increase ($+8\sigma$) in pre-prints posted in 2020 was driven by research related to the COVID-19 pandemic. 
 Some of these `COVID papers' are by biologists using arXiv for the first time, but many are written by researchers from other fields (e.g., physicists, mathematicians). 
-\del{As of mid-2021, quantitative biology pre-prints have returned to pre-pandemic levels, but 20\% of the research in this field is now related to the COVID-19 pandemic, demonstrating a strong shift in research focus.}
 \end{abstract}
@@
 These long-term trends are relatively predictable from year to year, allowing us to quantify any change in academic productivity due to the COVID-19 pandemic.
-We used the number of publications from January 2015 to December 2019 to train a predictive Gaussian process model\cite{Rasmussen:2006}.
+\add{The number of pre-prints posted varies substantially by category.
+Computer science (\texttt{cs}) is the growing the fastest with more than 5,000 pre-prints per month, followed by math (\texttt{math}) with about 3,500 per month.
+For contrast, only $\approx$200 pre-prints are posted each month quantitative biology (\texttt{q-bio}), and just $\approx$40 in nuclear experimental physics (\texttt{nucl-ex}).
+For this reason, we model the number of monthly pre-prints in each research field separately.} We used the number of publications from January 2015 to December 2019 to train a predictive Gaussian process model\cite{Rasmussen:2006}.
 This model treats the number of monthly pre-prints in a research field as a sequence of normally distributed random variables.
 The expected number of pre-prints was assumed to increase linearly with time, and a quasi-periodic kernel function\cite{Rasmussen:2006,Ambikasaran:2014} was used to capture seasonal variations (Figure~\ref{fig:1}).
-In nearly all research fields, the number of pre-prints posted since January 2020 agree closely with the model predictions (Table~1) indicating no meaningful collective impact on academic productivity due to the COVID-19 pandemic.
-In fact, the total number of \arxiv\ pre-prints in 2020 across all fields (168,630) exceeds our model predictions ($162{,}577 \pm 4{,}393$; $+1.4\sigma$).
+In nearly all research fields, the number of pre-prints posted since January 2020 agree closely with the model predictions (Table~1) indicating no significant overall decline in academic productivity due to the COVID-19 pandemic.
+%In fact, the total number of \arxiv\ pre-prints in 2020 across all fields (168,630) exceeds our model predictions ($162{,}577 \pm 4{,}393$; $+1.4\sigma$).
+% todo
+
+\add{The exception is physics (\texttt{physics}), which shows the largest decline in pre-prints by number since the pandemic began.
+In 2020 the number posted to physics actually exceeded our model predictions (15,242 compared to 14,746) by $1\sigma$, driven by a short peak of pre-prints posted to Physics and Society (\texttt{physics.soc-ph}) in early 2020.
+When we segment the number of pre-prints posted to physics by their sub-category we find that Computational Physics (\texttt{physics.comp-ph}) and Applied Physics (\texttt{physics.app-ph}) had significant drops in pre-print counts since the pandemic began.
+All other physics sub-categories remained largely stable with time.
+The category of mathetmatical physics (\texttt{math-ph}) also showed a decline in pre-prints since the pandemic began, but this category has about an order of magnitude fewer pre-prints posted than physics.
+The mathematical physics category does not have defined sub-categories, making it more difficult to evaluate whether this decline occurred in particular areas of study}.
+
+Experimental research projects often require specialised laboratories, or data to be collected over many years.
+The pandemic has forced many laboratories to close or operate with restricted access\cite{PhysicsWorld,ScienceMag,Castelvecchi}, which could lead to long-lasting delays in ongoing experiments or immediate drops in publication rates in part due to difficulty in accessing completed (or nearly completed) experiment data.
+\add{This could explain why there are declines in pre-prints in some sub-categories (e.g., Applied Physics), but there is not significant evidence of this in pre-print numbers of experimental-only fields (e.g., \texttt{hep-ex} or \texttt{nucl-ex}).
+In those fields the model predictions agree reasonably well with the pre-prints since 2020.}
 
@@
 The expected number of pre-prints from our Gaussian Process model is shown in red, and the red region shows the uncertainty in that expectation.
-The dashed light grey line shows predictions from an autoregressive integrated moving average model.}
+The blue line shows predictions from an autoregressive integrated moving average model.}
 \label{fig:1}
@@
 Hosting conferences online creates a mix of effects, as forgoing in-person experiences benefits those who could not otherwise participate due to visa issues or travel costs\cite{Guinnessy:2021}.
-The immediate impact of cancelling a conference is readily apparent in pre-print counts in lattice physics (Figure~\ref{fig:1}; \arxiv\ subject code \texttt{hep-lat}, see Methods for explanation), where most pre-prints are posted around December each year as conference proceedings from the International Symposium on Lattice Field Theory.
+The immediate impact of cancelling a conference is readily apparent in pre-print counts in lattice physics (Figure~\ref{fig:1}; \arxiv\ subject code \texttt{hep-lat}, see Methods), where most pre-prints are posted around December each year as conference proceedings from the International Symposium on Lattice Field Theory.
 In 2020 the conference was cancelled\cite{LatticeConferenceWebsite}, and no accompanying pre-prints exist.
-\add{The conference was held in 2021, but t}\del{T}he impact of conference cancellations and travel restrictions on academic research is likely to extend well beyond the output of a single field in a particular year, since discussions at conferences spark new research ideas and collaborations that can lead to publications many months or years later\cite{Wang:2017}.
+The conference was held in 2021 with nearly twice the number of corresponding authors than the 2019 conference\cite{LatticeConferenceWebsite2019,LatticeConferenceWebsite2021} (634 compared to 303), but the impact of conference cancellations and travel restrictions on academic research is likely to extend well beyond the output of a single field in a particular year, since discussions at conferences spark new research ideas and collaborations that can lead to publications many months or years later\cite{Wang:2017}.
 
-Experimental research projects often require specialised laboratories, or data to be collected over many years.
-The pandemic has forced many laboratories to close or operate with restricted access\cite{PhysicsWorld,ScienceMag,Castelvecchi}, which could lead to long-lasting delays in ongoing experiments or immediate drops in publication rates in part due to difficulty in accessing completed (or nearly completed) experiment data.
-\del{There is some evidence of this in pre-print numbers, with {20\%} fewer pre-prints in high energy physics (\texttt{hep-ex}) in 2020 than expected by our model (observed 615; predicted 818 $\pm$ 94)}
-\add{The number of pre-prints posted to \arxiv\ in the \texttt{physics} category spiked early in 2020, but the total number of physics pre-prints posted that year was only $1\sigma$ higher than the prediction from our model, because this spike was counter-balanced by a decline in the second half of 2020.
+%Throughout 2020, research in condensed matter (\texttt{cond-mat}) saw a $1.8\sigma$ increase above our model predictions (observed 16,188; predicted 15,325 $\pm$ 492).
+% todo: update for context of 2021
+%Segmenting condensed matter pre-prints by research topic shows that most of this increase (in \texttt{cond-mat}) was driven by a 30\% increase in material science pre-prints (\texttt{cond-mat.mtrl-sci}).
+% todo:
+
+%del{There is some evidence of this in pre-print numbers, with {20\%} fewer pre-prints in high energy physics (\texttt{hep-ex}) in 2020 than expected by our model (observed 615; predicted 818 $\pm$ 94)}
+%\add{The number of pre-prints posted to \arxiv\ in the \texttt{physics} category spiked early in 2020, but the total number of physics pre-prints posted that year was only $1\sigma$ higher than the prediction from our model, because this spike was counter-balanced by a decline in the second half of 2020.
 This decline has steadily continued throughout 2021 and 2022, with physics showing the most significant decline in pre-prints since the pandemic began.}
-But declines are not ubiquitous across all experimental fields.
+
+%But declines are not ubiquitous across all experimental fields.
 The closest comparable field of phenomenology in high energy physics (\texttt{hep-ph}) showed no substantial decline.
 Neither did nuclear experimental research (\texttt{nucl-ex}).
-And throughout 2020, research in condensed matter (\texttt{cond-mat}) saw a $1.8\sigma$ increase above our model predictions (observed 16,188; predicted 15,325 $\pm$ 492).
-Segmenting condensed matter pre-prints by research topic shows that most of this increase (in \texttt{cond-mat}) was driven by a 30\% increase in material science pre-prints (\texttt{cond-mat.mtrl-sci}).
 
-The field of quantitative biology (\texttt{q-bio}) showed the largest increase in pre-prints in 2020 above what is expected by our model.
+\begin{figure}[!t]
+\includegraphics[width=0.95\linewidth]{physics-pre-prints}
+\caption{\add{The number of pre-prints posted to \arxiv\ in the categories of Computational Physics (\texttt{physics.comp-ph}) and Applied Physics (\texttt{physics.app-ph}).
+The decline in physics pre-prints since the pandemic began (Figure~\ref{fig:1}) is largely explained by the drop in pre-prints in Computational Physics and Applied Physics.}}
+\label{fig:physics}
+\end{figure}
+
+The field of quantitative biology (\texttt{q-bio}) showed the largest increase in pre-prints in 2020 above what is expected by our model, before returning to pre-pandemic levels in mid-2021. % todo: 
 There were {2,790} quantitative biology pre-prints in 2020, {50\%} above the previous year, representing a $+8\sigma$ deviation from our model predictions (predicted 1,914 $\pm$ 109; 876 fewer than observed).
@@
 Pandemic-related pre-prints also appeared in other fields (computer science, physics, statistics, and quantitative finance, all peaking around {April 2020}), but these papers constituted fewer than 10\% of the total pre-prints in these fields.
-As of mid-2021 the number of monthly pre-prints in quantitative biology has dropped to pre-pandemic levels.
-Approximately 10--15\% of current pre-prints in quantitative biology are related to the ongoing pandemic.\del{, representing a strong shift in ongoing research focus.} \add{This represents a strong shift in research focus, but this fraction is declining with time.}
+Approximately 10--15\% of current pre-prints in quantitative biology are related to the ongoing pandemic.This represents a strong shift in research focus, but this fraction is declining with time.
+% todo:
 
@@
 
+\newcommand{\totalcell}[1]{\cellcolor{gray!25}\textbf{#1}}
+
 \begin{table}
-       \caption{The number of \arxiv\ pre-prints per primary research field in 2020 compared to 2019, and model predictions (see Methods).}
+       \caption{The number of \arxiv\ pre-prints per primary research field in 2020 and 2021, the percent change relative to 2019, predicted number of pre-prints using pre-pandemic data (see Methods), and the statistical significance between the predicted number and actual number of pre-prints.}
 \begin{center}
@@
     \label{tab:table1}
-    \begin{tabular}{|c|l|c|c|c|c|} 
+    \begin{tabular}{|c|cccc|cccc|} 
     \hline
-    	\arxiv\ & Field of research & Pre-prints & Yearly & Predicted & Significance \\
-	code &&(2020)& change (\%)&(2020)& ($\sigma$)\\
+\arxiv\ & \multicolumn{4}{c|}{2020}  & \multicolumn{4}{c|}{2021} \\
+primary & Actual & Change & Predicted & Sig. & Actual & Change & Predicted & Sig.  \\
+category	& & (\%) &  &  $(\sigma)$ & & (\%) &  & $(\sigma)$ \\
+%code &  &            & change (\%)&       & ($\sigma$)   &            & change (\%) &      & ($\sigma$) \\
 \hline
-\texttt{astro-ph}    & Astrophysics           & 14,893    & $ +3.4$ & 14,990 $\pm$ 653      & $-0.1$ \\
-\texttt{cond-mat}    & Condensed Matter       & 16,188    & $ +7.0$ & 15,325 $\pm$ 492      & $+1.8$ \\
-\texttt{cs      }    & Computer Science       & 54,808    & $+24.3$ & 51,673 $\pm$ 4,047    & $+0.8$ \\
-\texttt{gr-qc   }    & General Relativity and & 3,405     & $+12.4$ & 3,166 $\pm$ 154       & $+1.6$ \\
-					 & Quantum Cosmology      &&&&\\
-\texttt{hep-ex  }    & High Energy Physics    & 615       & $-19.8$ & 818 $\pm$ 94          & $-2.2$ \\
-                     & (Experiment)           &&&&\\
-\texttt{hep-lat }    & High Energy Physics    & 394       & $-19.3$ & 475 $\pm$ 170         & $-0.5$ \\
-                     & (Lattice)              &&&&\\
-\texttt{hep-ph  }    & High Energy Physics    & 4,586     & $ +1.3$ & 4,615 $\pm$ 329       & $-0.1$ \\
-                     & (Phenomenology)        &&&&\\
-\texttt{hep-th  }    & High Energy Physics    & 3,869     & $ +0.9$ & 4,000 $\pm$ 291       & $-0.4$ \\
-                     & (Theory)               &&&&\\
-\texttt{math    }    & Mathematics            & 35,367    & $ +5.5$ & 35,362 $\pm$ 1,177    & $+0.0$ \\
-\texttt{math-ph }    & Mathematical Physics   & 1,378     & $ -6.3$ & 1,460 $\pm$ 95        & $-0.9$ \\
-\texttt{nlin    }    & Nonlinear Sciences     & 885       & $ +6.1$ & 822 $\pm$ 63          & $+1.0$ \\
-\texttt{nucl-ex }    & Nuclear Experiment     & 499       & $ +2.9$ & 490 $\pm$ 61          & $+0.1$ \\
-\texttt{nucl-th }    & Nuclear Theory         & 1,212     & $ -3.2$ & 1,234 $\pm$ 111       & $-0.2$ \\
-\texttt{physics }    & Physics                & 15,241    & $+14.3$ & 14,746 $\pm$ 506      & $+1.0$ \\
-\texttt{q-bio   }    & Quantitative Biology   & 2,790     & $+49.0$ & 1,914 $\pm$ 109       & $+8.0$ \\
-\texttt{q-fin   }    & Quantitative Finance   & 979       & $+23.1$ & 805 $\pm$ 55          & $+3.2$ \\
-\texttt{quant-ph}    & Quantum Physics        & 6,341     & $+13.4$ & 5,858 $\pm$ 266       & $+1.8$ \\
-\texttt{stat    }    & Statistics             & 5,180     & $+20.2$ & 4,824 $\pm$ 490       & $+0.7$ \\
+astro-ph &  14,894 &  +3.4 &  14,990       & -0.1 &  14,388 &  -0.2 &  15,544        & -1.4 \\
+%&         &       &  $\pm$   653 &      &         &       &   $\pm$   857 &      \\                
+cond-mat &  16,187 &  +7.0 &  15,325       & +1.8 &  16,230 &  +7.3 &  15,561        & +1.0 \\
+%&         &       &  $\pm$   492 &      &         &       &   $\pm$   644 &      \\                
+cs &  54,802 & +24.3 &  51,669       & +0.8 &  61,396 & +39.3 &  59,433        & +0.3 \\
+%&         &       &  $\pm$ 4,047 &      &         &       &   $\pm$ 5,698 &      \\                        
+gr-qc &   3,404 & +12.3 &   3,166       & +1.5 &   3,521 & +16.2 &   3,316        & +1.0 \\
+%&         &       &  $\pm$   154 &      &         &       &   $\pm$   201 &      \\                    
+hep-ex &     615 & -19.8 &     818       & -2.2 &     828 &  +8.0 &     813        & +0.1 \\
+%&         &       &  $\pm$    94 &      &         &       &   $\pm$   123 &      \\                    
+hep-lat &     394 & -19.3 &     475       & -0.5 &     690 & +41.4 &     452        & +1.1 \\
+%&         &       &  $\pm$   170 &      &         &       &   $\pm$   222 &      \\                
+hep-ph &   4,585 &  +1.2 &   4,615       & -0.1 &   5,045 & +11.4 &   4,650        & +0.9 \\
+%&         &       &  $\pm$   329 &      &         &       &   $\pm$   428 &      \\                    
+hep-th &   3,870 &  +1.0 &   4,000       & -0.4 &   3,786 &  -1.2 &   4,163        & -1.0 \\
+%&         &       &  $\pm$   291 &      &         &       &   $\pm$   379 &      \\                    
+math &  35,367 &  +5.5 &  35,362       & +0.0 &  34,406 &  +2.6 &  37,091        & -1.6 \\
+%&         &       &  $\pm$ 1,177 &      &         &       &   $\pm$ 1,640 &      \\                    
+math-ph &   1,378 &  -6.3 &   1,460       & -0.9 &   1,117 & -24.0 &   1,468        & -2.8 \\
+%&         &       &  $\pm$    95 &      &         &       &   $\pm$   124 &      \\                
+nlin &     885 &  +6.1 &     822       & +1.0 &     787 &  -5.6 &     831        & -0.5 \\
+%&         &       &  $\pm$    63 &      &         &       &   $\pm$    82 &      \\                    
+nucl-ex &     499 &  +2.9 &     490       & +0.1 &     414 & -14.6 &     499        & -1.1 \\
+%&         &       &  $\pm$    61 &      &         &       &   $\pm$    79 &      \\                
+nucl-th &   1,212 &  -3.2 &   1,234       & -0.2 &   1,128 &  -9.9 &   1,223        & -0.7 \\
+%&         &       &  $\pm$   111 &      &         &       &   $\pm$   144 &      \\                
+physics &  15,242 & +14.3 &  14,746       & +1.0 &  13,917 &  +4.3 &  16,070        & -2.8 \\
+%&         &       &  $\pm$   506 &      &         &       &   $\pm$   776 &      \\                
+q-bio &   2,790 & +49.0 &   1,914       & +8.0 &   2,092 & +11.8 &   1,984        & +0.8 \\
+%&         &       &  $\pm$   109 &      &         &       &   $\pm$   142 &      \\                    
+q-fin &     979 & +23.1 &     805       & +3.2 &     850 &  +6.9 &     837        & +0.2 \\
+%&         &       &  $\pm$    55 &      &         &       &   $\pm$    71 &      \\                    
+quant-ph &   6,342 & +13.5 &   5,859       & +1.8 &   7,378 & +32.0 &   6,170        & +3.5 \\
+%&         &       &  $\pm$   266 &      &         &       &   $\pm$   347 &      \\                
+stat &   5,179 & +20.2 &   4,824       & +0.7 &   4,877 & +13.2 &   5,116        & -0.4 \\
+%&         &       &  $\pm$   490 &      &         &       &   $\pm$   640 &      \\                    
 \hline     
-&\cellcolor{gray!25}\textbf{Total} & \cellcolor{gray!25}$\mathbf{168{,}630}$ & \cellcolor{gray!25}$\mathbf{+12.6}$ & \cellcolor{gray!25}$\mathbf{162{,}577 \pm 4{,}393}$ \cellcolor{gray!25}& \cellcolor{gray!25}$\mathbf{+1.4}$ \\
+\totalcell{Total} & \totalcell{168,624} & \totalcell{+12.6} & \totalcell{162,574} & \totalcell{+1.4} & \totalcell{172,850} & \totalcell{+15.4} & \totalcell{175,221} & \totalcell{-0.4} \\
+% \pm 4{,}393
+% todo
 \hline 
@@
 The COVID-19 pandemic appears to have had minimal impact on collective research productivity to date.
+\add{However, some significant changes were noted within specific areas of research.
+The largest persistent decline in pre-prints has occurred in (primarily computational and applied) physics.
+As of mid-2022, this steady decline continues.} Amongst individual fields the largest \emph{relative} decline in a single field is explainable by a high energy lattice physics conference that was cancelled in 2020.
 In this study we have only focussed on the quantity of publications as a measure for productivity, and not their quality.
 While citations are the most used measure of impact of academic publications, that metric becomes a more biased statistic when many related pre-prints are all being posted nearly at the same time\cite{Fassin:2021}.
-\add{Amongst individual fields the largest \emph{relative} decline in a single field is explanable by a high energy lattice physics conference that was cancelled in 2020.}
-\add{The largest persistent decline in pre-prints has occurred in physics, a field where many researchers rely on unfettered laboratory access to conduct careful experiments.
-As of mid-2022, this steady decline in physics pre-prints continues.}
-\del{Amongst individual fields the largest fractional decline is explainable by a cancelled conference, and the few other drops in pre-prints are statistically insignificant.}
-\del{It's plausible that many of the pre-prints already posted are research projects that were well underway before the pandemic.} The relatively long timescales of academic research would suggest that the full impact on academic productivity due to the COVID-19 pandemic is yet to be seen.
+The relatively long timescales of academic research would suggest that the full impact on academic productivity due to the COVID-19 pandemic is yet to be seen.
 The methodology and results presented here provide a quantitative framework for future investigations of the longer term impacts of the pandemic.
@@
 In general, this errs on the side of name conflation and over-productivity, except for the relatively rare authors who change their last names over their career.
-Any common name is likely to appear in the literature early in the data set and will not impact the conclusions we draw about how publishing changed in 2020.
+Any common name is likely to appear in the literature early in the data set and will not impact the conclusions we draw about how publishing changed from 2020 January onwards.
 Indeed, we re-produced the analysis and figures in this paper by taking all initials given and it made no change to our conclusions.
@@
 
+\bibitem{LatticeConferenceWebsite2019}
+Helmholtz-Institut f\"ur Strahlen- und Kernphysik Indico Service (Indico).
+2021.
+\emph{The 38th International Symposium on Lattice Field Theory (Lattice 2020)}.
+[online] Available at: {https://indico.cern.ch/event/764552/} [Accessed 20 June 2022].
+
+\bibitem{LatticeConferenceWebsite2021}
+Helmholtz-Institut f\"ur Strahlen- und Kernphysik Indico Service (Indico).
+2021.
+\emph{The 38th International Symposium on Lattice Field Theory (Lattice 2021)}.
+[online] Available at: {https://indico.cern.ch/event/1006302/} [Accessed 20 June 2022].
+
 \bibitem{Wang:2017}

### 2022-06-20 plot-new-authors-segmented-by-field.py

@@ -24,7 +24,7 @@ else:
 bins = np.arange(2007, max(records["created_decimal_year"]) + 1/12, 1/12)
 
 xlim = (bins[0], bins[-1])
-xlim = (2010, bins[-1])
+xlim = (2017, 2023)
 
 
 def get_new_authors(arxiv_ids, authors, callback=None):
@@ -269,7 +269,7 @@ for ppc, y, z in data_new_authors:
         *pad_edges(bins[idx:], y[idx:]),
         drawstyle="steps-mid",
         c="k",
-        alpha=0.5,
+        alpha=0.75,
         zorder=100
     )
     tax.set_xlim(*xlim)

### 2022-06-20 plot-physics-declines.py

@@ -1,0 +1,96 @@
+from http.client import NOT_ACCEPTABLE
+from sysconfig import is_python_build
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.ticker import MaxNLocator
+from tqdm import tqdm
+
+from plot_utils import (load_records, load_authors, load_metadata, get_decimal_date, pad_edges, unique_ify)
+
+matplotlib.style.use("matplotlibrc.paper")
+
+try:
+    (records, authors, metadata)
+except NameError:
+    records = load_records()
+    authors = load_authors()
+    metadata = load_metadata()
+else:
+    print("WARNING: Using pre-loaded data.")
+
+
+xlim = (2014, 2023)
+key = "primary_category"
+is_physics = (records["primary_parent_category"] == "physics")
+
+categories = set(records[key][is_physics])
+N_categories = len(categories)
+
+bin_width = 1/12
+bins = np.arange(2010, max(records["created_decimal_year"]) + bin_width, bin_width)
+t = bins[:-1] + 0.5 * np.diff(bins)[0]
+
+
+fig, ax = plt.subplots()
+H, bin_edges = np.histogram(records["created_decimal_year"][is_physics], bins)
+ax.plot(
+    t,
+    H,
+    drawstyle="steps-mid",
+    c="k"
+)
+
+# show top most N
+show_top_most_N = 22
+
+K = 4
+L = int(show_top_most_N/K)
+
+fig, axes = plt.subplots(L, K, figsize=(10, 10))
+
+#for j, group in enumerate(records[is_physics].group_by([key]).groups):
+from collections import Counter
+for j, (category, count) in enumerate(Counter(records[key][is_physics]).most_common(show_top_most_N)):
+
+    group = records[is_physics * (records[key] == category)]
+    try:
+        ax = axes.flat[j]
+    except IndexError:
+        break
+    sub_cat = category.split(".")[1]
+    ax.set_title(r"$\textrm{{{0}}}$".format(sub_cat))
+
+    ys = group["created_decimal_year"]
+    
+    H, bin_edges = np.histogram(group["created_decimal_year"], bins)
+    
+    # Build a model.
+
+    ax.plot(
+        t,
+        H,
+        drawstyle="steps-mid",
+        c="k"
+    )
+
+for j, ax in enumerate(axes.flat[:show_top_most_N]):
[…]

### 2022-06-20 plot-pre-prints-segmented-by-field.py

@@ -69,7 +69,7 @@ else:
     print("WARNING: Using pre-determined counts")
 
 
-xlim = (2010, 2023)
+xlim = (2017, 2023)
 
 class LineModel(Model):
     parameter_names = ("m", "b")
@@ -86,6 +86,7 @@ num_preprints_var = np.zeros(shape)
 statistics = {}
 
 reports = []
+tabular_data = []
 
 fig, axes = plt.subplots(6, 3, figsize=(6, 8))
 assert len(axes.flat) >= len(ppcs)
@@ -175,7 +176,7 @@ for j, group in enumerate(records.group_by(["primary_parent_category"]).groups):
     )
 
     pred_as_bins = True
-    t_s, t_e = (2020, 2022.5)
+    t_s, t_e = (2020, 2023)
 
     if pred_as_bins:
         t_pred = np.arange(t_s + 0.5 * bin_width, t_e + 0.5 * bin_width, bin_width)
@@ -200,7 +201,7 @@ for j, group in enumerate(records.group_by(["primary_parent_category"]).groups):
             facecolor="tab:red",
             edgecolor=None,
             zorder=0,
-            alpha=1,
+            alpha=0.5,
             lw=0
         )
 
@@ -213,7 +214,8 @@ for j, group in enumerate(records.group_by(["primary_parent_category"]).groups):
         ax.plot(
             t_pred, y_pred,
             c="tab:red",
-            drawstyle="steps-mid",
+            #drawstyle="steps-mid",
+            lw=2,
             zorder=1
         )
 
@@ -221,7 +223,7 @@ for j, group in enumerate(records.group_by(["primary_parent_category"]).groups):
             t_pred,
             y_pred - np.sqrt(y_pred_var),
             y_pred + np.sqrt(y_pred_var),
-            step="mid",
+            #step="mid",
             facecolor="tab:red",
             edgecolor=None,
             zorder=0,
@@ -256,28 +258,43 @@ for j, group in enumerate(records.group_by(["primary_parent_category"]).groups):
         ax.plot(
             t_pred,
             y_pred_arima,
-            c="#666666",
-            lw=1,         
-            ls="-.",  
-            zorder=2
[…]

### 2022-06-21 article/ms.tex
add zenodo link, final comments

@@
 We find that that most individual fields were not measurably affected by the pandemic.
-However, we do observe a significant ongoing decline in physics pre-prints, as well as temporary declines in areas of high-energy physics, where the cancellation of a key conference played a key role.
+However, we do observe a significant ongoing decline in \add{general} physics pre-prints, as well as temporary declines in areas of high-energy physics, where the cancellation of a key conference played a key role.
 The most significant increase occurred in quantitative biology where a 50\%~increase ($+8\sigma$) in pre-prints posted in 2020 was driven by research related to the COVID-19 pandemic. 
@@
 These long-term trends are relatively predictable from year to year, allowing us to quantify any change in academic productivity due to the COVID-19 pandemic.
-\add{The number of pre-prints posted varies substantially by category.
+The number of pre-prints posted varies substantially by category.
 Computer science (\texttt{cs}) is the growing the fastest with more than 5,000 pre-prints per month, followed by math (\texttt{math}) with about 3,500 per month.
 For contrast, only $\approx$200 pre-prints are posted each month quantitative biology (\texttt{q-bio}), and just $\approx$40 in nuclear experimental physics (\texttt{nucl-ex}).
-For this reason, we model the number of monthly pre-prints in each research field separately.} We used the number of publications from January 2015 to December 2019 to train a predictive Gaussian process model\cite{Rasmussen:2006}.
+For this reason, we model the number of monthly pre-prints in each research field separately.
+We used the number of publications from January 2015 to December 2019 to train a predictive Gaussian process model\cite{Rasmussen:2006}.
 This model treats the number of monthly pre-prints in a research field as a sequence of normally distributed random variables.
@@
 
-\add{The exception is physics (\texttt{physics}), which shows the largest decline in pre-prints by number since the pandemic began.
-In 2020 the number posted to physics actually exceeded our model predictions (15,242 compared to 14,746) by $1\sigma$, driven by a short peak of pre-prints posted to Physics and Society (\texttt{physics.soc-ph}) in early 2020.
-When we segment the number of pre-prints posted to physics by their sub-category we find that Computational Physics (\texttt{physics.comp-ph}) and Applied Physics (\texttt{physics.app-ph}) had significant drops in pre-print counts since the pandemic began.
+The exception is \add{general physics (\arxiv\ category \texttt{physics}; not all physics-related domains}), which shows the largest decline in pre-prints by number since the pandemic began.
+In 2020 the number posted to general physics actually exceeded our model predictions (15,242 compared to 14,746) by $1\sigma$, driven by a short peak of pre-prints posted to Physics and Society (\texttt{physics.soc-ph}) in early 2020.
+When we segment the number of pre-prints posted to physics by their sub-category we find that Computational Physics (\texttt{physics.comp-ph}) and Applied Physics (\texttt{physics.app-ph}) had significant drops in pre-print counts since the pandemic began (Figure~\ref{fig:physics}).
 All other physics sub-categories remained largely stable with time.
 The category of mathetmatical physics (\texttt{math-ph}) also showed a decline in pre-prints since the pandemic began, but this category has about an order of magnitude fewer pre-prints posted than physics.
-The mathematical physics category does not have defined sub-categories, making it more difficult to evaluate whether this decline occurred in particular areas of study}.
+The mathematical physics category does not have defined sub-categories, making it more difficult to evaluate whether this decline occurred in particular areas of study.
 
@@
 The pandemic has forced many laboratories to close or operate with restricted access\cite{PhysicsWorld,ScienceMag,Castelvecchi}, which could lead to long-lasting delays in ongoing experiments or immediate drops in publication rates in part due to difficulty in accessing completed (or nearly completed) experiment data.
-\add{This could explain why there are declines in pre-prints in some sub-categories (e.g., Applied Physics), but there is not significant evidence of this in pre-print numbers of experimental-only fields (e.g., \texttt{hep-ex} or \texttt{nucl-ex}).
-In those fields the model predictions agree reasonably well with the pre-prints since 2020.}
+This could explain why there are declines in pre-prints in some sub-categories (e.g., Applied Physics), but there is not significant evidence of this in pre-print numbers of experimental-only fields (e.g., \texttt{hep-ex} or \texttt{nucl-ex}).
+In those fields the model predictions agree reasonably well with the pre-prints since 2020.
 
@@
 \includegraphics[width=0.95\linewidth]{physics-pre-prints}
-\caption{\add{The number of pre-prints posted to \arxiv\ in the categories of Computational Physics (\texttt{physics.comp-ph}) and Applied Physics (\texttt{physics.app-ph}).
-The decline in physics pre-prints since the pandemic began (Figure~\ref{fig:1}) is largely explained by the drop in pre-prints in Computational Physics and Applied Physics.}}
+\caption{The number of pre-prints posted to \arxiv\ in the categories of Computational Physics (\texttt{physics.comp-ph}) and Applied Physics (\texttt{physics.app-ph}).
+The decline in physics pre-prints since the pandemic began (Figure~\ref{fig:1}) is largely explained by the drop in pre-prints in Computational Physics and Applied Physics.}
 \label{fig:physics}
@@
 The COVID-19 pandemic appears to have had minimal impact on collective research productivity to date.
-\add{However, some significant changes were noted within specific areas of research.
-The largest persistent decline in pre-prints has occurred in (primarily computational and applied) physics.
+However, some significant changes were noted within specific areas of research.
+The largest persistent decline in pre-prints has occurred in \add{general physics (arXiv primary category \texttt{physics}), largely due to a decline in  pre-prints posted to computational and applied physics areas within the \texttt{physics} primary category.
 As of mid-2022, this steady decline continues.} Amongst individual fields the largest \emph{relative} decline in a single field is explainable by a high energy lattice physics conference that was cancelled in 2020.
@@
 
-\dataccess{\add{All of the data and software used in this work are publicly available.
+\dataccess{All of the data and software used in this work are publicly available.
 The dataset used here is from \arxiv.
@@
 We accessed this dataset on 06 June 2022.
-We have stored a time-stamped version of this dataset, as well as the code used to process the data, as Zenodo record \cite{ZenodoRecord}.
+We have stored a time-stamped version of this dataset, as well as the code used to process the data \cite{ZenodoRecord}.
 A public GitHub repository (https://github.com/andycasey/arxiv) stores all software used to preprocess and analyse these data, including the scripts to produce the figures shown here, and to compile this manuscript.
-The repository includes a \texttt{README.md} file that describes the software environment used to perform the analysis, and steps on how to reproduce our work in full.}}
+The repository includes a \texttt{README.md} file that describes the software environment used to perform the analysis, and steps on how to reproduce our work in full.}
 
-\aucontribute{\add{AC and IM conceived the idea to examine changes in academic productivity due to the pandemic.
-PK and IM helped guide the investigation and provided ideas on new avenues to explore and tests to perform.
+\aucontribute{AC and IM conceived the idea to examine changes in academic productivity due to the pandemic.
+\add{PR} and IM helped guide the investigation and provided ideas on new avenues to explore and tests to perform.
 AC conducted the analysis.
-AC led the writing of the manuscript, with all authors contributing significantly.} All authors read and iterated upon the text of the final manuscript.}
+AC led the writing of the manuscript, with all authors contributing significantly.
+All authors read and iterated upon the text of the final manuscript.}
 
@@
 \bibitem{Viglione:2020}
-{Viglione, G.}, {Are women publishing less during the pandemic? Here?s what the data say}. \emph{Nature}, \textbf{581}, 356-366, doi:10.1038/d41586-020-01294-9 \newblock (2020).
+{Viglione, G.}, {Are women publishing less during the pandemic? Here\add{'}s what the data say}. \emph{Nature}, \textbf{581}, 356-366, doi:10.1038/d41586-020-01294-9 \newblock (2020).
 
@@
 
+\add{\bibitem{ZenodoRecord} Casey, Andrew R. (2022). Impact of the COVID-19 pandemic on academic productivity [Data set]. Zenodo. https://doi.org/10.5281/zenodo.6615684}
+
 \end{thebibliography}

### 2022-06-22 article/ms.tex
returned to rsos

@@
 We find that that most individual fields were not measurably affected by the pandemic.
-However, we do observe a significant ongoing decline in \add{general} physics pre-prints, as well as temporary declines in areas of high-energy physics, where the cancellation of a key conference played a key role.
+However, we do observe a significant ongoing decline in general physics pre-prints, as well as temporary declines in areas of high-energy physics, where the cancellation of a key conference played a key role.
 The most significant increase occurred in quantitative biology where a 50\%~increase ($+8\sigma$) in pre-prints posted in 2020 was driven by research related to the COVID-19 pandemic. 
@@
 
-The exception is \add{general physics (\arxiv\ category \texttt{physics}; not all physics-related domains}), which shows the largest decline in pre-prints by number since the pandemic began.
+The exception is general physics (\arxiv\ category \texttt{physics}; not all physics-related domains), which shows the largest decline in pre-prints by number since the pandemic began.
 In 2020 the number posted to general physics actually exceeded our model predictions (15,242 compared to 14,746) by $1\sigma$, driven by a short peak of pre-prints posted to Physics and Society (\texttt{physics.soc-ph}) in early 2020.
@@
 However, some significant changes were noted within specific areas of research.
-The largest persistent decline in pre-prints has occurred in \add{general physics (arXiv primary category \texttt{physics}), largely due to a decline in  pre-prints posted to computational and applied physics areas within the \texttt{physics} primary category.
+The largest persistent decline in pre-prints has occurred in {general physics (arXiv primary category \texttt{physics}), largely due to a decline in  pre-prints posted to computational and applied physics areas within the \texttt{physics} primary category.
 As of mid-2022, this steady decline continues.} Amongst individual fields the largest \emph{relative} decline in a single field is explainable by a high energy lattice physics conference that was cancelled in 2020.
@@
 \aucontribute{AC and IM conceived the idea to examine changes in academic productivity due to the pandemic.
-\add{PR} and IM helped guide the investigation and provided ideas on new avenues to explore and tests to perform.
+{PR} and IM helped guide the investigation and provided ideas on new avenues to explore and tests to perform.
 AC conducted the analysis.
@@
 \bibitem{Viglione:2020}
-{Viglione, G.}, {Are women publishing less during the pandemic? Here\add{'}s what the data say}. \emph{Nature}, \textbf{581}, 356-366, doi:10.1038/d41586-020-01294-9 \newblock (2020).
+{Viglione, G.}, {Are women publishing less during the pandemic? Here's what the data say}. \emph{Nature}, \textbf{581}, 356-366, doi:10.1038/d41586-020-01294-9 \newblock (2020).
 
@@
 
-\add{\bibitem{ZenodoRecord} Casey, Andrew R. (2022). Impact of the COVID-19 pandemic on academic productivity [Data set]. Zenodo. https://doi.org/10.5281/zenodo.6615684}
+\bibitem{ZenodoRecord} Casey, Andrew R. (2022). Impact of the COVID-19 pandemic on academic productivity [Data set]. Zenodo. https://doi.org/10.5281/zenodo.6615684
 

### 2022-10-26 article/ms.tex
it's always reviewer 2

@@
 \newcommand{\todo}[1]{\textcolor{green}{#1}}
+\newcommand{\change}[1]{\textbf{#1}}
 
@@
 %%%% Article title to be placed here
-\title{The impact of the COVID-19 pandemic on academic productivity}
+\title{\change{The effect of the COVID-19 pandemic on arXiv pre-prints in maths, physics, computer science, and quantitative biology}}
 
@@
 
-\noindent Peer-reviewed publications are the most common measure of productivity in academia.
+\noindent Peer-reviewed publications are the most common measure of productivity in \change{many specialties of science and technology}.
 The \arxiv\cite{Ginsparg:2011} ({https://arxiv.org}) is a service for the distribution of research publications (pre-prints) prior to their appearance in a journal.
-The presence of a pre-print on \arxiv\ does not ensure that the contents have passed peer-review, but most material on \arxiv\ eventually goes through peer-review because it is now standard in many research fields to post to \arxiv\ either during or after the peer-review process \cite{Lariviere:2014}.
-For this reason, the number of pre-prints posted to \arxiv\ approximates the number of peer-reviewed publications written in these fields.
+The presence of a pre-print on \arxiv\ does not ensure that the contents have passed peer-review, but most material on \arxiv\ eventually goes through peer-review because it is now standard in many research specialties to post to \arxiv\ either during or after the peer-review process \cite{Lariviere:2014}.
+\change{The fraction of publications that are also posted to \arxiv\ increases each year but varies significantly by specialty, from 18.5\% in mathematics to nearly 100\% in high energy physics\cite{?,??}.
+Not all academic outputs (peer-reviewed publications or otherwise) in a given research area will appear on \arxiv, but in the specialties we investigate here (physics, mathematics, and other STEM-related fields) the number of pre-prints posted on \arxiv\ is proportional to (a proxy measure of) the number of peer-reviwed publications in that area.
+For this reason, even if the proportionality differs between specialities of research, we can still examine changes per specialty as a function of time before and during the COVID-19 pandemic}.
 
@@
 One quarter of the excess pre-prints posted to quantitative biology during 2020 were both related to the pandemic and written by authors who had never appeared in the quantitative biology literature (234 of the 876 excess pre-prints in \texttt{q-bio}).
-A careful examination reveals that about half of these (45\%; 107/234) were written by mathematicians or physicists, with another 26\% (61/234) led by engineers, computer scientists, or economists.
+A careful examination \change{(see Methods)} reveals that about half of these (45\%; 107/234) were written by mathematicians or physicists, with another 26\% (61/234) led by engineers, computer scientists, or economists.
 These pre-prints tend to focus on modelling the COVID-19 outbreak using public data sets, rather than quantitative biology research that requires more expert domain knowledge.
@@
 
+\subsection*{Identifying research specialisations of pandemic-related pre-prints posted by new authors in \texttt{q-bio}}
+
+\change{There were 234 pre-prints posted to the \texttt{q-bio} primary category in 2020 which include pandemic-related terms, and where the lead author had never before appeared on the \texttt{q-bio} pre-print before.
+We investigated every pre-print to extract the affiliation of those authors, and used this information to search the internet for the author's website.
+From this information we identified their research specialisations using their own words, or from their social media pages if no website could be found.
+In situations where an author's specialisation crossed multiple fields (e.g., mathematics and biology) we took a conservative approach and described the author as a biologist.
+However, we found that only 3\% (8/234) were led by established biologists who had not appeared in \texttt{q-bio} pre-prints before.
+About half (45\%; 107/234) were written by mathematicians or physicists, with another 26\% (61/234) led by engineers, computer scientists, or economists.}
+
 \vspace{\fill}

### 2022-10-27 article/ms.tex
revviewer comments

@@
 With a global pandemic that has changed the world, how has it changed academic productivity? 
-Here we show that academics are posting more papers on the arXiv pre-print server than would be expected had the pandemic not occurred: 341,474 pre-prints were posted between 2020-2021, up 14\% from 2019 and above the predictions of our Gaussian process model.
+Here we show that academics \change{in science and technology specialisations} are posting more \change{manuscripts} on the arXiv pre-print server than would be expected had the pandemic not occurred: 341,474 pre-prints were posted between 2020-2021, up 14\% from 2019 and above the predictions of our Gaussian process model.
 %168,630 were posted in 2020 , a +12.6\% change from 2019 and $+1.4\sigma$ deviation above the 162,577 predicted by our Gaussian process model. % todo: reword
@@
 \noindent Peer-reviewed publications are the most common measure of productivity in \change{many specialties of science and technology}.
-The \arxiv\cite{Ginsparg:2011} ({https://arxiv.org}) is a service for the distribution of research publications (pre-prints) prior to their appearance in a journal.
+The \arxiv\cite{Ginsparg:2011} ({https://arxiv.org}) is a service for the distribution of research manuscripts (pre-prints) prior to their appearance in a journal.
 The presence of a pre-print on \arxiv\ does not ensure that the contents have passed peer-review, but most material on \arxiv\ eventually goes through peer-review because it is now standard in many research specialties to post to \arxiv\ either during or after the peer-review process \cite{Lariviere:2014}.
-\change{The fraction of publications that are also posted to \arxiv\ increases each year but varies significantly by specialty, from 18.5\% in mathematics to nearly 100\% in high energy physics\cite{?,??}.
-Not all academic outputs (peer-reviewed publications or otherwise) in a given research area will appear on \arxiv, but in the specialties we investigate here (physics, mathematics, and other STEM-related fields) the number of pre-prints posted on \arxiv\ is proportional to (a proxy measure of) the number of peer-reviwed publications in that area.
-For this reason, even if the proportionality differs between specialities of research, we can still examine changes per specialty as a function of time before and during the COVID-19 pandemic}.
+\change{The fraction of peer-reviewed publications that are also posted to the \arxiv\ increases each year\cite{Davis:2007} but not all academic outputs (peer reviewed publications or otherwise) will appear on the \arxiv.
+Fundamental reports, findings from inquiries or research groups, and many other texts that are only published in peer-reviewed publications will never appear on the \arxiv.
+In the specialisations we investigate here (physics, mathematics, astronomy, computer science, and other STEM-related fields), the fraction of peer-reviewed literature that also appears on \arxiv\ is estimated to vary from 18.5\% in mathematics to nearly 100\% in high energy physics\cite{Davis:2007,GentilBeccot:2009}.
+Although the fraction may vary between discplines, within a given specialisation the number of pre-prints provides a proxy measure of the number of peer-reviewed publications, which we can use to examine changes in each specialty as a function of time.}
 
@@
 For contrast, only $\approx$200 pre-prints are posted each month quantitative biology (\texttt{q-bio}), and just $\approx$40 in nuclear experimental physics (\texttt{nucl-ex}).
-For this reason, we model the number of monthly pre-prints in each research field separately.
+For this reason, we model the number of monthly pre-prints in each specialisation separately.
 We used the number of publications from January 2015 to December 2019 to train a predictive Gaussian process model\cite{Rasmussen:2006}.
@@
 All other physics sub-categories remained largely stable with time.
-The category of mathetmatical physics (\texttt{math-ph}) also showed a decline in pre-prints since the pandemic began, but this category has about an order of magnitude fewer pre-prints posted than physics.
+The category of mathematical physics (\texttt{math-ph}) also showed a decline in pre-prints since the pandemic began, but this category has about an order of magnitude fewer pre-prints posted than \change{the general} physics \change{category}.
 The mathematical physics category does not have defined sub-categories, making it more difficult to evaluate whether this decline occurred in particular areas of study.
@@
 Pandemic-related pre-prints also appeared in other fields (computer science, physics, statistics, and quantitative finance, all peaking around {April 2020}), but these papers constituted fewer than 10\% of the total pre-prints in these fields.
-Approximately 10--15\% of current pre-prints in quantitative biology are related to the ongoing pandemic.This represents a strong shift in research focus, but this fraction is declining with time.
+Approximately 10--15\% of current pre-prints in quantitative biology are related to the ongoing pandemic.
+This represents a strong shift in research focus, but this fraction is declining with time.
 % todo:
@@
 
-The COVID-19 pandemic appears to have had minimal impact on collective research productivity to date.
-However, some significant changes were noted within specific areas of research.
+The COVID-19 pandemic appears to have had minimal impact on collective research productivity to date, \change{and that is before accounting for the tremendous work in scientific communication and contributions to public policy throughout the pandemic.} However, some significant changes were noted within specific areas of research.
 The largest persistent decline in pre-prints has occurred in {general physics (arXiv primary category \texttt{physics}), largely due to a decline in  pre-prints posted to computational and applied physics areas within the \texttt{physics} primary category.
 As of mid-2022, this steady decline continues.} Amongst individual fields the largest \emph{relative} decline in a single field is explainable by a high energy lattice physics conference that was cancelled in 2020.
-In this study we have only focussed on the quantity of publications as a measure for productivity, and not their quality.
+In this study we have only focussed on the quantity of \change{pre-prints} as a measure for productivity, and not their quality.
 While citations are the most used measure of impact of academic publications, that metric becomes a more biased statistic when many related pre-prints are all being posted nearly at the same time\cite{Fassin:2021}.
@@
 While our approach to name confusion is grossly simple, it is unlikely that these choices have any substantial impact on our inferences.
-In general, this errs on the side of name conflation and over-productivity, except for the relatively rare authors who change their last names over their career.
+\change{Any author who changes their given or family name (or initials) will incorrectly appear as a `new' author when they next appear on a pre-print, but we expect this fraction to vary smoothly in time, allowing us to examine sharp changes as a consequence of the pandemic.} In general, \change{our approach} errs on the side of name conflation and over-productivity.
 Any common name is likely to appear in the literature early in the data set and will not impact the conclusions we draw about how publishing changed from 2020 January onwards.
@@
 
-We identified a pre-print as being related to the COVID-19 pandemic if the title or abstract contained any of the following four (case-insensitive) terms: `pandemic', `COVID', `SARS-CoV-2', `lockdown', or `coronavirus'.
+We identified a pre-print as being related to the COVID-19 pandemic if the title or abstract contained any of the following four (case-insensitive) terms: `pandemic', `COVID', `SARS-CoV-2', `lockdown', \change{`spike glycoprotein`}, or `coronavirus'.
 Before 2020, these terms rarely appear in the title or abstract of any pre-print on \arxiv\ (Figure~\ref{fig:2}).
@@
 
-\change{There were 234 pre-prints posted to the \texttt{q-bio} primary category in 2020 which include pandemic-related terms, and where the lead author had never before appeared on the \texttt{q-bio} pre-print before.
-We investigated every pre-print to extract the affiliation of those authors, and used this information to search the internet for the author's website.
+\change{There were 234 pre-prints posted to the \texttt{q-bio} primary category in 2020 which include pandemic-related terms, and where no author had never appeared on the \texttt{q-bio} pre-print before.
+We investigated every pre-print to extract the affiliation of the lead author, and used this information to search the internet for the author's website.
 From this information we identified their research specialisations using their own words, or from their social media pages if no website could be found.
@@
 However, we found that only 3\% (8/234) were led by established biologists who had not appeared in \texttt{q-bio} pre-prints before.
-About half (45\%; 107/234) were written by mathematicians or physicists, with another 26\% (61/234) led by engineers, computer scientists, or economists.}
+About half (45\%; 107/234) were led by mathematicians or physicists, with another 26\% (61/234) led by engineers, computer scientists, or economists.}
 
@@
 {Larivi{\'e}re, V., et al.,} {arXiv E-prints and the journal of record: An analysis of roles and relationships}, \emph{Journal of the Association for Information Science and Technology}, \textbf{65}, 6, doi:10.1002/asi.23044 \newblock (2014).
+
+\bibitem{Davis:2007}
+{Davis P.
+M., Fromerth M.
+J.,} {Does the arXiv lead to higher citations and reduced publisher downloads for mathematics articles?}, \emph{Scientometrics}, \textbf{71}, 2, doi:10.1007/s11192-007-1661-8 \newblock (2007).
+
+\bibitem{GentilBeccot:2009}
+{Gentil-Beccot A., Mele S., \& Brooks T.,} {Citing and Reading Behaviours in High-Energy Physics.
+How a Community Stopped Worrying about Journals and Learned to Love Repositories}, Preprint at https://arxiv.org/abs/0906.5418 \newblock (2009).
 

### 2022-10-27 plot-pandemic-related-preprints.py

@@ -44,20 +44,41 @@ bin_max = max(records["created_decimal_year"])
 
 bins = np.arange(2011, bin_max, 1/12)
 counts = {}
+referee_counts = {}
 totals = {}
+referee_totals = {}
 
-patterns = ("pandemic", "covid", "sars-cov-2", "lockdown", "coronavirus")
+original_pattern_arxiv_ids = []
+referee_pattern_arxiv_ids = []
+
+patterns = ("pandemic", "covid", "sars-cov-2", "lockdown", "coronavirus", "spike glycoprotein")
+referee_patterns = [
+    "2019nCoV", "epidemic", "Corona Virus", "CoV 2", "Cov2", "nCoV",
+    "SARS2", "SARSCoV", "severe acurte respiratory syndrome", "spike glycoprotein",
+    "novel betacov"
+]
 
 q_bio_2020_ids = []
+referee_q_bio_2020_ids = []
+referee_matches = []
+
+diff_counts = {}
+why_matched = {}
+why_diff_matched = {}
 
 for ppc in ppcs:
     print(ppc)
 
     counts[ppc] = np.zeros_like(bins)
     totals[ppc] = np.zeros_like(bins)
+    diff_counts[ppc] = np.zeros_like(bins)
+
+    referee_counts[ppc] = np.zeros_like(bins)
+    referee_totals[ppc] = np.zeros_like(bins)
     
     mask = (records["primary_parent_category"] == ppc)
-
+    ppc_original_pattern_arxiv_ids = []
+    ppc_referee_pattern_arxiv_ids = []
     for arxiv_id in tqdm(records["id"][mask]):
         date = get_decimal_date(arxiv_id)
         if date < bins[0] or date > bins[-1]:
@@ -65,20 +86,45 @@ for ppc in ppcs:
 
         idx = np.digitize(date, bins)
         totals[ppc][idx] += 1
+        referee_totals[ppc][idx] += 1
 
         md = metadata[arxiv_id]
         context = f"{md['title']} {md['abstract']}".lower()
         for pattern in patterns:
             if pattern in context:
                 counts[ppc][idx] += 1
+                original_pattern_arxiv_ids.append(arxiv_id)
+                ppc_original_pattern_arxiv_ids.append(arxiv_id)
                 if ppc == "q-bio" and date > 2020 and date < 2021:
                     q_bio_2020_ids.append(arxiv_id)
                 break
+        
[…]

### 2022-10-28 article/ms.tex
minor

@@
 
-The COVID-19 pandemic appears to have had minimal impact on collective research productivity to date, \change{and that is before accounting for the tremendous work in scientific communication and contributions to public policy throughout the pandemic.} However, some significant changes were noted within specific areas of research.
+\change{The COVID-19 pandemic appears to have had minimal impact on collective research productivity {in the science and technology specialisations we examined}, {and that is before accounting for the tremendous work in scientific communication and contributions to public policy throughout the pandemic.} These findings are at odds with similar studies in fields that have larger research overlap with the SARS-CoV-2 virus.
+For example, publications in life sciences saw a 6.5 fold increase in subject headings related to COVID-19\cite{Riccaboni}.
+The effects of the pandemic clearly impact research specialisations differently, and we note some} significant changes within specific research areas.
 The largest persistent decline in pre-prints has occurred in {general physics (arXiv primary category \texttt{physics}), largely due to a decline in  pre-prints posted to computational and applied physics areas within the \texttt{physics} primary category.
 As of mid-2022, this steady decline continues.} Amongst individual fields the largest \emph{relative} decline in a single field is explainable by a high energy lattice physics conference that was cancelled in 2020.
+
 In this study we have only focussed on the quantity of \change{pre-prints} as a measure for productivity, and not their quality.
@@
 
-\change{There were 234 pre-prints posted to the \texttt{q-bio} primary category in 2020 which include pandemic-related terms, and where no author had never appeared on the \texttt{q-bio} pre-print before.
+\change{There were 234 pre-prints posted to the \texttt{q-bio} primary category in 2020 which include pandemic-related terms, and where no author had ever appeared on the \texttt{q-bio} pre-print before.
 We investigated every pre-print to extract the affiliation of the lead author, and used this information to search the internet for the author's website.
@@
 
+\bibitem{Riccaboni}
+Riccaboni, M., Verginer, L.
+{The impact of the COVID-19 pandemic on scientific research in the life sciences}, \emph{PLOS One}, \textbf{17}, 2, 1--16, doi:10.1371/journal.pone.0263001 \newblock (2022).
+
 \bibitem{Fassin:2021}

### 2022-10-31 article/ms.tex
co-author comments

@@
 \newcommand{\todo}[1]{\textcolor{green}{#1}}
-\newcommand{\change}[1]{\textbf{#1}}
+\newcommand{\change}[1]{\textcolor{blue}{\textbf{#1}}}
 
@@
 \author{%%%% Author details
-A. R. Casey$^{1,2}$,  I. Mandel$^{1,3,4}$ and P. K. Ray$^{5}$}
+{A. Casey$^{1,2}$,  I. Mandel$^{1,3,4}$ and P. K. Ray$^{5}$}}
 
@@
 However, we do observe a significant ongoing decline in general physics pre-prints, as well as temporary declines in areas of high-energy physics, where the cancellation of a key conference played a key role.
-The most significant increase occurred in quantitative biology where a 50\%~increase ($+8\sigma$) in pre-prints posted in 2020 was driven by research related to the COVID-19 pandemic. 
+The most significant increase occurred in \change{the \arxiv\ category of} quantitative biology, where a 50\%~increase ($+8\sigma$) in pre-prints posted in 2020 was driven by research related to the COVID-19 pandemic.
 Some of these `COVID papers' are by biologists using arXiv for the first time, but many are written by researchers from other fields (e.g., physicists, mathematicians).
@@
 
-\noindent Peer-reviewed publications are the most common measure of productivity in \change{many specialties of science and technology}.
-The \arxiv\cite{Ginsparg:2011} ({https://arxiv.org}) is a service for the distribution of research manuscripts (pre-prints) prior to their appearance in a journal.
-The presence of a pre-print on \arxiv\ does not ensure that the contents have passed peer-review, but most material on \arxiv\ eventually goes through peer-review because it is now standard in many research specialties to post to \arxiv\ either during or after the peer-review process \cite{Lariviere:2014}.
-\change{The fraction of peer-reviewed publications that are also posted to the \arxiv\ increases each year\cite{Davis:2007} but not all academic outputs (peer reviewed publications or otherwise) will appear on the \arxiv.
-Fundamental reports, findings from inquiries or research groups, and many other texts that are only published in peer-reviewed publications will never appear on the \arxiv.
-In the specialisations we investigate here (physics, mathematics, astronomy, computer science, and other STEM-related fields), the fraction of peer-reviewed literature that also appears on \arxiv\ is estimated to vary from 18.5\% in mathematics to nearly 100\% in high energy physics\cite{Davis:2007,GentilBeccot:2009}.
-Although the fraction may vary between discplines, within a given specialisation the number of pre-prints provides a proxy measure of the number of peer-reviewed publications, which we can use to examine changes in each specialty as a function of time.}
+\change{The \arxiv\ pre-print service\cite{Ginsparg:2011} ({https://arxiv.org}) has become an important platform for disseminating research in a growing number of scientific fields, most notably in physics, mathematics, and computer science.
+However, not all academic outputs will appear on the \arxiv: fundamental reports, findings from inquiries or research groups, and many other texts that are only published in peer-reviewed publications will never appear on the \arxiv, or any other pre-print server.
+In the specialisations we investigate (physics, mathematics, astronomy, computer science, and other STEM-related fields), the fraction of peer-reviewed literature that also appears on the \arxiv\ has been estimated to vary from 18.5\% in mathematics to nearly 100\% in high energy physics\cite{Davis:2007,GentilBeccot:2009}.}
 
@@
 We have instead adopted a broad perspective and present new quantitative results for entire scientific fields.
-The aim is to provide a “large-scale” view of scientific productivity which will serve as a foundation for future studies on the detailed impacts of COVID-19 on academic work.
+The aim is to provide a “large-scale” view of \change{one aspect of} scientific productivity which will serve as a foundation for future studies on the detailed impacts of COVID-19 on academic work.
 
@@
 
-The field of quantitative biology (\texttt{q-bio}) showed the largest increase in pre-prints in 2020 above what is expected by our model, before returning to pre-pandemic levels in mid-2021. % todo: 
+The \change{\arxiv\ category} of quantitative biology (\texttt{q-bio}) showed the largest increase in pre-prints in 2020 above what is expected by our model, before returning to pre-pandemic levels in mid-2021.
+% todo:
 There were {2,790} quantitative biology pre-prints in 2020, {50\%} above the previous year, representing a $+8\sigma$ deviation from our model predictions (predicted 1,914 $\pm$ 109; 876 fewer than observed).
-This increase is explainable by an increase in COVID-19 related research, as there were 864 quantitative biology pre-prints in 2020 with pandemic-related terms in their title or abstract (see Methods), and just 39 in the decade prior.
-Indeed, in {April 2020} nearly 60\% of the \arxiv\ pre-prints in quantitative biology were related to the pandemic (Figure~\ref{fig:2}).
+This increase is explainable by an increase in COVID-19 related research, as there were 864 \change{pre-prints in the quantitative biology \arxiv\ category} in 2020 with pandemic-related terms in their title or abstract (see Methods), and just 39 in the decade prior.
+Indeed, in {April 2020} nearly 60\% of the \change{pre-prints in the \arxiv\ category of quantitative biology} were related to the pandemic (Figure~\ref{fig:2}).
 Pandemic-related pre-prints also appeared in other fields (computer science, physics, statistics, and quantitative finance, all peaking around {April 2020}), but these papers constituted fewer than 10\% of the total pre-prints in these fields.
-Approximately 10--15\% of current pre-prints in quantitative biology are related to the ongoing pandemic.
+Approximately 10--15\% of current pre-prints in \change{the \arxiv\ category of} quantitative biology are related to the ongoing pandemic.
 This represents a strong shift in research focus, but this fraction is declining with time.
@@
 \caption{Pre-prints related to the COVID-19 pandemic peaked in April 2020, accounting for nearly 60\% the quantitative biology (\texttt{q-bio}) pre-prints in that month.
-Shown are five other research fields with the highest number of pandemic-related pre-prints in 2020.
+Shown are five other categories with the highest number of pandemic-related pre-prints in 2020.
 Monthly pre-prints are shown on the left, and the percent of pandemic-related pre-prints in that research field is shown on the right.}
@@
 An increase in quantitative biology\footnote{Quantitative biology is only a sub-field of biology, and most biology pre-prints are posted to bioRxiv ({https://biorxiv.org}).} research during a global pandemic is unsurprising.
-However, the question arises if this increase was driven by existing researchers in the field, or by authors from other fields entering quantitative biology due to the pandemic.
-In 2020 there was a peak in the number of new authors appearing for the first time in the quantitative biology literature (Figure~\ref{fig:3}) while the number of new authors in all other fields remained steady.
-The increase in pre-prints (and new authors) is driven by small (1-4) groups of authors, where many had never posted pre-prints to quantitative biology before.
+However, the question arises if this increase was driven by existing researchers in the field, or by authors from other fields \change{posting to the quantitative biology category} due to the pandemic.
+In 2020 there was a peak in the number of new authors appearing for the first time in the quantitative biology \change{\arxiv\ category} (Figure~\ref{fig:3}) while the number of new authors in all other fields remained steady.
+The increase in pre-prints (and new authors) is driven by small (1-4) groups of authors, where many had never posted pre-prints to \change{the \arxiv\ category of quantitative biology} before.
 
@@
 
-One quarter of the excess pre-prints posted to quantitative biology during 2020 were both related to the pandemic and written by authors who had never appeared in the quantitative biology literature (234 of the 876 excess pre-prints in \texttt{q-bio}).
+One quarter of the excess pre-prints posted to \change{the quantitative biology \arxiv\ category} during 2020 were both related to the pandemic and written by authors who had never appeared in \change{that \arxiv\ category before} (234 of the 876 excess pre-prints in \texttt{q-bio}).
 A careful examination \change{(see Methods)} reveals that about half of these (45\%; 107/234) were written by mathematicians or physicists, with another 26\% (61/234) led by engineers, computer scientists, or economists.
 These pre-prints tend to focus on modelling the COVID-19 outbreak using public data sets, rather than quantitative biology research that requires more expert domain knowledge.
-About 3\% (8/234) were led by established biologists who have not used the \arxiv\ before and are now doing so presumably to help ensure that their COVID-19 research is more widely available.
-This fraction is likely higher among pre-prints where at least one author has appeared before in the quantitative biology literature.
+About 3\% (8/234) were led by established biologists who have not used the \arxiv\ before.
+This fraction is likely higher among pre-prints where at least one author has appeared before in the quantitative biology \change{\arxiv\ category}.
 This represents a genuine increase in new researchers using the \arxiv, and a boost to making quantitative biology research more widely accessible.
@@
 
-\change{The COVID-19 pandemic appears to have had minimal impact on collective research productivity {in the science and technology specialisations we examined}, {and that is before accounting for the tremendous work in scientific communication and contributions to public policy throughout the pandemic.} These findings are at odds with similar studies in fields that have larger research overlap with the SARS-CoV-2 virus.
+The COVID-19 pandemic appears to have had minimal impact on collective research productivity \change{in the science and technology specialisations we examined}, \change{and that is before accounting for the tremendous work in scientific communication and contributions to public policy throughout the pandemic.} \change{These findings are at odds with similar studies in fields that have larger research overlap with the SARS-CoV-2 virus.
 For example, publications in life sciences saw a 6.5 fold increase in subject headings related to COVID-19\cite{Riccaboni}.
-The effects of the pandemic clearly impact research specialisations differently, and we note some} significant changes within specific research areas.
-The largest persistent decline in pre-prints has occurred in {general physics (arXiv primary category \texttt{physics}), largely due to a decline in  pre-prints posted to computational and applied physics areas within the \texttt{physics} primary category.
-As of mid-2022, this steady decline continues.} Amongst individual fields the largest \emph{relative} decline in a single field is explainable by a high energy lattice physics conference that was cancelled in 2020.
+The effects of the pandemic clearly impact research specialisations differently, and we note some significant changes within specific research areas.
+The largest persistent decline in pre-prints has occurred in general physics (arXiv primary category \texttt{physics}), largely due to a decline in  pre-prints posted to computational and applied physics areas within the \texttt{physics} primary category.
+As of mid-2022, this steady decline continues.
+Amongst individual fields the largest \emph{relative} decline in a single field is explainable by a high energy lattice physics conference that was cancelled in 2020.}
 
-In this study we have only focussed on the quantity of \change{pre-prints} as a measure for productivity, and not their quality.
+In this study we have only focussed on the quantity of \change{pre-prints}, and not their quality.
 While citations are the most used measure of impact of academic publications, that metric becomes a more biased statistic when many related pre-prints are all being posted nearly at the same time\cite{Fassin:2021}.
@@
 
-\bibitem{Lariviere:2014}
-{Larivi{\'e}re, V., et al.,} {arXiv E-prints and the journal of record: An analysis of roles and relationships}, \emph{Journal of the Association for Information Science and Technology}, \textbf{65}, 6, doi:10.1002/asi.23044 \newblock (2014).
+%\bibitem{Lariviere:2014}
+%{Larivi{\'e}re, V., et al.,} {arXiv E-prints and the journal of record: An analysis of roles and relationships}, \emph{Journal of the Association for Information Science and Technology}, \textbf{65}, 6, doi:10.1002/asi.23044 \newblock (2014).
 

### 2022-10-31 article/ms.tex
last ref

@@
 
-The COVID-19 pandemic appears to have had minimal impact on collective research productivity \change{in the science and technology specialisations we examined}, \change{and that is before accounting for the tremendous work in scientific communication and contributions to public policy throughout the pandemic.} \change{These findings are at odds with similar studies in fields that have larger research overlap with the SARS-CoV-2 virus.
+The COVID-19 pandemic appears to have had minimal impact on collective research productivity \change{in the science and technology specialisations we examined}, \change{and that is before accounting for the tremendous work in scientific communication and contributions to public policy throughout the pandemic.} \change{These findings are at odds with similar studies in fields that have larger research overlap with the SARS-CoV-2 virus\cite{Fraser}.
 For example, publications in life sciences saw a 6.5 fold increase in subject headings related to COVID-19\cite{Riccaboni}.
@@
 
+\bibitem{Fraser}
+Fraser N, Brierley L, Dey G, Polka JK, Pálfy M, Nanni F, et al., {The evolving role of preprints in the dissemination of COVID-19 research and their impact on the science communication landscape}, \emph{PLOS Biology}, \textbf{19}, 4, doi:10.1371/journal.pbio.3000959 \newblock (2021).
+
 \bibitem{Riccaboni}
